### 2015-09-07 paper/paper.tex
Start on a paper

\documentclass[letterpaper]{article}
\usepackage[pass]{geometry}

\usepackage{ijcai13}
\usepackage{times}

\title{Really Hard Instances of the Subgraph Isomorphism Problem}
\author{Ciaran McCreesh\thanks{This work was supported by the Engineering and Physical Sciences
    Research Council [grant number EP/K503058/1]} \and Patrick Prossser \and James Trimble \\
University of Glasgow, Glasgow, Scotland \\
c.mccreesh.1@research.gla.ac.uk}

\begin{document}

\maketitle

\begin{abstract}
    Really hard SIP instances.
\end{abstract}

\section{Introduction}

\section{Background and Definitions}

\section{Non-Induced Isomorphisms}

\subsection{Varying Densities}

\subsection{Varying Pattern Sizes}

\subsection{Location of the Phase Transition}

\section{Induced Isomorphisms}

\subsection{Varying Densities}

\subsection{Varying Pattern Sizes}

\subsection{SAT Encoding}

\section{Conclusion}

\bibliographystyle{named}
\bibliography{paper}

\end{document}

### 2015-09-07 paper/paper.tex
Write a bit

@@
+% vim: set spell spelllang=en tw=100 :
+
 \documentclass[letterpaper]{article}
@@
 \usepackage{times}
+\usepackage{complexity}
+\usepackage{microtype}
 
-\title{Really Hard Instances of the Subgraph Isomorphism Problem}
+\title{Really Hard Instances for the Subgraph Isomorphism Problem}
 \author{Ciaran McCreesh\thanks{This work was supported by the Engineering and Physical Sciences
@@
 
-\section{Background and Definitions}
+The \emph{non-induced subgraph isomorphism problem} is to find an injective mapping from a given
+pattern graph to a given target graph which preserves adjacency---in essence, we are ``finding a
+copy of'' the pattern inside the target. The \emph{induced} variant of the problem additionally
+requires that the mapping preserve non-adjacency, so there are no ``extra edges'' in the copy of the
+pattern that we find. Despite these problems being \NP-complete, modern practical subgraph
+isomorphism algorithms can handle problem instances with many hundreds of vertices in the pattern
+graph, and up to ten thousand vertices in the target graph
+\cite{Cordella:2004,Solnon:2010,Audemard:2014,McCreesh:2015}.
+
+However, these algorithms cannot handle \emph{arbitrary} instances of this size. The experimental
+evaluations in these papers were performed using a mix of real-world instances (from applications
+??), and random graph pairs.  Using random instances can be beneficial, because it provides a way of
+generating arbitrarily many instances cheaply, and reduces the risk of over-fitting when tuning
+design parameters. The instances used were generated by taking a subgraph of a random (using
+different models) graph and permuting the vertices.  This is not ideal: such instances are
+guaranteed to be satisfiable, and so existing benchmark suites contain relatively few non-trivial
+unsatisfiable instances. This can lead to bias.
+
+Here we present a new method for creating random pattern/target pairs which can generate both
+satisfiable and unsatisfiable instances. For non-induced isomorphisms, as with other hard problems,
+there is a phase transition from satisfiable to unsatisfiable as we alter a parameter, and for see
+the expected complexity peak occur near this phase transition when using three different solvers.
+
+?? Some stuff in there about clique, colouring and SAT.
+
+For induced isomorphisms, there are two phase transitions, and although instances become really hard
+near each transition, there is an even harder region in between the two transitions.
+
+?? Weakness not specific to SIP solvers, use a SAT encoding and it happens too.
 
@@
 
-\subsection{Varying Pattern Sizes}
-
-\subsection{Location of the Phase Transition}
+\subsection{Varying Sizes}
 
 \section{Induced Isomorphisms}
-
-\subsection{Varying Densities}
-
-\subsection{Varying Pattern Sizes}
 

### 2015-09-13 sat/sat_random_subgraph_isomorphism.cc
Simple SAT encoding

@@ -1,0 +1,207 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include <boost/program_options.hpp>
+
+#include <iostream>
+#include <exception>
+#include <cstdlib>
+#include <vector>
+
+namespace po = boost::program_options;
+
+auto create_random_graph(int size, double density, int seed) -> std::vector<std::vector<uint8_t>>
+{
+    std::mt19937 rand;
+    rand.seed(seed);
+    std::uniform_real_distribution<double> dist(0.0, 1.0);
+
+    std::vector<std::vector<uint8_t>> result(size, std::vector<uint8_t>(size, 0));
+
+    for (int e = 0 ; e < size ; ++e) {
+        for (int f = e + 1 ; f < size ; ++f) {
+            if (dist(rand) <= density) {
+                result[e][f] = 1;
+                result[f][e] = 1;
+            }
+        }
+    }
+
+    return result;
+}
+
+auto var(int a, int b, int target_size) -> int
+{
+    return a * target_size + b + 1;
+}
+
+auto main(int argc, char * argv[]) -> int
+{
+    try {
+        po::options_description display_options{ "Program options" };
+        display_options.add_options()
+            ("help",                                  "Display help information")
+            ("induced",                               "Induced version")
+            ;
+
+        po::options_description all_options{ "All options" };
+        all_options.add_options()
+            ("pattern-size",    po::value<int>(),     "Number of pattern vertices")
+            ("pattern-density", po::value<double>(),  "Pattern density")
+            ("pattern-seed",    po::value<int>(),     "Pattern seed")
+            ("target-size",     po::value<int>(),     "Number of target vertices")
+            ("target-density",  po::value<double>(),  "Target density")
+            ("target-seed",     po::value<int>(),     "Target seed")
+            ;
+
+        all_options.add(display_options);
+
+        po::positional_options_description positional_options;
+        positional_options
+            .add("pattern-size", 1)
+            .add("pattern-density", 1)
+            .add("pattern-seed", 1)
+            .add("target-size", 1)
+            .add("target-density", 1)
+            .add("target-seed", 1)
+            ;
+
+        po::variables_map options_vars;
+        po::store(po::command_line_parser(argc, argv)
+                .options(all_options)
+                .positional(positional_options)
+                .run(), options_vars);
+        po::notify(options_vars);
+
+        /* --help? Show a message, and exit. */
+        if (options_vars.count("help")) {
[…]

### 2015-09-14 paper/paper.tex
Start plotting

@@
 \usepackage{microtype}
+\usepackage{gnuplot-lua-tikz}
+\usepackage{amsmath}
+\usepackage{amssymb}
+
+% \usepackage{showframe}
 
@@
 
-\subsection{SAT Encoding}
+\section{Other Encodings}
+
+\begin{figure*}[tb]
+    \input{gen-graph-sat.tex}
+    \caption{Behaviour of other solvers on the induced variant. Each point is the average of ten
+        runs. For each plot, the x-axis is the pattern density and the y-axis is the target
+        density, both from 0 to 1. Along the top row, we show the proportion of instances which are
+        satisfiable; the white bands shows the phase transitions. On the second row, we show the
+        number of search nodes used by the Glasgow algorithm, and on the third row, the number of
+        conflicts recorded by the Glucose SAT solver: the dark regions indicate ``really hard''
+        instances. These results show that instances which are hard for the Glasgow algorithm tend
+        to be hard using other solvers too, but do not otherwise allow comparisons between solver
+        performances.}
+\end{figure*}
 

### 2015-09-14 paper/paper.tex
More plotting

@@
 
-\subsection{Varying Densities}
-
-\subsection{Varying Sizes}
+\begin{figure*}[p]
+    \input{gen-graph-non-induced.tex}
+    \setlength{\abovecaptionskip}{-3em}
+    \caption{Behaviour of algorithms on the non-induced variant. Each point is the average of ten
+        runs. For each plot, the x-axis is the pattern density and the y-axis is the target
+        density, both from 0 to 1. Along the top row, we show the proportion of instances which are
+        satisfiable; the white bands shows the phase transitions. On the second row, we show the
+        number of search nodes used by the Glasgow algorithm, and on the third row, the number of
+        search nodes used by the VF2 algorithm: the dark regions indicate ``really hard''
+        instances.}
+\end{figure*}
 
 \section{Induced Isomorphisms}
+
+\begin{figure*}[p]
+    \input{gen-graph-induced.tex}
+    \setlength{\abovecaptionskip}{-3em}
+    \caption{Behaviour of algorithms on the induced variant. Each point is the average of ten
+        runs. For each plot, the x-axis is the pattern density and the y-axis is the target
+        density, both from 0 to 1. Along the top row, we show the proportion of instances which are
+        satisfiable; the white bands shows the phase transitions. On the second row, we show the
+        number of search nodes used by the Glasgow algorithm, and on the third row, the number of
+        search nodes used by the VF2 algorithm: the dark regions indicate ``really hard''
+        instances.}
+\end{figure*}
 
@@
     \input{gen-graph-sat.tex}
+    \setlength{\abovecaptionskip}{-3em}
     \caption{Behaviour of other solvers on the induced variant. Each point is the average of ten
@@
         number of search nodes used by the Glasgow algorithm, and on the third row, the number of
-        conflicts recorded by the Glucose SAT solver: the dark regions indicate ``really hard''
+        decisions made by the Glucose SAT solver: the dark regions indicate ``really hard''
         instances. These results show that instances which are hard for the Glasgow algorithm tend
-        to be hard using other solvers too, but do not otherwise allow comparisons between solver
-        performances.}
+        to be hard using other solvers too.}
 \end{figure*}

### 2015-09-15 paper/paper.tex
More

@@
 \usepackage{amssymb}
+\usepackage{placeins}
+\usepackage{cleveref}
 
@@
 evaluations in these papers were performed using a mix of real-world instances (from applications
-??), and random graph pairs.  Using random instances can be beneficial, because it provides a way of
-generating arbitrarily many instances cheaply, and reduces the risk of over-fitting when tuning
-design parameters. The instances used were generated by taking a subgraph of a random (using
-different models) graph and permuting the vertices.  This is not ideal: such instances are
-guaranteed to be satisfiable, and so existing benchmark suites contain relatively few non-trivial
-unsatisfiable instances. This can lead to bias.
+??), and randomly generated graph pairs. Using random instances can be beneficial, because it
+provides a way of generating arbitrarily many instances cheaply, and reduces the risk of
+over-fitting when tuning design parameters. The instances used were generated by taking a subgraph
+of a random graph (using various models, including scale-free and bounded degree) and permuting the
+vertices.  This is not ideal: such instances are guaranteed to be satisfiable, and so existing
+benchmark suites contain relatively few non-trivial unsatisfiable instances. This introduces a new
+source of bias when designing algorithms.
 
-Here we present a new method for creating random pattern/target pairs which can generate both
-satisfiable and unsatisfiable instances. For non-induced isomorphisms, as with other hard problems,
-there is a phase transition from satisfiable to unsatisfiable as we alter a parameter, and for see
-the expected complexity peak occur near this phase transition when using three different solvers.
+However, generating non-trivial unsatisfiable random instances is not straightforward. Using a
+pattern graph from one of the random suites with the ``wrong'' target graph tends to give either a
+trivially unsatisfiable instance, or a satisfiable instance. In particular, it is \emph{not} the
+case that a relatively small random graph is very unlikely to appear in a larger random graph.
 
-?? Some stuff in there about clique, colouring and SAT.
+Here we present and evaluate a new method for creating random pattern/target pairs which can
+generate both satisfiable and unsatisfiable instances. Our work builds upon the widely studied phase
+transition phenomena observed in satisfiability and graph colouring problems. However, we have three
+control parameters rather than one: we can independently alter the density of the pattern graph, the
+density of the target graph, and the relative orders (number of vertices) of the pattern and target
+graphs.  For non-induced isomorphisms, with the correct choice of parameters there is a phase
+transition from satisfiable to unsatisfiable, and as expected we see a complexity peak occur (with
+three different solvers) near this phase transition.
 
-For induced isomorphisms, there are two phase transitions, and although instances become really hard
-near each transition, there is an even harder region in between the two transitions.
+For certain choices of parameter for induced isomorphisms, there are two phase transitions, going
+from satisfiable to unsatisfiable, and then from unsatisfiable back to satisfiable. Again, when
+going from satisfiable to unsatisfiable (from either direction), instances go from being trivial to
+really hard to solve. However, each of the three solvers we tested also finds the central
+unsatisfiable region to be really hard---this is surprising, since this region is apparently
+over-constrained. To show that this is not a simple weakness of subgraph isomorphism algorithms, we
+show that the same behaviour occurs when using use a SAT solver, a pseudo-boolean solver, and a
+clique encoding.
 
-?? Weakness not specific to SIP solvers, use a SAT encoding and it happens too.
+\subsection{Definitions}
 
-\section{Non-Induced Isomorphisms}
+?? Some stuff about order, density, random
 
-\begin{figure*}[p]
+\subsection{Experimental Setup}
+
+?? experimental setup.
+
+\section{Non-Induced Subgraph Isomorphisms}
+
+\begin{figure}[h]
     \input{gen-graph-non-induced.tex}
-    \setlength{\abovecaptionskip}{-3em}
+    \setlength{\abovecaptionskip}{-1.5em}
     \caption{Behaviour of algorithms on the non-induced variant. Each point is the average of ten
@@
         instances.}
-\end{figure*}
+    \label{figure:non-induced}
+\end{figure}
+
+In the top row of \cref{figure:non-induced} we show the phase transition for the non-induced
+variant, for patterns of order 10, 20 and 30, a target of order 150, and varying pattern (x-axis)
+and target (y-axis) densities. Inside the orange region, at the bottom right of each plot, every
+instance is unsatisfiable---here we are trying to find a dense pattern in a sparse target. In the
+purple region, at the top left, every instance is satisfiable---we are looking for a sparse pattern
+in a dense target (which is easy, since we only have to preserve adjacency, not non-adjacency). The
+white band between the regions shows the location of the phase transition: here, roughly half the
+instances are satisfiable.
+
+On subsequent rows, we show the average difficulty of different algorithms.
+
+We measure search nodes. Cannot be used to compare performance of algorithms directly, but this does
+show that each solver found the same set of instances difficult.
+
+Satisfiable is easy, until very close to the phase transition. Unsatisfiable is much harder than
+satisfiable.
+
+?? Are unsatisfiable instances on the phase transition much harder than satisfiable ones?
+
+\subsection{Locating the Phase Transition}
 
@@
 
-\begin{figure*}[p]
+\begin{figure*}
     \input{gen-graph-induced.tex}
-    \setlength{\abovecaptionskip}{-3em}
+    \setlength{\abovecaptionskip}{-2em}
     \caption{Behaviour of algorithms on the induced variant. Each point is the average of ten
@@
         search nodes used by the VF2 algorithm: the dark regions indicate ``really hard''
-        instances.}
+        instances.}\label{figure:induced}
 \end{figure*}
+
+\subsection{Locating the Phase Transition}
 
@@
 
-\begin{figure*}[tb]
-    \input{gen-graph-sat.tex}
-    \setlength{\abovecaptionskip}{-3em}
-    \caption{Behaviour of other solvers on the induced variant. Each point is the average of ten
-        runs. For each plot, the x-axis is the pattern density and the y-axis is the target
-        density, both from 0 to 1. Along the top row, we show the proportion of instances which are
-        satisfiable; the white bands shows the phase transitions. On the second row, we show the
-        number of search nodes used by the Glasgow algorithm, and on the third row, the number of
-        decisions made by the Glucose SAT solver: the dark regions indicate ``really hard''
-        instances. These results show that instances which are hard for the Glasgow algorithm tend
-        to be hard using other solvers too.}
+\begin{figure*}
+    \hskip-1.2cm\input{gen-graph-sat.tex}
+    \setlength{\abovecaptionskip}{-2em}
+    \caption{Behaviour of other solvers on the induced variant on smaller graphs, shown in the style of
+        \cref{figure:induced}. The second row shows the number of search nodes used by the Glasgow
+        algorithm, and the third row shows the number of decisions made by the Glucose SAT solver.}
 \end{figure*}

### 2015-09-15 code/clique.cc
Clique encoding, too

@@ -1,0 +1,196 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "clique.hh"
+#include "bit_graph.hh"
+#include "template_voodoo.hh"
+#include "degree_sort.hh"
+
+#include <algorithm>
+#include <limits>
+
+namespace
+{
+    auto modular_product(const Graph & g1, const Graph & g2) -> Graph
+    {
+        Graph result(g1.size() * g2.size());
+
+        for (int v1 = 0 ; v1 < g1.size() ; ++v1)
+            for (int v2 = 0 ; v2 < g2.size() ; ++v2)
+                for (int w1 = 0 ; w1 < g1.size() ; ++w1)
+                    for (int w2 = 0 ; w2 < g2.size() ; ++w2) {
+                        unsigned p1 = v2 * g1.size() + v1;
+                        unsigned p2 = w2 * g1.size() + w1;
+                        if (p1 != p2 && v1 != w1 && v2 != w2 && (g1.adjacent(v1, w1) == g2.adjacent(v2, w2))) {
+                            result.add_edge(p1, p2);
+                        }
+                    }
+        return result;
+    }
+
+    auto noninduced_modular_product(const Graph & g1, const Graph & g2) -> Graph
+    {
+        Graph result(g1.size() * g2.size());
+
+        for (int v1 = 0 ; v1 < g1.size() ; ++v1)
+            for (int v2 = 0 ; v2 < g2.size() ; ++v2)
+                for (int w1 = 0 ; w1 < g1.size() ; ++w1)
+                    for (int w2 = 0 ; w2 < g2.size() ; ++w2) {
+                        unsigned p1 = v2 * g1.size() + v1;
+                        unsigned p2 = w2 * g1.size() + w1;
+                        if (p1 != p2 && v1 != w1 && v2 != w2 && ((! g1.adjacent(v1, w1)) || g2.adjacent(v2, w2))) {
+                            result.add_edge(p1, p2);
+                        }
+                    }
+
+        return result;
+    }
+
+    auto unproduct(const Graph & g1, const Graph &, int v) -> std::pair<int, int>
+    {
+        return std::make_pair(v % g1.size(), v / g1.size());
+    }
+
+    template <unsigned n_words_>
+    struct CliqueSubgraphIsomorphism
+    {
+        const Params & params;
+        const Graph & pattern;
+        const Graph & target;
+
+        FixedBitGraph<n_words_> graph;
+        std::vector<int> order;
+        Result result;
[…]

### 2015-09-15 code/random_subgraph_isomorphism.cc

@@ -1,6 +1,7 @@
 /* vim: set sw=4 sts=4 et foldmethod=syntax : */
 
 #include "sequential.hh"
+#include "clique.hh"
 
 #include <boost/program_options.hpp>
 
@@ -105,6 +106,7 @@ auto main(int argc, char * argv[]) -> int
             ("help",                                  "Display help information")
             ("timeout",            po::value<int>(),  "Abort after this many seconds")
             ("induced",                               "Induced version")
+            ("clique",                                "Use clique algorithm")
             ;
 
         po::options_description all_options{ "All options" };
@@ -163,11 +165,20 @@ auto main(int argc, char * argv[]) -> int
 
         /* Do the actual run. */
         bool aborted = false;
-        auto result = run_this(sequential_subgraph_isomorphism)(
-                graphs,
-                params,
-                aborted,
-                options_vars.count("timeout") ? options_vars["timeout"].as<int>() : 0);
+        Result result;
+
+        if (options_vars.count("clique"))
+            result = run_this(clique_subgraph_isomorphism)(
+                    graphs,
+                    params,
+                    aborted,
+                    options_vars.count("timeout") ? options_vars["timeout"].as<int>() : 0);
+        else
+            result = run_this(sequential_subgraph_isomorphism)(
+                    graphs,
+                    params,
+                    aborted,
+                    options_vars.count("timeout") ? options_vars["timeout"].as<int>() : 0);
 
         /* Stop the clock. */
         auto overall_time = duration_cast<milliseconds>(steady_clock::now() - params.start_time);

### 2015-09-15 code/clique.cc
dynexdegree sorting

@@ -73,7 +73,7 @@ namespace
         {
             // populate our order with every vertex initially
             std::iota(order.begin(), order.end(), 0);
-            degree_sort(g, order, false);
+            dynexdegree_sort(g, order, false);
 
             // re-encode graph as a bit graph
             graph.resize(g.size());

### 2015-09-15 code/degree_sort.cc

@@ -17,3 +17,35 @@ auto degree_sort(const Graph & graph, std::vector<int> & p, bool reverse) -> voi
             [&] (int a, int b) { return (! reverse) ^ (degrees[a] < degrees[b] || (degrees[a] == degrees[b] && a > b)); });
 }
 
+auto dynexdegree_sort(const Graph & graph, std::vector<int> & p, bool reverse) -> void
+{
+    // pre-calculate degrees and exdegrees
+    std::vector<int> degrees;
+    std::transform(p.begin(), p.end(), std::back_inserter(degrees),
+            [&] (int v) { return graph.degree(v); });
+
+    std::vector<int> exdegrees((graph.size()));
+    for (int i = 0 ; i < graph.size() ; ++i)
+        for (int j = 0 ; j < graph.size() ; ++j)
+            if (graph.adjacent(i, j))
+                exdegrees[i] += degrees[j];
+
+    auto unsorted = p.begin(), unsorted_end = p.end();
+
+    while (unsorted != unsorted_end) {
+        // sort on degree
+        std::sort(unsorted, unsorted_end,
+                [&] (int a, int b) { return (! reverse) ^ (
+                    (degrees[a] < degrees[b]) ||
+                    (degrees[a] == degrees[b] && exdegrees[a] < exdegrees[b]) ||
+                    (degrees[a] == degrees[b] && exdegrees[a] == exdegrees[b] && a > b)
+                    ); });
+
+        for (int i = 0 ; i < graph.size() ; ++i)
+            if (graph.adjacent(i, *prev(unsorted_end)))
+                --degrees[i];
+
+        --unsorted_end;
+    }
+}
+

### 2015-09-15 code/clique.cc
Don't dynex

@@ -73,7 +73,7 @@ namespace
         {
             // populate our order with every vertex initially
             std::iota(order.begin(), order.end(), 0);
-            dynexdegree_sort(g, order, false);
+            degree_sort(g, order, false);
 
             // re-encode graph as a bit graph
             graph.resize(g.size());

### 2015-09-15 code/degree_sort.cc

@@ -17,35 +17,3 @@ auto degree_sort(const Graph & graph, std::vector<int> & p, bool reverse) -> voi
             [&] (int a, int b) { return (! reverse) ^ (degrees[a] < degrees[b] || (degrees[a] == degrees[b] && a > b)); });
 }
 
-auto dynexdegree_sort(const Graph & graph, std::vector<int> & p, bool reverse) -> void
-{
-    // pre-calculate degrees and exdegrees
-    std::vector<int> degrees;
-    std::transform(p.begin(), p.end(), std::back_inserter(degrees),
-            [&] (int v) { return graph.degree(v); });
-
-    std::vector<int> exdegrees((graph.size()));
-    for (int i = 0 ; i < graph.size() ; ++i)
-        for (int j = 0 ; j < graph.size() ; ++j)
-            if (graph.adjacent(i, j))
-                exdegrees[i] += degrees[j];
-
-    auto unsorted = p.begin(), unsorted_end = p.end();
-
-    while (unsorted != unsorted_end) {
-        // sort on degree
-        std::sort(unsorted, unsorted_end,
-                [&] (int a, int b) { return (! reverse) ^ (
-                    (degrees[a] < degrees[b]) ||
-                    (degrees[a] == degrees[b] && exdegrees[a] < exdegrees[b]) ||
-                    (degrees[a] == degrees[b] && exdegrees[a] == exdegrees[b] && a > b)
-                    ); });
-
-        for (int i = 0 ; i < graph.size() ; ++i)
-            if (graph.adjacent(i, *prev(unsorted_end)))
-                --degrees[i];
-
-        --unsorted_end;
-    }
-}
-

### 2015-09-15 paper/paper.tex
More picturation

@@
         \cref{figure:induced}. The second row shows the number of search nodes used by the Glasgow
-        algorithm, and the third row shows the number of decisions made by the Glucose SAT solver.}
+        algorithm, the third row shows the number of decisions made by the Glucose SAT solver, and
+        the third row shows the number of search nodes used by a clique encoding.}
 \end{figure*}

### 2015-09-16 paper/paper.tex
Bit more

@@
 \begin{abstract}
-    Really hard SIP instances.
+    Really hard SIP instances. Phase transitions. Alternative models and reduction.
 \end{abstract}
@@
 generate both satisfiable and unsatisfiable instances. Our work builds upon the widely studied phase
-transition phenomena observed in satisfiability and graph colouring problems. However, we have three
-control parameters rather than one: we can independently alter the density of the pattern graph, the
-density of the target graph, and the relative orders (number of vertices) of the pattern and target
-graphs.  For non-induced isomorphisms, with the correct choice of parameters there is a phase
-transition from satisfiable to unsatisfiable, and as expected we see a complexity peak occur (with
-three different solvers) near this phase transition.
+transition phenomena observed in satisfiability and graph colouring problems ?? cite stuff. However,
+we have three control parameters rather than one: we can independently alter the density of the
+pattern graph, the density of the target graph, and the relative orders (number of vertices) of the
+pattern and target graphs.  For non-induced isomorphisms, with the correct choice of parameters
+there is a phase transition from satisfiable to unsatisfiable, and as expected we see a complexity
+peak occur (with three different solvers) near this phase transition.
 
@@
 over-constrained. To show that this is not a simple weakness of subgraph isomorphism algorithms, we
-show that the same behaviour occurs when using use a SAT solver, a pseudo-boolean solver, and a
-clique encoding.
+verify that the same behaviour occurs when using use a SAT solver, a pseudo-boolean solver, and a
+clique encoding. ?? Except maybe it doesn't...
 
@@
 
-?? experimental setup.
+?? Software versions.
+
+Our experiments were performed on systems with Intel Xeon E5-4650 v2 (Q1'14) CPUs and 768GBytes RAM
+(although this much RAM was not needed), running Scientific Linux release 6.7. Software was compiled
+using GCC 4.9.
 
@@
 \begin{figure*}
-    \input{gen-graph-induced.tex}
+    \hskip-1.2cm\input{gen-graph-induced.tex}
     \setlength{\abovecaptionskip}{-2em}
@@
 
-\subsection{Locating the Phase Transition}
+\subsection{Predicting Induced Behaviour}
 
-\section{Other Encodings}
+?? Graph showing the product of satisfiable with vertical flip, min (unsat) and max (sat) of search
+nodes with vertical flip.
+
+\section{Other Encodings and Solvers}
 
@@
 
+?? \cite{Anton:2009}
+
+?? \cite{Lipets:2009}
+
 \section{Conclusion}

### 2015-09-20 lad/lad_random_subgraph_isomorphism.cc
Try LAD too

@@ -1,0 +1,108 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include <boost/program_options.hpp>
+
+#include <iostream>
+#include <exception>
+#include <cstdlib>
+#include <vector>
+
+namespace po = boost::program_options;
+
+auto create_random_graph(int size, double density, int seed) -> std::vector<std::vector<uint8_t>>
+{
+    std::mt19937 rand;
+    rand.seed(seed);
+    std::uniform_real_distribution<double> dist(0.0, 1.0);
+
+    std::vector<std::vector<uint8_t>> result(size, std::vector<uint8_t>(size, 0));
+
+    for (int e = 0 ; e < size ; ++e) {
+        for (int f = e + 1 ; f < size ; ++f) {
+            if (dist(rand) <= density) {
+                result[e][f] = 1;
+                result[f][e] = 1;
+            }
+        }
+    }
+
+    return result;
+}
+
+auto var(int a, int b, int target_size) -> int
+{
+    return a * target_size + b + 1;
+}
+
+auto main(int argc, char * argv[]) -> int
+{
+    try {
+        po::options_description display_options{ "Program options" };
+        display_options.add_options()
+            ("help",                                  "Display help information")
+            ;
+
+        po::options_description all_options{ "All options" };
+        all_options.add_options()
+            ("size",    po::value<int>(),     "Number of vertices")
+            ("density", po::value<double>(),  "Density")
+            ("seed",    po::value<int>(),     "Seed")
+            ;
+
+        all_options.add(display_options);
+
+        po::positional_options_description positional_options;
+        positional_options
+            .add("size", 1)
+            .add("density", 1)
+            .add("seed", 1)
+            ;
+
+        po::variables_map options_vars;
+        po::store(po::command_line_parser(argc, argv)
+                .options(all_options)
+                .positional(positional_options)
+                .run(), options_vars);
+        po::notify(options_vars);
+
+        /* --help? Show a message, and exit. */
+        if (options_vars.count("help")) {
+            std::cout << "Usage: " << argv[0] << " [options] size density seed" << std::endl;
+            std::cout << std::endl;
+            std::cout << display_options << std::endl;
+            return EXIT_SUCCESS;
+        }
+
+        /* No algorithm or no input file specified? Show a message and exit. */
[…]

### 2015-09-21 paper/paper.tex
More results
More
More plotting

@@
 unsatisfiable region to be really hard---this is surprising, since this region is apparently
-over-constrained. To show that this is not a simple weakness of subgraph isomorphism algorithms, we
-verify that the same behaviour occurs when using use a SAT solver, a pseudo-boolean solver, and a
-clique encoding. ?? Except maybe it doesn't...
+over-constrained. ?? Except really it's because solvers can't handle edges and non-edges
+simultaneously. To show that this is not a simple weakness of subgraph isomorphism algorithms, we
+verify that this region is also hard when using use a SAT solver, a pseudo-boolean solver, and a
+clique encoding.
 
@@
 
-?? Some stuff about order, density, random
+?? Some stuff about order, density, random, $\rightarrowtail$ and $\hookrightarrow$.
 
@@
 
-\begin{figure}[h]
-    \input{gen-graph-non-induced.tex}
-    \setlength{\abovecaptionskip}{-1.5em}
+\begin{figure}[tb]
+    \hskip-0.2cm\input{gen-graph-non-induced.tex}
+    \setlength{\abovecaptionskip}{-2.5em}
     \caption{Behaviour of algorithms on the non-induced variant. Each point is the average of ten
@@
         satisfiable; the white bands shows the phase transitions. On the second row, we show the
-        number of search nodes used by the Glasgow algorithm, and on the third row, the number of
-        search nodes used by the VF2 algorithm: the dark regions indicate ``really hard''
-        instances.}
+        number of search nodes used by the Glasgow algorithm, on the third row, the number of
+        search nodes used by the LAD algorithm, and the fourth, VF2; the dark regions indicate
+        ``really hard'' instances.}
     \label{figure:non-induced}
@@
 
-On subsequent rows, we show the average difficulty of different algorithms.
-
-We measure search nodes. Cannot be used to compare performance of algorithms directly, but this does
-show that each solver found the same set of instances difficult.
+On subsequent rows, we show the average difficulty of different algorithms.  We measure search
+nodes. Cannot be used to compare performance of algorithms directly, but this does show that each
+solver found the same set of instances difficult.
 
 Satisfiable is easy, until very close to the phase transition. Unsatisfiable is much harder than
-satisfiable.
+satisfiable. This is solver-independent, although VF2 is terrible.
 
@@
 
+?? Can we predict where that line is?
+
 \section{Induced Isomorphisms}
 
-\begin{figure*}
+In \cref{figure:induced} we show the induced behaviour (modified Glasgow to include the complement
+as a supplemental graph).
+
+Pattern size 10, what we expect. Patterns 20 and 30 have large unsatisfiable regions in the middle,
+but these remain hard. Paattern sizes 14, 15 and 16 show the transition between the two styles.
+
+Odd unsat with LAD at the top right (not EHPs), trivial for Glasgow: mapping a non-clique into a
+clique. Glasgow looks at non-edges and can see immediately that no mapping is possible.
+
+Note that this is solver-independent.
+
+\begin{figure*}[tb]
     \hskip-1.2cm\input{gen-graph-induced.tex}
     \setlength{\abovecaptionskip}{-2em}
-    \caption{Behaviour of algorithms on the induced variant. Each point is the average of ten
-        runs. For each plot, the x-axis is the pattern density and the y-axis is the target
-        density, both from 0 to 1. Along the top row, we show the proportion of instances which are
-        satisfiable; the white bands shows the phase transitions. On the second row, we show the
-        number of search nodes used by the Glasgow algorithm, and on the third row, the number of
-        search nodes used by the VF2 algorithm: the dark regions indicate ``really hard''
-        instances.}\label{figure:induced}
+    \caption{Behaviour of algorithms on the induced variant, shown in the style
+    of \cref{figure:non-induced}.}\label{figure:induced}
 \end{figure*}
@@
 
+This suggests that we're not able to make use of adjacency and non-adjacency simultaneously.
+
 \section{Other Encodings and Solvers}
 
-\begin{figure*}
+\begin{figure*}[tb]
     \hskip-1.2cm\input{gen-graph-sat.tex}
@@
     \caption{Behaviour of other solvers on the induced variant on smaller graphs, shown in the style of
-        \cref{figure:induced}. The second row shows the number of search nodes used by the Glasgow
+        \cref{figure:non-induced}. The second row shows the number of search nodes used by the Glasgow
         algorithm, the third row shows the number of decisions made by the Glucose SAT solver, and
-        the third row shows the number of search nodes used by a clique encoding.}
+    the fourth row shows the number of search nodes used by a clique encoding.}\label{figure:alt}
 \end{figure*}
+
+Is this just a weakness of current subgraph isomorphism solvers? In \cref{figure:alt} we repeat the
+experiments on smaller pattern and target graphs, using different solving techniques. Although these
+techniques are not competitive in absolute terms, we wish to see if the same pattern of behaviour
+occurs.
+
+Direct SAT encoding: variables for each assignment, cardinality constraints, edge and non-edge
+constraints. We use the Glucose SAT solver. We see that hard instances remain hard, including inside
+the central region, and that the easy satisfiable instances remain easy.
+
+Clique encoding: association graph, a version of BBMC modified to solve the decision problem. The
+hard instances in the central region remain hard. Additionally, some of the easy satisfiable
+instances become hard. Possible weakness of clique algorithms for this kind of input.
+
+?? Pseudo-boolean, incomplete solvers?
 

### 2015-09-22 paper/paper.tex
More
A phase transition

@@
 \begin{figure}[tb]
+    \input{gen-graph-phase-transition.tex}
+    \caption{With a fixed a pattern size of 20, a target size of 150, a target density of 0.4, and
+    varying pattern density, we observe a phase transition and complexity peak in the non-induced
+    variant. Each point represents one instance.}
+    \label{figure:phase-transition}
+\end{figure}
+
+Arbitrarily decide upon a pattern size of 20, a target size of 150, and a fixed target density of
+0.40. \cref{figure:phase-transition}.
+
+\begin{figure}[tb]
     \hskip-0.2cm\input{gen-graph-non-induced.tex}
@@
     \hskip-1.2cm\input{gen-graph-induced.tex}
-    \setlength{\abovecaptionskip}{-2em}
+    \setlength{\abovecaptionskip}{-3em}
     \caption{Behaviour of algorithms on the induced variant, shown in the style
@@
     \hskip-1.2cm\input{gen-graph-sat.tex}
-    \setlength{\abovecaptionskip}{-2em}
+    \setlength{\abovecaptionskip}{-3em}
     \caption{Behaviour of other solvers on the induced variant on smaller graphs, shown in the style of

### 2015-09-29 pb/clasp-3.1.3/libclasp/tests/program_builder_test.cpp
Add clasp

@@ -1,0 +1,2588 @@
+// 
+// Copyright (c) 2006, Benjamin Kaufmann
+// 
+// This file is part of Clasp. See http://www.cs.uni-potsdam.de/clasp/ 
+// 
+// Clasp is free software; you can redistribute it and/or modify
+// it under the terms of the GNU General Public License as published by
+// the Free Software Foundation; either version 2 of the License, or
+// (at your option) any later version.
+// 
+// Clasp is distributed in the hope that it will be useful,
+// but WITHOUT ANY WARRANTY; without even the implied warranty of
+// MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+// GNU General Public License for more details.
+// 
+// You should have received a copy of the GNU General Public License
+// along with Clasp; if not, write to the Free Software
+// Foundation, Inc., 51 Franklin St, Fifth Floor, Boston, MA  02110-1301  USA
+//
+#include "test.h"
+#include <clasp/solver.h>
+#include <clasp/minimize_constraint.h>
+#include <clasp/unfounded_check.h>
+
+#include <sstream>
+using namespace std;
+namespace Clasp { namespace Test {
+using namespace Clasp::Asp;
+struct ClauseObserver : public DecisionHeuristic {
+	Literal doSelect(Solver&){return Literal();}
+	void updateVar(const Solver&, Var, uint32) {}
+	void newConstraint(const Solver&, const Literal* first, LitVec::size_type size, ConstraintType) {
+		Clause c;
+		for (LitVec::size_type i = 0; i < size; ++i, ++first) {
+			c.push_back(*first);
+		}
+		std::sort(c.begin(), c.end());
+		clauses_.push_back(c);
+	}
+	typedef std::vector<Literal> Clause;
+	typedef std::vector<Clause> Clauses;
+	Clauses clauses_;
+};
+
[…]

### 2015-09-29 pb/si_pb.py

@@ -1,0 +1,74 @@
+import itertools
+from random import random
+import sys
+
+def read_graph(filename):
+	i = -1
+	for line in open(filename, "r"):
+		if i == -1:
+			n = int(line)
+			g = [[False]*n for _ in range(n)]
+		else:
+			tokens = map(int, line.split())
+			for j in range(1, tokens[0]+1):
+				g[i][tokens[j]] = True
+		i += 1
+	return g
+
+def show_graph(g):
+       for row in g:
+               print ["1 " if b else "0 " for b in row]
+
+def si(p, t, induced=False):
+	np = len(p)
+	nt = len(t)
+
+	constraints = []
+
+	# Each pattern vertex maps to exactly one target vertex
+	for i in range(np):
+		constraints.append(
+			" ".join("+1 x{}".format(i*nt + j + 1) for j in range(nt)) + \
+			"= 1;")
+
+	# Each target vertex maps to at most one pattern vertex
+	for i in range(nt):
+		constraints.append(
+			" ".join("-1 x{}".format(j*nt + i + 1) for j in range(np)) + \
+			">= -1;")
+
+	# If v maps to w, then for every neighbour u of v,
+	# u must map to a neighbour of w in T
+	for v in range(np):
+		for w in range(nt):
+			for u in range(np):
+				if p[u][v]:
+					constraints.append("-1 x{} ".format(v*nt+w+1) \
+						+ " ".join("+1 x{}".format(u*nt+x+1) for x in range(nt) if t[w][x]) \
+						+ " >= 0;")
+
+	# Induced equivalent
+	if induced:
+		for v in range(np):
+			for w in range(nt):
+				for u in range(np):
+					if u!=v and not p[u][v]:
+						constraints.append("-1 x{} ".format(v*nt+w+1) \
+							+ " ".join("+1 x{}".format(u*nt+x+1) for x in range(nt) if not t[w][x]) \
+							+ " >= 0;")
+
+	constraints = set(constraints)
+	print "* #variable= {} #constraint= {}".format(nt*np, len(constraints))
+	for c in constraints:
+		print c
+
+if __name__=="__main__":
+    if len(sys.argv) < 3:
+        print "Usage: python {} pattern_file target_file [induced]".format(sys.argv[0])
+    else:
+        p_file = sys.argv[1]
+        t_file = sys.argv[2]
+        induced = len(sys.argv) > 3 and sys.argv[3][0].lower()=="i"
[…]

### 2015-09-29 mip/si_mip_cplex.py
mip

@@ -1,0 +1,82 @@
+import itertools
+import sys
+
+def read_graph(filename):
+	i = -1
+	for line in open(filename, "r"):
+		if i == -1:
+			n = int(line)
+			g = [[False]*n for _ in range(n)]
+		else:
+			tokens = map(int, line.split())
+			for j in range(1, tokens[0]+1):
+				g[i][tokens[j]] = True
+		i += 1
+	return g
+
+def show_graph(g):
+       for row in g:
+               print ["1 " if b else "0 " for b in row]
+
+def si(p, t, induced=False):
+	np = len(p)
+	nt = len(t)
+
+	var_names = ["p{}t{}".format(i, j) for i in range(np) for j in range(nt)] 
+
+	print "Maximize"
+	print " obj: + " + " +  ".join(var_names)
+
+	print
+	print "Subject To"
+
+	# Each pattern vertex maps to exactly one target vertex
+	for i in range(np):
+		print " p{}_to_at_most_1_vtx:".format(i), \
+			" + ".join("p{}t{}".format(i, j) for j in range(nt)), \
+			" = 1"
+
+	# Each target vertex maps to at most one pattern vertex
+	print
+	for i in range(nt):
+		print " t{}_to_at_most_1_vtx:".format(i), \
+			" + ".join("p{}t{}".format(j, i) for j in range(np)), \
+			" <= 1"
+
+	# If v maps to w, then for every neighbour u of v,
+	# u must map to a neighbour of w in T
+	print
+	for v in range(np):
+		for w in range(nt):
+			for u in range(np):
+				if p[u][v]:
+					print " c_{}_{}_{}: + p{}t{} ".format(v, w, u, v, w) \
+						+ " ".join("-p{}t{}".format(u, x) for x in range(nt) if t[w][x]) \
+						+ " <= 0"
+
+	# Induced equivalent
+	if induced:
+		print
+		for v in range(np):
+			for w in range(nt):
+				for u in range(np):
+					if u!=v and not p[u][v]:
+						print " ci_{}_{}_{}: + p{}t{} ".format(v, w, u, v, w) \
+							+ " ".join("-p{}t{}".format(u, x) for x in range(nt) if not t[w][x]) \
+							+ " <= 0"
+
+	print 
+	print "Binary"
+	print " " + " ".join(var_names)
+	print "End"
+
+if __name__=="__main__":
+    if len(sys.argv) < 3:
[…]

### 2015-10-19 data-presentation-talk/talk.tex
Experimental talk

% vim: set spell spelllang=en tw=100 et sw=4 sts=4 foldmethod=marker foldmarker={{{,}}} :

\documentclass{beamer}

\usepackage{tikz}
\usepackage{xcolor}
\usepackage{complexity}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage{amsmath}                   % \operatorname
\usepackage{amsfonts}                  % \mathcal
\usepackage{amssymb}                   % \nexists
\usepackage{gnuplot-lua-tikz}          % graphs
\usepackage[vlined]{algorithm2e} % algorithms
\usepackage{centernot}
\usepackage{mathtools}
\usepackage{listings}

\usetikzlibrary{shapes, arrows, shadows, calc, positioning, fit}
\usetikzlibrary{decorations.pathreplacing, decorations.pathmorphing, shapes.misc}
\usetikzlibrary{tikzmark}

\definecolor{uofguniversityblue}{rgb}{0, 0.219608, 0.396078}

\definecolor{uofgheather}{rgb}{0.356863, 0.32549, 0.490196}
\definecolor{uofgaquamarine}{rgb}{0.603922, 0.72549, 0.678431}
\definecolor{uofgslate}{rgb}{0.309804, 0.34902, 0.380392}
\definecolor{uofgrose}{rgb}{0.823529, 0.470588, 0.709804}
\definecolor{uofgmocha}{rgb}{0.709804, 0.564706, 0.47451}
\definecolor{uofgsandstone}{rgb}{0.321569, 0.278431, 0.231373}
\definecolor{uofgforest}{rgb}{0, 0.2, 0.129412}
\definecolor{uofglawn}{rgb}{0.517647, 0.741176, 0}
\definecolor{uofgcobalt}{rgb}{0, 0.615686, 0.92549}
\definecolor{uofgturquoise}{rgb}{0, 0.709804, 0.819608}
\definecolor{uofgsunshine}{rgb}{1.0, 0.862745, 0.211765}
\definecolor{uofgpumpkin}{rgb}{1.0, 0.72549, 0.282353}
\definecolor{uofgthistle}{rgb}{0.584314, 0.070588, 0.447059}
\definecolor{uofgrust}{rgb}{0.603922, 0.227451, 0.023529}
\definecolor{uofgburgundy}{rgb}{0.490196, 0.133333, 0.223529}
\definecolor{uofgpillarbox}{rgb}{0.701961, 0.047059, 0}
\definecolor{uofglavendar}{rgb}{0.356863, 0.301961, 0.580392}

\tikzset{vertex/.style={draw, circle, inner sep=0pt, minimum size=0.5cm, font=\small\bfseries}}
\tikzset{notvertex/.style={vertex, color=white, text=black}}
\tikzset{plainvertex/.style={vertex}}
\tikzset{vertexc1/.style={vertex, fill=uofgburgundy, text=white}}
\tikzset{vertexc2/.style={vertex, fill=uofgsandstone, text=white}}
\tikzset{vertexc3/.style={vertex, fill=uofgforest, text=white}}
\tikzset{vertexc4/.style={vertex, fill=uofgheather, text=white}}
\tikzset{edge/.style={color=black!50!white}}
\tikzset{bedge/.style={ultra thick}}
\tikzset{edged/.style={color=screengrey, dashed}}
\tikzset{edgel3/.style={color=uofgrose, ultra thick}}

% {{{ theme things
\useoutertheme[footline=authortitle]{miniframes}
\useinnertheme{rectangles}

\setbeamerfont{block title}{size={}}
\setbeamerfont{title}{size=\large,series=\bfseries}
\setbeamerfont{section title}{size=\large,series=\mdseries}
\setbeamerfont{author}{size=\normalsize,series=\mdseries}
\setbeamercolor*{structure}{fg=uofguniversityblue}
\setbeamercolor*{palette primary}{use=structure,fg=black,bg=white}
\setbeamercolor*{palette secondary}{use=structure,fg=white,bg=uofgcobalt}
\setbeamercolor*{palette tertiary}{use=structure,fg=white,bg=uofguniversityblue}
\setbeamercolor*{palette quaternary}{fg=white,bg=black}

\setbeamercolor*{titlelike}{parent=palette primary}

\beamertemplatenavigationsymbolsempty

\setbeamertemplate{title page}
{
    \begin{tikzpicture}[remember picture, overlay]
        \node at (current page.north west) {
            \begin{tikzpicture}[remember picture, overlay]
                \fill [fill=uofguniversityblue, anchor=north west] (0, 0) rectangle (\paperwidth, -2.6cm);
            \end{tikzpicture}
        };

        \node (logo) [anchor=north east, shift={(-0.6cm,-0.6cm)}] at (current page.north east) {
            \includegraphics*[keepaspectratio=true,scale=0.7]{UoG_keyline.pdf}
        };

        \node [anchor=west, xshift=0.2cm] at (current page.west |- logo.west) {
            \begin{minipage}{0.65\paperwidth}\raggedright
                {\usebeamerfont{title}\usebeamercolor[white]{}\inserttitle}\\[0.1cm]
                {\usebeamerfont{author}\usebeamercolor[white]{}\insertauthor}
            \end{minipage}
        };
    \end{tikzpicture}
}

\setbeamertemplate{section page}
{
    \begin{centering}
        \begin{beamercolorbox}[sep=12pt,center]{part title}
            \usebeamerfont{section title}\insertsection\par
        \end{beamercolorbox}
    \end{centering}
}

\newcommand{\frameofframes}{/}
\newcommand{\setframeofframes}[1]{\renewcommand{\frameofframes}{#1}}

\makeatletter
\setbeamertemplate{footline}
{%
    \begin{beamercolorbox}[colsep=1.5pt]{upper separation line foot}
    \end{beamercolorbox}
    \begin{beamercolorbox}[ht=2.5ex,dp=1.125ex,%
        leftskip=.3cm,rightskip=.3cm plus1fil]{author in head/foot}%
        \leavevmode{\usebeamerfont{author in head/foot}\insertshortauthor}%
        \hfill%
        {\usebeamerfont{institute in head/foot}\usebeamercolor[fg]{institute in head/foot}\insertshortinstitute}%
    \end{beamercolorbox}%
    \begin{beamercolorbox}[ht=2.5ex,dp=1.125ex,%
        leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
        {\usebeamerfont{title in head/foot}\insertshorttitle}%
        \hfill%
        {\usebeamerfont{frame number}\usebeamercolor[fg]{frame number}\insertframenumber~\frameofframes~\inserttotalframenumber}
    \end{beamercolorbox}%
    \begin{beamercolorbox}[colsep=1.5pt]{lower separation line foot}
    \end{beamercolorbox}
}

% }}}

\title[When is Finding Little Graphs Inside Big Graphs Hard?]{When is Finding Little Graphs \\ Inside Big Graphs Hard?}
\author[Ciaran McCreesh and Patrick Prosser]{\textbf{Ciaran McCreesh} and Patrick Prosser}

\begin{document}

{
    \usebackgroundtemplate{
        \tikz[overlay, remember picture]
        \node[at=(current page.south), anchor=south, inner sep=0pt]{\includegraphics*[keepaspectratio=true, width=\paperwidth]{background.jpg}};
    }
    \begin{frame}[plain,noframenumbering]
        \titlepage
    \end{frame}
}

\begin{frame}{Non-Induced Subgraph Isomorphism}

    \begin{center}\begin{tikzpicture}
        \node[draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (Na) at (1,  0) {1};
        \node[draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (Nb) at (1, -1.5) {2};
        \node[draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (Nc) at (0, -3) {3};
        \node[draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (Nd) at (2, -3) {4};

        \draw [ultra thick] (Na) -- (Nb);
        \draw [ultra thick] (Nb) -- (Nc);
        \draw [ultra thick] (Nc) -- (Nd);
        \draw [ultra thick] (Nb) -- (Nd);

        \node[draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (N1) at (5.5,  0) {1};
        \node[draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N2) at (7.5,  0) {2};
        \node[draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (N3) at (5.5, -1.5) {3};
        \node[draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N4) at (7.5, -1.5) {4};
        \node[draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (N5) at (5.5, -3) {5};
        \node[draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (N6) at (7.5, -3) {6};

        \draw [thick, color=uofgsandstone!50] (N1) -- (N2);
        \draw [ultra thick] (N1) -- (N3);
        \draw [thick, color=uofgsandstone!50] (N1) -- (N4);
        \draw [thick, color=uofgsandstone!50] (N2) -- (N4);
        \draw [ultra thick] (N3) -- (N5);
        \draw [ultra thick] (N3) -- (N6);
        \draw [thick, color=uofgsandstone!50] (N4) -- (N6);
        \draw [ultra thick] (N5) -- (N6);
        \draw [thick, color=uofgsandstone!50] (N2) to [in=45, out=315] (N6);
        \draw [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);
    \end{tikzpicture}\end{center}

\end{frame}

\begin{frame}{Induced Subgraph Isomorphism}

    \begin{center}\begin{tikzpicture}
        \node[draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (Na) at (1,  0) {1};
        \node[draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (Nb) at (1, -1.5) {2};
        \node[draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (Nc) at (0, -3) {3};
        \node[draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (Nd) at (2, -3) {4};

        \draw [ultra thick] (Na) -- (Nb);
        \draw [ultra thick] (Nb) -- (Nc);
        \draw [ultra thick] (Nc) -- (Nd);
        \draw [ultra thick] (Nb) -- (Nd);

        \node[draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N1) at (5.5,  0) {1};
        \node[draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (N2) at (7.5,  0) {2};
        \node[draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (N3) at (5.5, -1.5) {3};
        \node[draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N4) at (7.5, -1.5) {4};
        \node[draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (N5) at (5.5, -3) {5};
        \node[draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (N6) at (7.5, -3) {6};

        \draw [thick, color=uofgsandstone!50] (N1) -- (N2);
        \draw [thick, color=uofgsandstone!50] (N1) -- (N3);
        \draw [thick, color=uofgsandstone!50] (N1) -- (N4);
        \draw [thick, color=uofgsandstone!50] (N2) -- (N4);
        \draw [ultra thick] (N3) -- (N5);
        \draw [ultra thick] (N3) -- (N6);
        \draw [thick, color=uofgsandstone!50] (N4) -- (N6);
        \draw [ultra thick] (N5) -- (N6);
        \draw [ultra thick] (N2) to [in=45, out=315] (N6);
        \draw [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);

    \end{tikzpicture}\end{center}
\end{frame}

\begin{frame}{Benchmarking}

    \begin{itemize}
        \item Based upon chemical and computer vision datasets, we can handle patterns with 1,000
            vertices and targets with 10,000 vertices.
        \item Do these results reflect the worst case, or are they too optimistic?
        \item Can we create ``hard'' benchmark instances?
    \end{itemize}

\end{frame}

\begin{frame}{Randomly Selected Subgraphs}

    \begin{itemize}
        \item Start with a random target graph.
        \item Pick vertices at random to make a pattern.
        \item Shuffle the numbering.
        \item <2-> These instances will always be satisfiable!
    \end{itemize}

\end{frame}

\begin{frame}{Independently Random Subgraphs}

    \begin{itemize}
        \item Make a random target graph.
        \item Independently, make a random pattern graph.
        \item <2-> Will these instances ever be satisfiable?
    \end{itemize}

\end{frame}

\begin{frame}{A Phase Transition}
\end{frame}

\begin{frame}{2D Phase Transitions}
\end{frame}

\begin{frame}{3D Phase Transitions}
\end{frame}

\begin{frame}{Reductions}
\end{frame}

\begin{frame}[plain,noframenumbering]
    \begin{tikzpicture}[remember picture, overlay]
        \node at (current page.north west) {
            \begin{tikzpicture}[remember picture, overlay]
                \fill [fill=uofguniversityblue, anchor=north west] (0, 0) rectangle (\paperwidth, -1.7cm);
            \end{tikzpicture}
        };

        \node (logo) [anchor=north east, shift={(-0.3cm,-0.2cm)}] at (current page.north east) {
            \includegraphics*[keepaspectratio=true,scale=0.55]{UoG_keyline.pdf}
        };
    \end{tikzpicture}
\end{frame}

\end{document}

### 2015-10-20 data-presentation-talk/talk.tex
Talk

@@
                 {\usebeamerfont{title}\usebeamercolor[white]{}\inserttitle}\\[0.1cm]
-                {\usebeamerfont{author}\usebeamercolor[white]{}\insertauthor}
+                {\usebeamerfont{author}\usebeamercolor[white]{}\scriptsize\insertauthor}
             \end{minipage}
@@
 \title[When is Finding Little Graphs Inside Big Graphs Hard?]{When is Finding Little Graphs \\ Inside Big Graphs Hard?}
-\author[Ciaran McCreesh and Patrick Prosser]{\textbf{Ciaran McCreesh} and Patrick Prosser}
+\author[Ciaran McCreesh, Patrick Prosser and James Trimble]{\textbf{Ciaran McCreesh}, Patrick Prosser and James Trimble}
 
@@
 
-\begin{frame}{A Phase Transition}
+\begin{frame}[plain,t]{A Phase Transition}
+
+    \only<1>{
+        \input{gen-graph-phase-1}
+    }
+
+    \only<2>{
+        \input{gen-graph-phase-2}
+    }
+
+    \only<3>{
+        \input{gen-graph-phase-3}
+    }
+
+\end{frame}
+
+\begin{frame}{This Looks Familiar\ldots}
+    \centering
+    \includegraphics*[keepaspectratio=true,scale=0.25]{sat.jpg}
+
+    \par\flushleft
+    Understanding the Empirical Hardness of NP-Complete Problems. Kevin Leyton-Brown, Holger H.
+    Hoos, Frank Hutter, Lin Xu.  Communications of the ACM, Vol. 57 No. 5, Pages 98-107
 \end{frame}
@@
 \begin{frame}{2D Phase Transitions}
+
+    \only<1>{
+        \input{gen-graph-2d-1}
+    }
+
+    \only<2>{
+        \input{gen-graph-2d-2}
+    }
+
+    \only<3>{
+        \input{gen-graph-2d-3}
+    }
+
 \end{frame}
 
-\begin{frame}{3D Phase Transitions}
+\begin{frame}[plain,t]{3D Phase Transitions?}
+
+    \only<1>{
+        \input{gen-graph-3d-1}
+    }
+
+    \only<2>{
+        \input{gen-graph-3d-2}
+    }
+
+\end{frame}
+
+\begin{frame}[t,plain]{What About Induced?}
+
+    \only<1>{
+        \input{gen-graph-induced-2d-1}
+    }
+
+    \only<2>{
+        \input{gen-graph-induced-2d-2}
+    }
+
+    \only<3>{
+        \input{gen-graph-induced-2d-3}
+    }
+
+    \only<4>{
+        \input{gen-graph-induced-2d-4}
+    }
+
+    \only<5>{
+        \input{gen-graph-induced-2d-5}
+    }
+
+    \only<6>{
+        \input{gen-graph-induced-3d-1}
+    }
+
+    \only<7>{
+        \input{gen-graph-induced-3d-2}
+    }
+
+\end{frame}
+
+\begin{frame}{Other Solvers?}
+
+    \only<1>{
+        \input{gen-graph-alt-1}
+    }
+
+    \only<2>{
+        \input{gen-graph-alt-2}
+    }
+
+    \only<3>{
+        \input{gen-graph-alt-3}
+    }
+
 \end{frame}
@@
 \begin{frame}{Reductions}
+
+    \only<1>{
+        \input{gen-graph-red-1}
+    }
+
+    \only<2>{
+        \input{gen-graph-red-2}
+    }
+
+    \only<3>{
+        \input{gen-graph-red-3}
+    }
+
+    \only<4>{
+        \input{gen-graph-red-4}
+    }
+
+    \only<5>{
+        \input{gen-graph-red-5}
+    }
+
+\end{frame}
+
+\begin{frame}{Why the Weird Shape?}
+
+    \only<1>{
+        \input{gen-graph-shape-1}
+    }
+
+    \only<2>{
+        \input{gen-graph-shape-2}
+    }
+
+    \only<3>{
+        \input{gen-graph-shape-3}
+    }
+
+    \only<4>{
+        \input{gen-graph-shape-4}
+    }
+
+    \only<5>{
+        \input{gen-graph-shape-5}
+    }
+
 \end{frame}

### 2015-10-20 data-presentation-talk/talk.tex
A bit more
Tweaks

@@
     \begin{center}\begin{tikzpicture}
-        \node[draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (Na) at (1,  0) {1};
-        \node[draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (Nb) at (1, -1.5) {2};
-        \node[draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (Nc) at (0, -3) {3};
-        \node[draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (Nd) at (2, -3) {4};
+        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Na) at (1,  0) {1};
+        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Nb) at (1, -1.5) {2};
+        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Nc) at (0, -3) {3};
+        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Nd) at (2, -3) {4};
 
-        \draw [ultra thick] (Na) -- (Nb);
-        \draw [ultra thick] (Nb) -- (Nc);
-        \draw [ultra thick] (Nc) -- (Nd);
-        \draw [ultra thick] (Nb) -- (Nd);
+        \node <2> [draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (Na) at (1,  0) {1};
+        \node <2> [draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (Nb) at (1, -1.5) {2};
+        \node <2> [draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (Nc) at (0, -3) {3};
+        \node <2> [draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (Nd) at (2, -3) {4};
 
-        \node[draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (N1) at (5.5,  0) {1};
-        \node[draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N2) at (7.5,  0) {2};
-        \node[draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (N3) at (5.5, -1.5) {3};
-        \node[draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N4) at (7.5, -1.5) {4};
-        \node[draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (N5) at (5.5, -3) {5};
-        \node[draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (N6) at (7.5, -3) {6};
+        \draw <1> [thick, color=uofgsandstone!50] (Na) -- (Nb);
+        \draw <1> [thick, color=uofgsandstone!50] (Nb) -- (Nc);
+        \draw <1> [thick, color=uofgsandstone!50] (Nc) -- (Nd);
+        \draw <1> [thick, color=uofgsandstone!50] (Nb) -- (Nd);
 
-        \draw [thick, color=uofgsandstone!50] (N1) -- (N2);
-        \draw [ultra thick] (N1) -- (N3);
-        \draw [thick, color=uofgsandstone!50] (N1) -- (N4);
-        \draw [thick, color=uofgsandstone!50] (N2) -- (N4);
-        \draw [ultra thick] (N3) -- (N5);
-        \draw [ultra thick] (N3) -- (N6);
-        \draw [thick, color=uofgsandstone!50] (N4) -- (N6);
-        \draw [ultra thick] (N5) -- (N6);
-        \draw [thick, color=uofgsandstone!50] (N2) to [in=45, out=315] (N6);
-        \draw [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);
+        \draw <2> [ultra thick] (Na) -- (Nb);
+        \draw <2> [ultra thick] (Nb) -- (Nc);
+        \draw <2> [ultra thick] (Nc) -- (Nd);
+        \draw <2> [ultra thick] (Nb) -- (Nd);
+
+        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N1) at (5.5,  0) {1};
+        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N2) at (7.5,  0) {2};
+        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N3) at (5.5, -1.5) {3};
+        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N4) at (7.5, -1.5) {4};
+        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N5) at (5.5, -3) {5};
+        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N6) at (7.5, -3) {6};
+
+        \node <2> [draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (N1) at (5.5,  0) {1};
+        \node <2> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N2) at (7.5,  0) {2};
+        \node <2> [draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (N3) at (5.5, -1.5) {3};
+        \node <2> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N4) at (7.5, -1.5) {4};
+        \node <2> [draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (N5) at (5.5, -3) {5};
+        \node <2> [draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (N6) at (7.5, -3) {6};
+
+        \draw <1> [thick, color=uofgsandstone!50] (N1) -- (N2);
+        \draw <1> [thick, color=uofgsandstone!50] (N1) -- (N3);
+        \draw <1> [thick, color=uofgsandstone!50] (N1) -- (N4);
+        \draw <1> [thick, color=uofgsandstone!50] (N2) -- (N4);
+        \draw <1> [thick, color=uofgsandstone!50] (N3) -- (N5);
+        \draw <1> [thick, color=uofgsandstone!50] (N3) -- (N6);
+        \draw <1> [thick, color=uofgsandstone!50] (N4) -- (N6);
+        \draw <1> [thick, color=uofgsandstone!50] (N5) -- (N6);
+        \draw <1> [thick, color=uofgsandstone!50] (N2) to [in=45, out=315] (N6);
+        \draw <1> [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);
+
+        \draw <2> [thick, color=uofgsandstone!50] (N1) -- (N2);
+        \draw <2> [ultra thick] (N1) -- (N3);
+        \draw <2> [thick, color=uofgsandstone!50] (N1) -- (N4);
+        \draw <2> [thick, color=uofgsandstone!50] (N2) -- (N4);
+        \draw <2> [ultra thick] (N3) -- (N5);
+        \draw <2> [ultra thick] (N3) -- (N6);
+        \draw <2> [thick, color=uofgsandstone!50] (N4) -- (N6);
+        \draw <2> [ultra thick] (N5) -- (N6);
+        \draw <2> [thick, color=uofgsandstone!50] (N2) to [in=45, out=315] (N6);
+        \draw <2> [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);
     \end{tikzpicture}\end{center}
@@
     \begin{center}\begin{tikzpicture}
-        \node[draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (Na) at (1,  0) {1};
-        \node[draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (Nb) at (1, -1.5) {2};
-        \node[draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (Nc) at (0, -3) {3};
-        \node[draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (Nd) at (2, -3) {4};
+        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Na) at (1,  0) {1};
+        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Nb) at (1, -1.5) {2};
+        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Nc) at (0, -3) {3};
+        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Nd) at (2, -3) {4};
 
-        \draw [ultra thick] (Na) -- (Nb);
-        \draw [ultra thick] (Nb) -- (Nc);
-        \draw [ultra thick] (Nc) -- (Nd);
-        \draw [ultra thick] (Nb) -- (Nd);
+        \node <2> [draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (Na) at (1,  0) {1};
+        \node <2> [draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (Nb) at (1, -1.5) {2};
+        \node <2> [draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (Nc) at (0, -3) {3};
+        \node <2> [draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (Nd) at (2, -3) {4};
 
-        \node[draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N1) at (5.5,  0) {1};
-        \node[draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (N2) at (7.5,  0) {2};
-        \node[draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (N3) at (5.5, -1.5) {3};
-        \node[draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N4) at (7.5, -1.5) {4};
-        \node[draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (N5) at (5.5, -3) {5};
-        \node[draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (N6) at (7.5, -3) {6};
+        \draw <1> [thick, color=uofgsandstone!50] (Na) -- (Nb);
+        \draw <1> [thick, color=uofgsandstone!50] (Nb) -- (Nc);
+        \draw <1> [thick, color=uofgsandstone!50] (Nc) -- (Nd);
+        \draw <1> [thick, color=uofgsandstone!50] (Nb) -- (Nd);
 
-        \draw [thick, color=uofgsandstone!50] (N1) -- (N2);
-        \draw [thick, color=uofgsandstone!50] (N1) -- (N3);
-        \draw [thick, color=uofgsandstone!50] (N1) -- (N4);
-        \draw [thick, color=uofgsandstone!50] (N2) -- (N4);
-        \draw [ultra thick] (N3) -- (N5);
-        \draw [ultra thick] (N3) -- (N6);
-        \draw [thick, color=uofgsandstone!50] (N4) -- (N6);
-        \draw [ultra thick] (N5) -- (N6);
-        \draw [ultra thick] (N2) to [in=45, out=315] (N6);
-        \draw [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);
+        \draw <2> [ultra thick] (Na) -- (Nb);
+        \draw <2> [ultra thick] (Nb) -- (Nc);
+        \draw <2> [ultra thick] (Nc) -- (Nd);
+        \draw <2> [ultra thick] (Nb) -- (Nd);
 
+        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N1) at (5.5,  0) {1};
+        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N2) at (7.5,  0) {2};
+        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N3) at (5.5, -1.5) {3};
+        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N4) at (7.5, -1.5) {4};
+        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N5) at (5.5, -3) {5};
+        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N6) at (7.5, -3) {6};
+
+        \node <2> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N1) at (5.5,  0) {1};
+        \node <2> [draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (N2) at (7.5,  0) {2};
+        \node <2> [draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (N3) at (5.5, -1.5) {3};
+        \node <2> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N4) at (7.5, -1.5) {4};
+        \node <2> [draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (N5) at (5.5, -3) {5};
+        \node <2> [draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (N6) at (7.5, -3) {6};
+
+        \draw <1> [thick, color=uofgsandstone!50] (N1) -- (N2);
+        \draw <1> [thick, color=uofgsandstone!50] (N1) -- (N3);
+        \draw <1> [thick, color=uofgsandstone!50] (N1) -- (N4);
+        \draw <1> [thick, color=uofgsandstone!50] (N2) -- (N4);
+        \draw <1> [thick, color=uofgsandstone!50] (N3) -- (N5);
+        \draw <1> [thick, color=uofgsandstone!50] (N3) -- (N6);
+        \draw <1> [thick, color=uofgsandstone!50] (N4) -- (N6);
+        \draw <1> [thick, color=uofgsandstone!50] (N5) -- (N6);
+        \draw <1> [thick, color=uofgsandstone!50] (N2) to [in=45, out=315] (N6);
+        \draw <1> [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);
+
+        \draw <2> [thick, color=uofgsandstone!50] (N1) -- (N2);
+        \draw <2> [thick, color=uofgsandstone!50] (N1) -- (N3);
+        \draw <2> [thick, color=uofgsandstone!50] (N1) -- (N4);
+        \draw <2> [thick, color=uofgsandstone!50] (N2) -- (N4);
+        \draw <2> [ultra thick] (N3) -- (N5);
+        \draw <2> [ultra thick] (N3) -- (N6);
+        \draw <2> [thick, color=uofgsandstone!50] (N4) -- (N6);
+        \draw <2> [ultra thick] (N5) -- (N6);
+        \draw <2> [ultra thick] (N2) to [in=45, out=315] (N6);
+        \draw <2> [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);
     \end{tikzpicture}\end{center}
@@
 
-\begin{frame}[t,plain]{What About Induced?}
+\begin{frame}{What About Induced?}
 
@@
     }
+\end{frame}
 
-    \only<6>{
+\begin{frame}[plain,t]{Induced in 3D?}
+
+    \only<1>{
         \input{gen-graph-induced-3d-1}
@@
 
-    \only<7>{
+    \only<2>{
         \input{gen-graph-induced-3d-2}
@@
 
+    \only<6>{
+        \input{gen-graph-shape-6}
+    }
+
 \end{frame}

### 2015-10-20 data-presentation-talk/talk.tex
nearlyplainframe

@@
 \setbeamercolor*{structure}{fg=uofguniversityblue}
-\setbeamercolor*{palette primary}{use=structure,fg=black,bg=white}
+\setbeamercolor*{palette primary}{use=structure,fg=black,bg=}
 \setbeamercolor*{palette secondary}{use=structure,fg=white,bg=uofgcobalt}
@@
     \end{beamercolorbox}
+}
+\makeatother
+
+\newenvironment{nearlyplainframe}[2][]{
+    \makeatletter
+    \def\beamer@entrycode{\vspace*{-\headheight}}
+
+    \setbeamertemplate{headline}
+    {%
+        \begin{beamercolorbox}[colsep=1.5pt]{upper separation line head}
+        \end{beamercolorbox}
+        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
+            leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
+        \end{beamercolorbox}%
+        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
+            leftskip=.3cm,rightskip=.3cm plus1fil]{author in head/foot}%
+        \end{beamercolorbox}%
+        \begin{beamercolorbox}[colsep=1.5pt]{lower separation line head}
+        \end{beamercolorbox}
+        \vspace*{\headheight}
+    }
+
+    \setbeamertemplate{footline}
+    {%
+        \begin{beamercolorbox}[colsep=1.5pt]{upper separation line foot}
+        \end{beamercolorbox}
+        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
+            leftskip=.3cm,rightskip=.3cm plus1fil]{author in head/foot}%
+        \end{beamercolorbox}%
+        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
+            leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
+        \end{beamercolorbox}%
+        \begin{beamercolorbox}[colsep=1.5pt]{lower separation line foot}
+        \end{beamercolorbox}
+    }
+    \makeatother
+
+    \begin{frame}[#1]{#2}
+    }{
+    \end{frame}
 }
@@
 
-\begin{frame}[plain,t]{A Phase Transition}
-
+\begin{nearlyplainframe}[t]{A Phase Transition}
     \only<1>{
@@
 
-\end{frame}
+\end{nearlyplainframe}
 
@@
 
-\begin{frame}[plain,t]{3D Phase Transitions?}
+\begin{nearlyplainframe}[t]{3D Phase Transitions?}
 
@@
 
-\begin{frame}[plain,t]{Induced in 3D?}
+\begin{nearlyplainframe}[t]{Induced in 3D?}
 

### 2015-11-12 code/random_subgraph_isomorphism.cc
Density model

@@ -12,6 +12,7 @@
 #include <thread>
 #include <mutex>
 #include <condition_variable>
+#include <algorithm>
 
 namespace po = boost::program_options;
 
@@ -19,19 +20,32 @@ using std::chrono::steady_clock;
 using std::chrono::duration_cast;
 using std::chrono::milliseconds;
 
-auto create_random_graph(int size, double density, int seed) -> Graph
+auto create_random_graph(int size, double density, int seed, bool edgemodel) -> Graph
 {
     std::mt19937 rand;
     rand.seed(seed);
-    std::uniform_real_distribution<double> dist(0.0, 1.0);
 
     Graph result;
     result.resize(size);
 
-    for (int e = 0 ; e < size ; ++e) {
-        for (int f = e + 1 ; f < size ; ++f) {
-            if (dist(rand) <= density)
-                result.add_edge(e, f);
+    if (edgemodel) {
+        std::vector<std::pair<int, int> > edges;
+        for (int e = 0 ; e < size ; ++e)
+            for (int f = e + 1 ; f < size ; ++f)
+                edges.emplace_back(e, f);
+
+        std::shuffle(edges.begin(), edges.end(), rand);
+
+        for (int i = 0 ; i < (size * (size - 1) * density / 2) ; ++i)
+            result.add_edge(edges.at(i).first, edges.at(i).second);
+    }
+    else {
+        std::uniform_real_distribution<double> dist(0.0, 1.0);
+        for (int e = 0 ; e < size ; ++e) {
+            for (int f = e + 1 ; f < size ; ++f) {
+                if (dist(rand) <= density)
+                    result.add_edge(e, f);
+            }
         }
     }
 
@@ -107,6 +121,7 @@ auto main(int argc, char * argv[]) -> int
             ("timeout",            po::value<int>(),  "Abort after this many seconds")
             ("induced",                               "Induced version")
             ("clique",                                "Use clique algorithm")
+            ("density",                               "Generate using density instead of probability")
             ;
 
         po::options_description all_options{ "All options" };
@@ -160,8 +175,10 @@ auto main(int argc, char * argv[]) -> int
 
         /* Create graphs */
         auto graphs = std::make_pair(
-                create_random_graph(options_vars["pattern-size"].as<int>(), options_vars["pattern-density"].as<double>(), options_vars["pattern-seed"].as<int>()),
-                create_random_graph(options_vars["target-size"].as<int>(), options_vars["target-density"].as<double>(), options_vars["target-seed"].as<int>()));
[…]

### 2015-11-14 prediction/induced.rb
Prediction

@@ -1,0 +1,41 @@
+#!/usr/bin/ruby
+# vim: set sw=4 sts=4 et tw=80 :
+
+require 'bigdecimal'
+
+ps = ARGV[0].to_i
+ts = ARGV[1].to_i
+
+0.upto(50) do | y |
+    0.upto(50) do | x |
+        pd = BigDecimal.new(x) / 50;
+        td = BigDecimal.new(y) / 50;
+
+        cliqueedges = BigDecimal.new(ps * (ps - 1) / 2);
+        nedges = pd * cliqueedges;
+        nnonedges = (1 - pd) * cliqueedges;
+
+        edgesokprob = td ** nedges;
+        nonedgesokprob = (1 - td) ** nnonedges;
+
+        solprob = edgesokprob * nonedgesokprob;
+
+        nstates = BigDecimal.new(1);
+        1.upto(ps) do | n |
+            nstates *= (ts - n + 1);
+        end
+
+        expsols = nstates * solprob;
+
+        if (expsols < 1)
+            expsols = 0
+        elsif (expsols > 1)
+            expsols = 1
+        else
+            expsols = 0.5;
+        end
+
+        print expsols.to_s + " "
+    end
+    puts
+end

### 2015-11-14 prediction/noninduced.rb

@@ -1,0 +1,39 @@
+#!/usr/bin/ruby
+# vim: set sw=4 sts=4 et tw=80 :
+
+require 'bigdecimal'
+
+ps = ARGV[0].to_i
+ts = ARGV[1].to_i
+
+0.upto(50) do | y |
+    0.upto(50) do | x |
+        pd = BigDecimal.new(x) / 50;
+        td = BigDecimal.new(y) / 50;
+
+        cliqueedges = BigDecimal.new(ps * (ps - 1) / 2);
+        nedges = pd * cliqueedges;
+
+        edgesokprob = td ** nedges;
+
+        solprob = edgesokprob;
+
+        nstates = BigDecimal.new(1);
+        1.upto(ps) do | n |
+            nstates *= (ts - n + 1);
+        end
+
+        expsols = nstates * solprob;
+
+        if (expsols < 1)
+            expsols = 0
+        elsif (expsols > 1)
+            expsols = 1
+        else
+            expsols = 0.5;
+        end
+
+        print expsols.to_s + " "
+    end
+    puts
+end

### 2015-11-14 paper/paper.tex
More stuff
Predictions and layout changes

--- 2015-09-21 paper/paper.tex
@@
 \usepackage{cleveref}
+\usepackage{natbib}
 
 % \usepackage{showframe}
+\usepackage{lipsum}
+
+\usetikzlibrary{decorations, decorations.pathreplacing, calc, backgrounds}
+
+\definecolor{uofguniversityblue}{rgb}{0, 0.219608, 0.396078}
+
+\definecolor{uofgheather}{rgb}{0.356863, 0.32549, 0.490196}
+\definecolor{uofgaquamarine}{rgb}{0.603922, 0.72549, 0.678431}
+\definecolor{uofgslate}{rgb}{0.309804, 0.34902, 0.380392}
+\definecolor{uofgrose}{rgb}{0.823529, 0.470588, 0.709804}
+\definecolor{uofgmocha}{rgb}{0.709804, 0.564706, 0.47451}
+\definecolor{uofgsandstone}{rgb}{0.321569, 0.278431, 0.231373}
+\definecolor{uofgforest}{rgb}{0, 0.2, 0.129412}
+\definecolor{uofglawn}{rgb}{0.517647, 0.741176, 0}
+\definecolor{uofgcobalt}{rgb}{0, 0.615686, 0.92549}
+\definecolor{uofgturquoise}{rgb}{0, 0.709804, 0.819608}
+\definecolor{uofgsunshine}{rgb}{1.0, 0.862745, 0.211765}
+\definecolor{uofgpumpkin}{rgb}{1.0, 0.72549, 0.282353}
+\definecolor{uofgthistle}{rgb}{0.584314, 0.070588, 0.447059}
+\definecolor{uofgrust}{rgb}{0.603922, 0.227451, 0.023529}
+\definecolor{uofgburgundy}{rgb}{0.490196, 0.133333, 0.223529}
+\definecolor{uofgpillarbox}{rgb}{0.701961, 0.047059, 0}
+\definecolor{uofglavendar}{rgb}{0.356863, 0.301961, 0.580392}
+
+\renewcommand{\topfraction}{0.9}
+\renewcommand{\bottomfraction}{0.8}
 
@@
 \begin{abstract}
-    Really hard SIP instances. Phase transitions. Alternative models and reduction.
+    We show how to generate ``really hard'' random instances for variants of the subgraph
+    isomorphism problem. For the non-induced variant of the problem, we predict and observe a phase
+    transition between satisfiable and unsatisfiable instances, with a corresponding complexity peak
+    seen in three different solvers. For the induced variant of the problem, the behaviour is more
+    complicated, with two phase transitions and an unexpectedly hard region in between. We show that
+    this region remains hard when using pseudo-boolean and mixed integer encodings, and under
+    reduction to the clique problem.
 \end{abstract}
@@
 requires that the mapping preserve non-adjacency, so there are no ``extra edges'' in the copy of the
-pattern that we find. Despite these problems being \NP-complete, modern practical subgraph
-isomorphism algorithms can handle problem instances with many hundreds of vertices in the pattern
-graph, and up to ten thousand vertices in the target graph
-\cite{Cordella:2004,Solnon:2010,Audemard:2014,McCreesh:2015}.
+pattern that we find. We illustrate both variants in \cref{figure:sip}.
+Despite these problems being \NP-complete, modern practical subgraph isomorphism algorithms can
+handle problem instances with many hundreds of vertices in the pattern graph, and up to ten thousand
+vertices in the target graph \citep{Cordella:2004,Solnon:2010,Audemard:2014,McCreesh:2015}.
 
 However, these algorithms cannot handle \emph{arbitrary} instances of this size. The experimental
-evaluations in these papers were performed using a mix of real-world instances (from applications
-??), and randomly generated graph pairs. Using random instances can be beneficial, because it
-provides a way of generating arbitrarily many instances cheaply, and reduces the risk of
-over-fitting when tuning design parameters. The instances used were generated by taking a subgraph
-of a random graph (using various models, including scale-free and bounded degree) and permuting the
-vertices.  This is not ideal: such instances are guaranteed to be satisfiable, and so existing
-benchmark suites contain relatively few non-trivial unsatisfiable instances. This introduces a new
-source of bias when designing algorithms.
+evaluations of these algorithms were performed using a mix of real-world graphs, graphs that encode
+biochemistry and computer vision problems, and randomly generated graph pairs. Using random
+instances to evaluate algorithm behaviour can be beneficial, because it provides a way of generating
+many instances cheaply, and reduces the risk of over-fitting when tuning design parameters. The
+random instances used in each case came from common datasets (cite ??), which were generated by
+taking a random subgraph of a random graph (using various models, including scale-free and bounded
+degree) and permuting the vertices.  This is not ideal: such instances are guaranteed to be
+satisfiable, and so existing benchmark suites contain relatively few non-trivial unsatisfiable
+instances. These instances also tend to be computationally fairly easy, with most of the difficulty
+being in dealing with the size of the model.
 
@@
 pattern graph from one of the random suites with the ``wrong'' target graph tends to give either a
-trivially unsatisfiable instance, or a satisfiable instance. In particular, it is \emph{not} the
-case that a relatively small random graph is very unlikely to appear in a larger random graph.
+trivially unsatisfiable instance, or a satisfiable instance. (In particular, it is \emph{not} the
+case that a relatively small random graph is unlikely to appear in a larger random graph.)
 
-Here we present and evaluate a new method for creating random pattern/target pairs which can
-generate both satisfiable and unsatisfiable instances. Our work builds upon the widely studied phase
-transition phenomena observed in satisfiability and graph colouring problems ?? cite stuff. However,
-we have three control parameters rather than one: we can independently alter the density of the
-pattern graph, the density of the target graph, and the relative orders (number of vertices) of the
-pattern and target graphs.  For non-induced isomorphisms, with the correct choice of parameters
-there is a phase transition from satisfiable to unsatisfiable, and as expected we see a complexity
-peak occur (with three different solvers) near this phase transition.
+\begin{figure}[b]
+    \centering
+    \begin{tikzpicture}[scale=0.5]
+        \node [draw, circle, fill=uofgcobalt, inner sep=1.5pt] (Na) at (1,  0) {1};
+        \node [draw, circle, fill=uofgpumpkin, inner sep=1.5pt] (Nb) at (1, -1.5) {2};
+        \node [draw, circle, fill=uofglawn, inner sep=1.5pt] (Nc) at (0, -3) {3};
+        \node [draw, circle, fill=uofgthistle, inner sep=1.5pt] (Nd) at (2, -3) {4};
 
-For certain choices of parameter for induced isomorphisms, there are two phase transitions, going
+        \draw [ultra thick] (Na) -- (Nb);
+        \draw [ultra thick] (Nb) -- (Nc);
+        \draw [ultra thick] (Nc) -- (Nd);
+        \draw [ultra thick] (Nb) -- (Nd);
+
+        \node [draw, circle, fill=uofgcobalt, inner sep=1.5pt] (N1) at (5.5,  0) {1};
+        \node [draw, circle, fill=white, inner sep=1.5pt] (N2) at (7.5,  0) {2};
+        \node [draw, circle, fill=uofgpumpkin, inner sep=1.5pt] (N3) at (5.5, -1.5) {3};
+        \node [draw, circle, fill=white, inner sep=1.5pt] (N4) at (7.5, -1.5) {4};
+        \node [draw, circle, fill=uofglawn, inner sep=1.5pt] (N5) at (5.5, -3) {5};
+        \node [draw, circle, fill=uofgthistle, inner sep=1.5pt] (N6) at (7.5, -3) {6};
+
+        \draw [thick, color=uofgsandstone!50] (N1) -- (N2);
+        \draw [ultra thick] (N1) -- (N3);
+        \draw [thick, color=uofgsandstone!50] (N1) -- (N4);
+        \draw [thick, color=uofgsandstone!50] (N2) -- (N4);
+        \draw [ultra thick] (N3) -- (N5);
+        \draw [ultra thick] (N3) -- (N6);
+        \draw [thick, color=uofgsandstone!50] (N4) -- (N6);
+        \draw [ultra thick] (N5) -- (N6);
+        \draw [thick, color=uofgsandstone!50] (N2) to [in=45, out=315] (N6);
+        \draw [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);
+
+        \node [draw, circle, fill=white, inner sep=1.5pt] (M1) at (-5.5,  0) {1};
+        \node [draw, circle, fill=uofgcobalt, inner sep=1.5pt] (M2) at (-3.5,  0) {2};
+        \node [draw, circle, fill=uofgthistle, inner sep=1.5pt] (M3) at (-5.5, -1.5) {3};
+        \node [draw, circle, fill=white, inner sep=1.5pt] (M4) at (-3.5, -1.5) {4};
+        \node [draw, circle, fill=uofglawn, inner sep=1.5pt] (M5) at (-5.5, -3) {5};
+        \node [draw, circle, fill=uofgpumpkin, inner sep=1.5pt] (M6) at (-3.5, -3) {6};
+
+        \draw [thick, color=uofgsandstone!50] (M1) -- (M2);
+        \draw [thick, color=uofgsandstone!50] (M1) -- (M3);
+        \draw [thick, color=uofgsandstone!50] (M1) -- (M4);
+        \draw [thick, color=uofgsandstone!50] (M2) -- (M4);
+        \draw [ultra thick] (M3) -- (M5);
+        \draw [ultra thick] (M3) -- (M6);
+        \draw [thick, color=uofgsandstone!50] (M4) -- (M6);
+        \draw [ultra thick] (M5) -- (M6);
+        \draw [ultra thick] (M2) to [in=45, out=315] (M6);
+        \draw [thick, color=uofgsandstone!50] (M1) to [in=135, out=225] (M5);
+
+        \node [anchor=center, font=\Large] (A1) at (-1.5, -1.5) { $\hookleftarrow$ };
+        \node [anchor=center, font=\Large] (A2) at ( 3.5, -1.5) { $\rightarrowtail$ };
+    \end{tikzpicture}
+
+    \caption{There is a non-induced isomorphism from the pattern graph, in the center, to the target
+    on the right, mapping vertex 1 to 1, 2 to 3, 3 to 5 and 4 to 6. This is not an induced
+    isomorphism, since there is an edge between 1 and 5 in the target but not between 1 and 3 in the
+    pattern. The mapping from the pattern to the (same) target on the left, sending 1 to 2, 2 to 6, 3 to
+    5 and 4 to 3, is both non-induced and induced.}
+    \label{figure:sip}
+\end{figure}
+
+Here we present and evaluate a new method for creating random pattern/target pairs. This method
+generates both satisfiable and unsatisfiable instances, and can produce computationally challenging
+instances with only a few tens of vertices in the pattern, and 150 vertices in the target. Our work
+builds upon the phase transition phenomena observed in satisfiability and graph colouring problems
+first described by \citet{Cheeseman:1991} and \citet{Mitchell:1992}---?? cite later stuff and
+explain really hard.
+
+For subgraph isomorphism we have three relevant control parameters: we can independently alter the
+edge probability of the pattern graph, the edge probability of the target graph, and the relative
+orders (number of vertices) of the pattern and target graphs.  For non-induced isomorphisms, with
+the correct choice of parameters we see results very similar to those observed with boolean
+satisfiability problems: there is a phase transition from satisfiable to unsatisfiable (and we can
+predict its location), and we see a complexity peak occur (with three different solvers) near this
+phase transition.
+
+For certain choices of parameters for induced isomorphisms, there are two phase transitions, going
 from satisfiable to unsatisfiable, and then from unsatisfiable back to satisfiable. Again, when
@@
 really hard to solve. However, each of the three solvers we tested also finds the central
-unsatisfiable region to be really hard---this is surprising, since this region is apparently
-over-constrained. ?? Except really it's because solvers can't handle edges and non-edges
-simultaneously. To show that this is not a simple weakness of subgraph isomorphism algorithms, we
-verify that this region is also hard when using use a SAT solver, a pseudo-boolean solver, and a
-clique encoding.
+unsatisfiable region to be really hard---this may be unexpected, since this region is not near a
+phase transition and is apparently over-constrained. To show that this is not a simple weakness of
+current subgraph isomorphism algorithms, we verify that this region is also hard when using use a
+pseudo-boolean encoding, a mixed integer model, and under reduction to the clique problem.
 
@@
 
-?? Some stuff about order, density, random, $\rightarrowtail$ and $\hookrightarrow$.
+Throughout, our graphs are unlabelled, undirected, and do not have any loops (vertices which are
+adjacent to themselves). We write $\operatorname{V}(G)$ for the vertex set of a graph $G$. A
+\emph{non-induced subgraph isomorphism} from a graph $P$ (called the \emph{pattern}) to a graph $T$
+(called the \emph{target}) is an injective mapping from $\operatorname{V}(P)$ to
+$\operatorname{V}(T)$ which preserves adjacency---that is, for every adjacent $v$ and $w$ in
+$\operatorname{V}(P)$, the vertices $i(v)$ and $i(w)$ are adjacent in $T$. An \emph{induced subgraph
+isomorphism} additionally preserves non-adjacency---that is, if $v$ and $w$ are not adjacent in $P$,
+then $i(v)$ and $i(w)$ must not be adjacent in $T$. We use the notation $i : P \rightarrowtail T$
+for a non-induced isomorphism, and $i : P \hookrightarrow T$ for an induced isomorphism.
+
+The \emph{order} of a graph is the cardinality of its vertex set.  We write $G(v, p)$ for a random
+graph with $v$ vertices, and an edge between each distinct pair of vertices with independent
+probability $p$.
 
@@
 
-?? Software versions.
+Our experiments were performed on systems with Intel Xeon E5-4650 v2 (Q1'14) CPUs and 768GBytes RAM
+(although this much RAM was not needed), running Scientific Linux release 6.7.   ?? Software
+versions. Clasp 3.1.3, Gurobi 6.0.5, Glasgow, LAD version 2, VFLib. Software was compiled using GCC
+4.9.
 
-Our experiments were performed on systems with Intel Xeon E5-4650 v2 (Q1'14) CPUs and 768GBytes RAM
-(although this much RAM was not needed), running Scientific Linux release 6.7. Software was compiled
-using GCC 4.9.
+?? Stuff about runtimes and recursive calls. We are not aiming to compare solvers; rather, we are
+looking for solver-independent behaviour.
 
@@
 
+\begin{figure}[t]
+    \input{gen-graph-phase-transition.tex}
+    \caption{With a fixed a pattern graph order of 20, a target graph order of 150, a target edge
+        probability of 0.4, and varying pattern edge probability, we observe a phase transition and
+        complexity peak with the Glasgow solver in the non-induced variant. Each point represents
+        one instance. The lines show mean search effort, and mean proportion satisfiable, using a
+    larger sample size.}
+    \label{figure:phase-transition}
+\end{figure}
+
+Suppose we arbitrarily decide upon a pattern graph order of 20, a target graph order of 150, and a
+fixed target edge probability of 0.40. As we vary the pattern edge probability from 0 to 1, we would
+expect to see a shift from entirely satisfiable instances (with no edges in the pattern, we can
+always find a match) to entirely unsatisfiable instances (a maximum clique in this order and edge
+probability of target graph will usually have between 9 and 12 vertices). Indeed,
+\cref{figure:phase-transition} shows that this is the case: the line (and the points show a subset
+of the samples) ?? describe this a bit more, find the 0.5 SAT mark.
+
+Some stuff about the complexity peak in the Glasgow solver. This looks remarkably similar to random
+3SAT problems---compare, for example, Figure 1 of \citet{LeytonBrown:2014}. In particular,
+satisfiable instances tend to be easier but show greater variation than unsatisfiable instances, and
+there are exceptionally hard satisfiable instances. (The parallel version of the Glasgow solver is
+designed to eliminate these.)
+
 \begin{figure}[tb]
-    \hskip-0.2cm\input{gen-graph-non-induced.tex}
-    \setlength{\abovecaptionskip}{-2.5em}
+    \input{gen-graph-non-induced.tex}
     \caption{Behaviour of algorithms on the non-induced variant. Each point is the average of ten
-        runs. For each plot, the x-axis is the pattern density and the y-axis is the target
-        density, both from 0 to 1. Along the top row, we show the proportion of instances which are
-        satisfiable; the white bands shows the phase transitions. On the second row, we show the
-        number of search nodes used by the Glasgow algorithm, on the third row, the number of
-        search nodes used by the LAD algorithm, and the fourth, VF2; the dark regions indicate
-        ``really hard'' instances.}
+        runs. For each plot, the x-axis is the pattern edge probability and the y-axis is the target
+        edge probability, both from 0 to 1. Along the top row, we show the proportion of instances which are
+        satisfiable; the white bands shows the phase transitions. The second row is our prediction
+        of what the first row will look like. On the third row, we show the number of search nodes
+        used by the Glasgow algorithm, on the fourth row, the number of search nodes used by the LAD
+        algorithm, and the fifth, VF2; the dark regions indicate ``really hard'' instances.}
     \label{figure:non-induced}
@@
 
+In figure \cref{figure:phase-transition-bands} we show what happens for larger patterns.
+
+\begin{figure}[tb]
+    \input{gen-graph-phase-transition-bands.tex}
+    \caption{The location of the phase transition, as the pattern grows. Shown are points where
+    between 20\% and 80\% of instances are satisfiable.}
+    \label{figure:phase-transition-bands}
+\end{figure}
+
 \subsection{Locating the Phase Transition}
 
-?? Can we predict where that line is?
+To predict the location of the phase transition, we can calculate the expected number of solutions
+by making a few dubious assumptions involving rounding and independence. Since we are trying to find
+an \emph{injective} mapping from a pattern $P = G(p, d_p)$ to a target $T = G(t, d_t)$, there are
+$t^{\underline{p}} = t \cdot (t - 1)\cdots(t - p + 1)$ possible assignments of target vertices to
+pattern vertices.  We expect the pattern to have $d_p \cdot \binom{p}{2}$ edges, so we obtain the
+probability of all of these edges being mapped to edges in the target by raising $d_t$ to this
+power, giving an expected number of solutions of \[ t^{\underline{p}} \cdot {d_t}^{d_p \cdot
+\binom{p}{2}} \textnormal{,} \] which we plot in ?? somewhere.
 
@@
 
-Pattern size 10, what we expect. Patterns 20 and 30 have large unsatisfiable regions in the middle,
-but these remain hard. Paattern sizes 14, 15 and 16 show the transition between the two styles.
+Pattern order 10, what we expect. Patterns 20 and 30 have large unsatisfiable regions in the middle,
+but these remain hard. Pattern orders 14, 15 and 16 show the transition between the two styles.
 
@@
 
-\begin{figure*}[tb]
-    \hskip-1.2cm\input{gen-graph-induced.tex}
-    \setlength{\abovecaptionskip}{-2em}
+\begin{figure*}[h]
+    \input{gen-graph-induced.tex}
     \caption{Behaviour of algorithms on the induced variant, shown in the style
-    of \cref{figure:non-induced}.}\label{figure:induced}
+    of \cref{figure:non-induced}. The second row shows a bound on the satisfiable region, by
+    considering where a \emph{non-}induced isomorphism may also be a non-induced isomorphism between
+    complement graphs.}\label{figure:induced}
 \end{figure*}
@@
 \subsection{Predicting Induced Behaviour}
+
+By repeating the argument for locating the non-induced phase transition and additionally considering
+non-edges, we get an expected number of solutions of \[t^{\underline{p}} \cdot {d_t}^{d_p \cdot
+\binom{p}{2}} \cdot {d_{\overline{t}}}^{d_{\overline{p}} \cdot \binom{p}{2}} \textnormal{,} \]
+where $d_{\overline{p}} = 1 - d_p$ and $d_{\overline{t}} = 1 - d_t$. We plot this ?? somewhere.
 
@@
 
-\begin{figure*}[tb]
-    \hskip-1.2cm\input{gen-graph-sat.tex}
-    \setlength{\abovecaptionskip}{-2em}
+\begin{figure*}[t]
+    \input{gen-graph-sat.tex}
     \caption{Behaviour of other solvers on the induced variant on smaller graphs, shown in the style of
         \cref{figure:non-induced}. The second row shows the number of search nodes used by the Glasgow
-        algorithm, the third row shows the number of decisions made by the Glucose SAT solver, and
-    the fourth row shows the number of search nodes used by a clique encoding.}\label{figure:alt}
+        algorithm, the third row shows the number of decisions made by the Clasp pseudo-boolean solver,
+        the fourth row shows the number of search nodes used by a clique encoding, and the fifth a mixed
+        integer encoding with Gurobi.}\label{figure:alt}
 \end{figure*}
@@
 
-Direct SAT encoding: variables for each assignment, cardinality constraints, edge and non-edge
-constraints. We use the Glucose SAT solver. We see that hard instances remain hard, including inside
-the central region, and that the easy satisfiable instances remain easy.
+Direct pseudo-boolean encoding: variables for each assignment, cardinality constraints, edge and
+non-edge constraints. We use the Clasp solver. We see that hard instances remain hard, including
+inside the central region, and that the easy satisfiable instances remain easy. (Although not
+presented here, we also tried an equivalent SAT encoding with the Glucose solver, and saw very
+similar performance.)
+
+\begin{figure}[tb]
+    \centering
+    \begin{tikzpicture}[scale=0.5]
+
+        \foreach \n in {1, ..., 6}{
+            \node [draw, circle, fill=uofgcobalt, inner sep=1.5pt] (N1\n) at (\n, 0) {\n};
+        }
+
+        \draw [decorate, decoration={brace, raise=0.2cm}, very thick] (N11.north west) -- (N16.north
+            east) node [midway, above=0.3cm] { $1 \mapsto$ };
+
+        \foreach \n in {1, ..., 6}{
+            \node [draw, circle, fill=uofgpumpkin, inner sep=1.5pt] (N2\n) at (0, -\n) {\n};
+        }
+
+        \draw [decorate, decoration={brace, raise=0.2cm, mirror}, very thick] (N21.north west) --
+            (N26.south west) node [midway, left=0.3cm] { $2 \mapsto$ };
+
+        \foreach \n in {1, ..., 6}{
+            \node [draw, circle, fill=uofglawn, inner sep=1.5pt] (N3\n) at (7, -\n) {\n};
+        }
+
+        \draw [decorate, decoration={brace, raise=0.2cm}, very thick] (N31.north east) -- (N36.south
+            east) node [midway, right=0.3cm] { $3 \mapsto$ };
+
+        \foreach \n in {1, ..., 6}{
+            \node [draw, circle, fill=uofgthistle, inner sep=1.5pt] (N4\n) at (\n, -7) {\n};
+        }
+
+        \draw [decorate, decoration={brace, raise=0.2cm, mirror}, very thick] (N41.south west) --
+            (N46.south east) node [midway, below=0.3cm] { $4 \mapsto$ };
+
+        \begin{scope}[on background layer]
+            \draw [color=uofgsandstone!50] ($(N11)!0.8!(N11.south)$) -- ($(N22)!0.8!(N22.east)$);
+            \draw [color=uofgsandstone!50] ($(N11)!0.8!(N11.south)$) -- ($(N23)!0.8!(N23.east)$);
+            \draw [color=uofgsandstone!50] ($(N11)!0.8!(N11.south)$) -- ($(N24)!0.8!(N24.east)$);
+            \draw [color=uofgsandstone!50] ($(N11)!0.8!(N11.south)$) -- ($(N25)!0.8!(N25.east)$);
+            \draw [color=uofgsandstone!50] ($(N11)!0.8!(N11.south)$) -- ($(N36)!0.8!(N36.west)$);
+            \draw [color=uofgsandstone!50] ($(N11)!0.8!(N11.south)$) -- ($(N46)!0.8!(N46.north)$);
+            \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N12)!0.8!(N12.south)$);
+            \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N32)!0.8!(N32.west)$);
+            \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N42)!0.8!(N42.north)$);
+            \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N13)!0.8!(N13.south)$);
+            \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N33)!0.8!(N33.west)$);
+            \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N43)!0.8!(N43.north)$);
+            \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N14)!0.8!(N14.south)$);
+            \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N34)!0.8!(N34.west)$);
+            \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N44)!0.8!(N44.north)$);
+            \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N15)!0.8!(N15.south)$);
+            \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N35)!0.8!(N35.west)$);
+            \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N45)!0.8!(N45.north)$);
+            \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N22)!0.8!(N22.east)$);
+            \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N42)!0.8!(N42.north)$);
+            \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N23)!0.8!(N23.east)$);
+            \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N43)!0.8!(N43.north)$);
+            \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N24)!0.8!(N24.east)$);
+            \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N44)!0.8!(N44.north)$);
+            \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N25)!0.8!(N25.east)$);
+            \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N45)!0.8!(N45.north)$);
+            \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N16)!0.8!(N16.south)$);
+            \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N22)!0.8!(N22.east)$);
+            \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N32)!0.8!(N32.west)$);
+            \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N23)!0.8!(N23.east)$);
+            \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N33)!0.8!(N33.west)$);
+            \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N24)!0.8!(N24.east)$);
+            \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N34)!0.8!(N34.west)$);
+            \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N25)!0.8!(N25.east)$);
+            \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N35)!0.8!(N35.west)$);
+            \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N16)!0.8!(N16.south)$);
+            \draw [color=uofgsandstone!50] ($(N12)!0.8!(N12.south)$) -- ($(N33)!0.8!(N33.west)$);
+            \draw [color=uofgsandstone!50] ($(N12)!0.8!(N12.south)$) -- ($(N24)!0.8!(N24.east)$);
+            \draw [color=uofgsandstone!50] ($(N12)!0.8!(N12.south)$) -- ($(N45)!0.8!(N45.north)$);
+            \draw [color=uofgsandstone!50] ($(N22)!0.8!(N22.east)$) -- ($(N14)!0.8!(N14.south)$);
+            \draw [color=uofgsandstone!50] ($(N22)!0.8!(N22.east)$) -- ($(N34)!0.8!(N34.west)$);
+            \draw [color=uofgsandstone!50] ($(N22)!0.8!(N22.east)$) -- ($(N44)!0.8!(N44.north)$);
+            \draw [color=uofgsandstone!50] ($(N22)!0.8!(N22.east)$) -- ($(N16)!0.8!(N16.south)$);
+            \draw [color=uofgsandstone!50] ($(N22)!0.8!(N22.east)$) -- ($(N36)!0.8!(N36.west)$);
+            \draw [color=uofgsandstone!50] ($(N22)!0.8!(N22.east)$) -- ($(N46)!0.8!(N46.north)$);
+            \draw [color=uofgsandstone!50] ($(N32)!0.8!(N32.west)$) -- ($(N13)!0.8!(N13.south)$);
+            \draw [color=uofgsandstone!50] ($(N32)!0.8!(N32.west)$) -- ($(N24)!0.8!(N24.east)$);
+            \draw [color=uofgsandstone!50] ($(N32)!0.8!(N32.west)$) -- ($(N44)!0.8!(N44.north)$);
+            \draw [color=uofgsandstone!50] ($(N32)!0.8!(N32.west)$) -- ($(N15)!0.8!(N15.south)$);
+            \draw [color=uofgsandstone!50] ($(N32)!0.8!(N32.west)$) -- ($(N26)!0.8!(N26.east)$);
+            \draw [color=uofgsandstone!50] ($(N32)!0.8!(N32.west)$) -- ($(N46)!0.8!(N46.north)$);
+            \draw [color=uofgsandstone!50] ($(N42)!0.8!(N42.north)$) -- ($(N13)!0.8!(N13.south)$);
+            \draw [color=uofgsandstone!50] ($(N42)!0.8!(N42.north)$) -- ($(N24)!0.8!(N24.east)$);
+            \draw [color=uofgsandstone!50] ($(N42)!0.8!(N42.north)$) -- ($(N34)!0.8!(N34.west)$);
+            \draw [color=uofgsandstone!50] ($(N42)!0.8!(N42.north)$) -- ($(N15)!0.8!(N15.south)$);
+            \draw [color=uofgsandstone!50] ($(N42)!0.8!(N42.north)$) -- ($(N26)!0.8!(N26.east)$);
+            \draw [color=uofgsandstone!50] ($(N42)!0.8!(N42.north)$) -- ($(N36)!0.8!(N36.west)$);
+            \draw [color=uofgsandstone!50] ($(N13)!0.8!(N13.south)$) -- ($(N34)!0.8!(N34.west)$);
+            \draw [color=uofgsandstone!50] ($(N13)!0.8!(N13.south)$) -- ($(N44)!0.8!(N44.north)$);
+            \draw [color=uofgsandstone!50] ($(N13)!0.8!(N13.south)$) -- ($(N25)!0.8!(N25.east)$);
+            \draw [color=uofgsandstone!50] ($(N13)!0.8!(N13.south)$) -- ($(N26)!0.8!(N26.east)$);
+            \draw [color=uofgsandstone!50] ($(N23)!0.8!(N23.east)$) -- ($(N15)!0.8!(N15.south)$);
+            \draw [color=uofgsandstone!50] ($(N23)!0.8!(N23.east)$) -- ($(N35)!0.8!(N35.west)$);
+            \draw [color=uofgsandstone!50] ($(N23)!0.8!(N23.east)$) -- ($(N45)!0.8!(N45.north)$);
+            \draw [color=uofgsandstone!50] ($(N23)!0.8!(N23.east)$) -- ($(N16)!0.8!(N16.south)$);
+            \draw [color=uofgsandstone!50] ($(N23)!0.8!(N23.east)$) -- ($(N36)!0.8!(N36.west)$);
+            \draw [color=uofgsandstone!50] ($(N23)!0.8!(N23.east)$) -- ($(N46)!0.8!(N46.north)$);
+            \draw [color=uofgsandstone!50] ($(N33)!0.8!(N33.west)$) -- ($(N14)!0.8!(N14.south)$);
+            \draw [color=uofgsandstone!50] ($(N33)!0.8!(N33.west)$) -- ($(N25)!0.8!(N25.east)$);
+            \draw [color=uofgsandstone!50] ($(N33)!0.8!(N33.west)$) -- ($(N45)!0.8!(N45.north)$);
+            \draw [color=uofgsandstone!50] ($(N33)!0.8!(N33.west)$) -- ($(N26)!0.8!(N26.east)$);
+            \draw [color=uofgsandstone!50] ($(N33)!0.8!(N33.west)$) -- ($(N46)!0.8!(N46.north)$);
+            \draw [color=uofgsandstone!50] ($(N43)!0.8!(N43.north)$) -- ($(N14)!0.8!(N14.south)$);
+            \draw [color=uofgsandstone!50] ($(N43)!0.8!(N43.north)$) -- ($(N25)!0.8!(N25.east)$);
+            \draw [color=uofgsandstone!50] ($(N43)!0.8!(N43.north)$) -- ($(N36)!0.8!(N36.west)$);
+            \draw [color=uofgsandstone!50] ($(N14)!0.8!(N14.south)$) -- ($(N35)!0.8!(N35.west)$);
+            \draw [color=uofgsandstone!50] ($(N14)!0.8!(N14.south)$) -- ($(N45)!0.8!(N45.north)$);
+            \draw [color=uofgsandstone!50] ($(N14)!0.8!(N14.south)$) -- ($(N26)!0.8!(N26.east)$);
+            \draw [color=uofgsandstone!50] ($(N24)!0.8!(N24.east)$) -- ($(N16)!0.8!(N16.south)$);
+            \draw [color=uofgsandstone!50] ($(N24)!0.8!(N24.east)$) -- ($(N36)!0.8!(N36.west)$);
+            \draw [color=uofgsandstone!50] ($(N24)!0.8!(N24.east)$) -- ($(N46)!0.8!(N46.north)$);
+            \draw [color=uofgsandstone!50] ($(N34)!0.8!(N34.west)$) -- ($(N15)!0.8!(N15.south)$);
+            \draw [color=uofgsandstone!50] ($(N34)!0.8!(N34.west)$) -- ($(N26)!0.8!(N26.east)$);
+            \draw [color=uofgsandstone!50] ($(N34)!0.8!(N34.west)$) -- ($(N46)!0.8!(N46.north)$);
+            \draw [color=uofgsandstone!50] ($(N44)!0.8!(N44.north)$) -- ($(N15)!0.8!(N15.south)$);
+            \draw [color=uofgsandstone!50] ($(N44)!0.8!(N44.north)$) -- ($(N26)!0.8!(N26.east)$);
+            \draw [color=uofgsandstone!50] ($(N44)!0.8!(N44.north)$) -- ($(N36)!0.8!(N36.west)$);
+            \draw [color=uofgsandstone!50] ($(N15)!0.8!(N15.south)$) -- ($(N26)!0.8!(N26.east)$);
+            \draw [color=uofgsandstone!50] ($(N25)!0.8!(N25.east)$) -- ($(N16)!0.8!(N16.south)$);
+            \draw [color=uofgsandstone!50] ($(N25)!0.8!(N25.east)$) -- ($(N36)!0.8!(N36.west)$);
+            \draw [color=uofgsandstone!50] ($(N25)!0.8!(N25.east)$) -- ($(N46)!0.8!(N46.north)$);
+            \draw [color=uofgsandstone!50] ($(N35)!0.8!(N35.west)$) -- ($(N46)!0.8!(N46.north)$);
+            \draw [color=uofgsandstone!50] ($(N45)!0.8!(N45.north)$) -- ($(N26)!0.8!(N26.east)$);
+            \draw [color=uofgsandstone!50] ($(N45)!0.8!(N45.north)$) -- ($(N36)!0.8!(N36.west)$);
+            \draw [thick] ($(N12)!0.8!(N12.south)$) -- ($(N43)!0.8!(N43.north)$);
+            \draw [thick] ($(N12)!0.8!(N12.south)$) -- ($(N35)!0.8!(N35.west)$);
+            \draw [thick] ($(N12)!0.8!(N12.south)$) -- ($(N26)!0.8!(N26.east)$);
+            \draw [thick] ($(N43)!0.8!(N43.north)$) -- ($(N35)!0.8!(N35.west)$);
+            \draw [thick] ($(N43)!0.8!(N43.north)$) -- ($(N26)!0.8!(N26.east)$);
+            \draw [thick] ($(N35)!0.8!(N35.west)$) -- ($(N26)!0.8!(N26.east)$);
+        \end{scope}
+
+    \end{tikzpicture}
+    \caption{The association graph encoding of the induced version of \cref{figure:sip}. The clique
+    of size four corresponds to the mapping sending pattern vertex 1 to target vertex 2, 2 to 6, 3
+    to 5, and 4 to 3.}\label{figure:assoc}
+\end{figure}
 
@@
 
-?? Pseudo-boolean, incomplete solvers?
+MIP encoding with Gurobi.
 
-?? \cite{Anton:2009}
+?? \citep{Anton:2009}
 
-?? \cite{Lipets:2009}
+?? \citep{Lipets:2009}
 

### 2015-11-14 paper/paper.tex
More

@@
 \usepackage{natbib}
+\usepackage{dblfloatfix}
 
 % \usepackage{showframe}
-\usepackage{lipsum}
 
@@
 \renewcommand{\bottomfraction}{0.8}
+\renewcommand{\dbltopfraction}{0.9}
+\renewcommand{\textfraction}{0.05}
+\setcounter{dbltopnumber}{2}
+\renewcommand{\floatpagefraction}{0.85}
+\renewcommand{\dblfloatpagefraction}{0.85}
 
@@
 
-However, generating non-trivial unsatisfiable random instances is not straightforward. Using a
-pattern graph from one of the random suites with the ``wrong'' target graph tends to give either a
-trivially unsatisfiable instance, or a satisfiable instance. (In particular, it is \emph{not} the
-case that a relatively small random graph is unlikely to appear in a larger random graph.)
-
 \begin{figure}[b]
@@
     \begin{tikzpicture}[scale=0.5]
+        \begin{scope}
             \node [draw, circle, fill=uofgcobalt, inner sep=1.5pt] (Na) at (1,  0) {1};
@@
             \node [anchor=center, font=\Large] (A2) at ( 3.5, -1.5) { $\rightarrowtail$ };
-    \end{tikzpicture}
-
-    \caption{There is a non-induced isomorphism from the pattern graph, in the center, to the target
-    on the right, mapping vertex 1 to 1, 2 to 3, 3 to 5 and 4 to 6. This is not an induced
-    isomorphism, since there is an edge between 1 and 5 in the target but not between 1 and 3 in the
-    pattern. The mapping from the pattern to the (same) target on the left, sending 1 to 2, 2 to 6, 3 to
-    5 and 4 to 3, is both non-induced and induced.}
-    \label{figure:sip}
-\end{figure}
-
-Here we present and evaluate a new method for creating random pattern/target pairs. This method
-generates both satisfiable and unsatisfiable instances, and can produce computationally challenging
-instances with only a few tens of vertices in the pattern, and 150 vertices in the target. Our work
-builds upon the phase transition phenomena observed in satisfiability and graph colouring problems
-first described by \citet{Cheeseman:1991} and \citet{Mitchell:1992}---?? cite later stuff and
-explain really hard.
-
-For subgraph isomorphism we have three relevant control parameters: we can independently alter the
-edge probability of the pattern graph, the edge probability of the target graph, and the relative
-orders (number of vertices) of the pattern and target graphs.  For non-induced isomorphisms, with
-the correct choice of parameters we see results very similar to those observed with boolean
-satisfiability problems: there is a phase transition from satisfiable to unsatisfiable (and we can
-predict its location), and we see a complexity peak occur (with three different solvers) near this
-phase transition.
-
-For certain choices of parameters for induced isomorphisms, there are two phase transitions, going
-from satisfiable to unsatisfiable, and then from unsatisfiable back to satisfiable. Again, when
-going from satisfiable to unsatisfiable (from either direction), instances go from being trivial to
-really hard to solve. However, each of the three solvers we tested also finds the central
-unsatisfiable region to be really hard---this may be unexpected, since this region is not near a
-phase transition and is apparently over-constrained. To show that this is not a simple weakness of
-current subgraph isomorphism algorithms, we verify that this region is also hard when using use a
-pseudo-boolean encoding, a mixed integer model, and under reduction to the clique problem.
-
-\subsection{Definitions}
-
-Throughout, our graphs are unlabelled, undirected, and do not have any loops (vertices which are
-adjacent to themselves). We write $\operatorname{V}(G)$ for the vertex set of a graph $G$. A
-\emph{non-induced subgraph isomorphism} from a graph $P$ (called the \emph{pattern}) to a graph $T$
-(called the \emph{target}) is an injective mapping from $\operatorname{V}(P)$ to
-$\operatorname{V}(T)$ which preserves adjacency---that is, for every adjacent $v$ and $w$ in
-$\operatorname{V}(P)$, the vertices $i(v)$ and $i(w)$ are adjacent in $T$. An \emph{induced subgraph
-isomorphism} additionally preserves non-adjacency---that is, if $v$ and $w$ are not adjacent in $P$,
-then $i(v)$ and $i(w)$ must not be adjacent in $T$. We use the notation $i : P \rightarrowtail T$
-for a non-induced isomorphism, and $i : P \hookrightarrow T$ for an induced isomorphism.
-
-The \emph{order} of a graph is the cardinality of its vertex set.  We write $G(v, p)$ for a random
-graph with $v$ vertices, and an edge between each distinct pair of vertices with independent
-probability $p$.
-
-\subsection{Experimental Setup}
-
-Our experiments were performed on systems with Intel Xeon E5-4650 v2 (Q1'14) CPUs and 768GBytes RAM
-(although this much RAM was not needed), running Scientific Linux release 6.7.   ?? Software
-versions. Clasp 3.1.3, Gurobi 6.0.5, Glasgow, LAD version 2, VFLib. Software was compiled using GCC
-4.9.
-
-?? Stuff about runtimes and recursive calls. We are not aiming to compare solvers; rather, we are
-looking for solver-independent behaviour.
-
-\section{Non-Induced Subgraph Isomorphisms}
-
-\begin{figure}[t]
-    \input{gen-graph-phase-transition.tex}
-    \caption{With a fixed a pattern graph order of 20, a target graph order of 150, a target edge
-        probability of 0.4, and varying pattern edge probability, we observe a phase transition and
-        complexity peak with the Glasgow solver in the non-induced variant. Each point represents
-        one instance. The lines show mean search effort, and mean proportion satisfiable, using a
-    larger sample size.}
-    \label{figure:phase-transition}
-\end{figure}
-
-Suppose we arbitrarily decide upon a pattern graph order of 20, a target graph order of 150, and a
-fixed target edge probability of 0.40. As we vary the pattern edge probability from 0 to 1, we would
-expect to see a shift from entirely satisfiable instances (with no edges in the pattern, we can
-always find a match) to entirely unsatisfiable instances (a maximum clique in this order and edge
-probability of target graph will usually have between 9 and 12 vertices). Indeed,
-\cref{figure:phase-transition} shows that this is the case: the line (and the points show a subset
-of the samples) ?? describe this a bit more, find the 0.5 SAT mark.
-
-Some stuff about the complexity peak in the Glasgow solver. This looks remarkably similar to random
-3SAT problems---compare, for example, Figure 1 of \citet{LeytonBrown:2014}. In particular,
-satisfiable instances tend to be easier but show greater variation than unsatisfiable instances, and
-there are exceptionally hard satisfiable instances. (The parallel version of the Glasgow solver is
-designed to eliminate these.)
-
-\begin{figure}[tb]
-    \input{gen-graph-non-induced.tex}
-    \caption{Behaviour of algorithms on the non-induced variant. Each point is the average of ten
-        runs. For each plot, the x-axis is the pattern edge probability and the y-axis is the target
-        edge probability, both from 0 to 1. Along the top row, we show the proportion of instances which are
-        satisfiable; the white bands shows the phase transitions. The second row is our prediction
-        of what the first row will look like. On the third row, we show the number of search nodes
-        used by the Glasgow algorithm, on the fourth row, the number of search nodes used by the LAD
-        algorithm, and the fifth, VF2; the dark regions indicate ``really hard'' instances.}
-    \label{figure:non-induced}
-\end{figure}
-
-In the top row of \cref{figure:non-induced} we show the phase transition for the non-induced
-variant, for patterns of order 10, 20 and 30, a target of order 150, and varying pattern (x-axis)
-and target (y-axis) densities. Inside the orange region, at the bottom right of each plot, every
-instance is unsatisfiable---here we are trying to find a dense pattern in a sparse target. In the
-purple region, at the top left, every instance is satisfiable---we are looking for a sparse pattern
-in a dense target (which is easy, since we only have to preserve adjacency, not non-adjacency). The
-white band between the regions shows the location of the phase transition: here, roughly half the
-instances are satisfiable.
-
-On subsequent rows, we show the average difficulty of different algorithms.  We measure search
-nodes. Cannot be used to compare performance of algorithms directly, but this does show that each
-solver found the same set of instances difficult.
-
-Satisfiable is easy, until very close to the phase transition. Unsatisfiable is much harder than
-satisfiable. This is solver-independent, although VF2 is terrible.
-
-?? Are unsatisfiable instances on the phase transition much harder than satisfiable ones?
-
-In figure \cref{figure:phase-transition-bands} we show what happens for larger patterns.
-
-\begin{figure}[tb]
-    \input{gen-graph-phase-transition-bands.tex}
-    \caption{The location of the phase transition, as the pattern grows. Shown are points where
-    between 20\% and 80\% of instances are satisfiable.}
-    \label{figure:phase-transition-bands}
-\end{figure}
-
-\subsection{Locating the Phase Transition}
-
-To predict the location of the phase transition, we can calculate the expected number of solutions
-by making a few dubious assumptions involving rounding and independence. Since we are trying to find
-an \emph{injective} mapping from a pattern $P = G(p, d_p)$ to a target $T = G(t, d_t)$, there are
-$t^{\underline{p}} = t \cdot (t - 1)\cdots(t - p + 1)$ possible assignments of target vertices to
-pattern vertices.  We expect the pattern to have $d_p \cdot \binom{p}{2}$ edges, so we obtain the
-probability of all of these edges being mapped to edges in the target by raising $d_t$ to this
-power, giving an expected number of solutions of \[ t^{\underline{p}} \cdot {d_t}^{d_p \cdot
-\binom{p}{2}} \textnormal{,} \] which we plot in ?? somewhere.
-
-\section{Induced Isomorphisms}
-
-In \cref{figure:induced} we show the induced behaviour (modified Glasgow to include the complement
-as a supplemental graph).
-
-Pattern order 10, what we expect. Patterns 20 and 30 have large unsatisfiable regions in the middle,
-but these remain hard. Pattern orders 14, 15 and 16 show the transition between the two styles.
-
-Odd unsat with LAD at the top right (not EHPs), trivial for Glasgow: mapping a non-clique into a
-clique. Glasgow looks at non-edges and can see immediately that no mapping is possible.
-
-Note that this is solver-independent.
-
-\begin{figure*}[h]
-    \input{gen-graph-induced.tex}
-    \caption{Behaviour of algorithms on the induced variant, shown in the style
-    of \cref{figure:non-induced}. The second row shows a bound on the satisfiable region, by
-    considering where a \emph{non-}induced isomorphism may also be a non-induced isomorphism between
-    complement graphs.}\label{figure:induced}
-\end{figure*}
-
-\subsection{Predicting Induced Behaviour}
-
-By repeating the argument for locating the non-induced phase transition and additionally considering
-non-edges, we get an expected number of solutions of \[t^{\underline{p}} \cdot {d_t}^{d_p \cdot
-\binom{p}{2}} \cdot {d_{\overline{t}}}^{d_{\overline{p}} \cdot \binom{p}{2}} \textnormal{,} \]
-where $d_{\overline{p}} = 1 - d_p$ and $d_{\overline{t}} = 1 - d_t$. We plot this ?? somewhere.
-
-?? Graph showing the product of satisfiable with vertical flip, min (unsat) and max (sat) of search
-nodes with vertical flip.
-
-This suggests that we're not able to make use of adjacency and non-adjacency simultaneously.
-
-\section{Other Encodings and Solvers}
-
-\begin{figure*}[t]
-    \input{gen-graph-sat.tex}
-    \caption{Behaviour of other solvers on the induced variant on smaller graphs, shown in the style of
-        \cref{figure:non-induced}. The second row shows the number of search nodes used by the Glasgow
-        algorithm, the third row shows the number of decisions made by the Clasp pseudo-boolean solver,
-        the fourth row shows the number of search nodes used by a clique encoding, and the fifth a mixed
-        integer encoding with Gurobi.}\label{figure:alt}
-\end{figure*}
-
-Is this just a weakness of current subgraph isomorphism solvers? In \cref{figure:alt} we repeat the
-experiments on smaller pattern and target graphs, using different solving techniques. Although these
-techniques are not competitive in absolute terms, we wish to see if the same pattern of behaviour
-occurs.
-
-Direct pseudo-boolean encoding: variables for each assignment, cardinality constraints, edge and
-non-edge constraints. We use the Clasp solver. We see that hard instances remain hard, including
-inside the central region, and that the easy satisfiable instances remain easy. (Although not
-presented here, we also tried an equivalent SAT encoding with the Glucose solver, and saw very
-similar performance.)
-
-\begin{figure}[tb]
-    \centering
+        \end{scope}
+    \end{tikzpicture}\\[0.5cm]
     \begin{tikzpicture}[scale=0.5]
-
+        \begin{scope}
             \foreach \n in {1, ..., 6}{
@@
                 \draw [color=uofgsandstone!50] ($(N45)!0.8!(N45.north)$) -- ($(N36)!0.8!(N36.west)$);
-            \draw [thick] ($(N12)!0.8!(N12.south)$) -- ($(N43)!0.8!(N43.north)$);
-            \draw [thick] ($(N12)!0.8!(N12.south)$) -- ($(N35)!0.8!(N35.west)$);
-            \draw [thick] ($(N12)!0.8!(N12.south)$) -- ($(N26)!0.8!(N26.east)$);
-            \draw [thick] ($(N43)!0.8!(N43.north)$) -- ($(N35)!0.8!(N35.west)$);
-            \draw [thick] ($(N43)!0.8!(N43.north)$) -- ($(N26)!0.8!(N26.east)$);
-            \draw [thick] ($(N35)!0.8!(N35.west)$) -- ($(N26)!0.8!(N26.east)$);
+                \draw [ultra thick] ($(N12)!0.8!(N12.south)$) -- ($(N43)!0.8!(N43.north)$);
+                \draw [ultra thick] ($(N12)!0.8!(N12.south)$) -- ($(N35)!0.8!(N35.west)$);
+                \draw [ultra thick] ($(N12)!0.8!(N12.south)$) -- ($(N26)!0.8!(N26.east)$);
+                \draw [ultra thick] ($(N43)!0.8!(N43.north)$) -- ($(N35)!0.8!(N35.west)$);
+                \draw [ultra thick] ($(N43)!0.8!(N43.north)$) -- ($(N26)!0.8!(N26.east)$);
+                \draw [ultra thick] ($(N35)!0.8!(N35.west)$) -- ($(N26)!0.8!(N26.east)$);
             \end{scope}
+        \end{scope}
+    \end{tikzpicture}
 
-    \end{tikzpicture}
-    \caption{The association graph encoding of the induced version of \cref{figure:sip}. The clique
-    of size four corresponds to the mapping sending pattern vertex 1 to target vertex 2, 2 to 6, 3
-    to 5, and 4 to 3.}\label{figure:assoc}
+    \caption{On top, there is a non-induced isomorphism from the pattern graph, in the center, to the target
+    on the right, mapping vertex 1 to 1, 2 to 3, 3 to 5 and 4 to 6. This is not an induced
+    isomorphism, since there is an edge between 1 and 5 in the target but not between 1 and 3 in the
+    pattern. The mapping from the pattern to the (same) target on the left, sending 1 to 2, 2 to 6, 3 to
+    5 and 4 to 3, is both non-induced and induced.
+    Below, the association graph encoding for the induced version. The bold clique of size
+    four corresponds to the same mapping.}
+    \label{figure:sip}
 \end{figure}
+
+Generating non-trivial unsatisfiable random instances is not straightforward. Using a
+pattern graph from one of the random suites with the ``wrong'' target graph tends to give either a
+trivially unsatisfiable instance, or a satisfiable instance. (In particular, it is \emph{not} the
+case that a relatively small random graph is unlikely to appear in a larger random graph.)
+
+Here we present and evaluate a new method for creating random pattern/target pairs. This method
+generates both satisfiable and unsatisfiable instances, and can produce computationally challenging
+instances with only a few tens of vertices in the pattern, and 150 vertices in the target. Our work
+builds upon the phase transition phenomena observed in satisfiability and graph colouring problems
+first described by \citet{Cheeseman:1991} and \citet{Mitchell:1992}---?? cite later stuff and
+explain really hard.
+
+For subgraph isomorphism we have three relevant control parameters: we can independently alter the
+edge probability of the pattern graph, the edge probability of the target graph, and the relative
+orders (number of vertices) of the pattern and target graphs.  For non-induced isomorphisms, with
+the correct choice of parameters we see results very similar to those observed with boolean
+satisfiability problems: there is a phase transition from satisfiable to unsatisfiable (and we can
+predict its location), and we see a complexity peak occur (with three different solvers) near this
+phase transition.
+
+For certain choices of parameters for induced isomorphisms, there are two phase transitions, going
+from satisfiable to unsatisfiable, and then from unsatisfiable back to satisfiable. Again, when
+going from satisfiable to unsatisfiable (from either direction), instances go from being trivial to
+really hard to solve. However, each of the three solvers we tested also finds the central
+unsatisfiable region to be really hard---this may be unexpected, since this region is not near a
+phase transition and is apparently over-constrained. To show that this is not a simple weakness of
+current subgraph isomorphism algorithms, we verify that this region is also hard when using use a
+pseudo-boolean encoding, a mixed integer model, and under reduction to the clique problem.
+
+\subsection{Definitions}
+
+Throughout, our graphs are unlabelled, undirected, and do not have any loops (vertices which are
+adjacent to themselves). We write $\operatorname{V}(G)$ for the vertex set of a graph $G$. A
+\emph{non-induced subgraph isomorphism} from a graph $P$ (called the \emph{pattern}) to a graph $T$
+(called the \emph{target}) is an injective mapping from $\operatorname{V}(P)$ to
+$\operatorname{V}(T)$ which preserves adjacency---that is, for every adjacent $v$ and $w$ in
+$\operatorname{V}(P)$, the vertices $i(v)$ and $i(w)$ are adjacent in $T$. An \emph{induced subgraph
+isomorphism} additionally preserves non-adjacency---that is, if $v$ and $w$ are not adjacent in $P$,
+then $i(v)$ and $i(w)$ must not be adjacent in $T$. We use the notation $i : P \rightarrowtail T$
+for a non-induced isomorphism, and $i : P \hookrightarrow T$ for an induced isomorphism.
+
+The \emph{order} of a graph is the cardinality of its vertex set.  We write $G(v, p)$ for a random
+graph with $v$ vertices, and an edge between each distinct pair of vertices with independent
+probability $p$.
+
+The \emph{association graph encoding} of a subgraph isomorphism problem is constructed by creating a
+new graph with a vertex for each pair $(p, t)$ of vertices from the pattern and target graphs
+respectively. There is an edge between vertex $(p_1, t_1)$ and vertex $(p_2, t_2)$ if mapping $p_1$
+to $t_1$ and $p_2$ to $t_2$ simultaneously is permitted. For the non-induced problem, permitted
+means either $p_1$ is not adjacent to $p_2$, or both $p_1$ and $p_2$, and $t_1$ and $t_2$, are
+adjacent; for the induced problem, it means that $p_1$ is adjacent to $p_2$ if and only if $t_1$ is
+adjacent to $t_2$. Either way, it can be seen that a clique (a subset of vertices, each of which is
+adjacent to every other in the subset) of size equal to the order of the pattern graph exists in the
+association graph if and only if the problem is satisfiable (and the vertices selected correspond to
+the assignments of the solution).
+
+?? Cite max common subgraph.
+
+\subsection{Experimental Setup}
+
+Our experiments were performed on systems with Intel Xeon E5-4650 v2 (Q1'14) CPUs and 768GBytes RAM
+(although this much RAM was not needed), running Scientific Linux release 6.7.   ?? Software
+versions. Clasp 3.1.3, Gurobi 6.0.5, Glasgow, LAD version 2, VFLib. Software was compiled using GCC
+4.9.
+
+?? Stuff about runtimes and recursive calls. We are not aiming to compare solvers; rather, we are
+looking for solver-independent behaviour.
+
+\section{Non-Induced Subgraph Isomorphisms}
+
+\begin{figure}[tb]
+    \input{gen-graph-phase-transition.tex}
+    \caption{With a fixed a pattern graph order of 20, a target graph order of 150, a target edge
+        probability of 0.4, and varying pattern edge probability, we observe a phase transition and
+        complexity peak with the Glasgow solver in the non-induced variant. Each point represents
+        one instance. The lines show mean search effort, and mean proportion satisfiable, using a
+    larger sample size.}
+    \label{figure:phase-transition}
+\end{figure}
+
+Suppose we arbitrarily decide upon a pattern graph order of 20, a target graph order of 150, and a
+fixed target edge probability of 0.40. As we vary the pattern edge probability from 0 to 1, we would
+expect to see a shift from entirely satisfiable instances (with no edges in the pattern, we can
+always find a match) to entirely unsatisfiable instances (a maximum clique in this order and edge
+probability of target graph will usually have between 9 and 12 vertices). Indeed,
+\cref{figure:phase-transition} shows that this is the case: the line (and the points show a subset
+of the samples) ?? describe this a bit more, find the 0.5 SAT mark.
+
+Some stuff about the complexity peak in the Glasgow solver. This looks remarkably similar to random
+3SAT problems---compare, for example, Figure 1 of \citet{LeytonBrown:2014}. In particular,
+satisfiable instances tend to be easier but show greater variation than unsatisfiable instances, and
+there are exceptionally hard satisfiable instances. (The parallel version of the Glasgow solver is
+designed to eliminate these.)
+
+\begin{figure}[tb]
+    \input{gen-graph-non-induced.tex}
+    \caption{Behaviour of algorithms on the non-induced variant. Each point is the average of ten
+        runs. For each plot, the x-axis is the pattern edge probability and the y-axis is the target
+        edge probability, both from 0 to 1. Along the top row, we show the proportion of instances which are
+        satisfiable; the white bands shows the phase transitions. The second row is our prediction
+        of what the first row will look like. On the third row, we show the number of search nodes
+        used by the Glasgow algorithm, on the fourth row, the number of search nodes used by the LAD
+        algorithm, and the fifth, VF2; the dark regions indicate ``really hard'' instances.}
+    \label{figure:non-induced}
+\end{figure}
+
+In the top row of \cref{figure:non-induced} we show the phase transition for the non-induced
+variant, for patterns of order 10, 20 and 30, a target of order 150, and varying pattern (x-axis)
+and target (y-axis) densities. Inside the orange region, at the bottom right of each plot, every
+instance is unsatisfiable---here we are trying to find a dense pattern in a sparse target. In the
+purple region, at the top left, every instance is satisfiable---we are looking for a sparse pattern
+in a dense target (which is easy, since we only have to preserve adjacency, not non-adjacency). The
+white band between the regions shows the location of the phase transition: here, roughly half the
+instances are satisfiable.
+
+On subsequent rows, we show the average difficulty of different algorithms.  We measure search
+nodes. Cannot be used to compare performance of algorithms directly, but this does show that each
+solver found the same set of instances difficult.
+
+Satisfiable is easy, until very close to the phase transition. Unsatisfiable is much harder than
+satisfiable. This is solver-independent, although VF2 is terrible.
+
+?? Are unsatisfiable instances on the phase transition much harder than satisfiable ones?
+
+In figure \cref{figure:phase-transition-bands} we show what happens for larger patterns.
+
+\begin{figure}[tb]
+    \input{gen-graph-phase-transition-bands.tex}
+    \caption{The location of the phase transition, as the pattern grows. Shown are points where
+    between 20\% and 80\% of instances are satisfiable.}
+    \label{figure:phase-transition-bands}
+\end{figure}
+
+\subsection{Locating the Phase Transition}
+
+To predict the location of the phase transition, we can calculate the expected number of solutions
+by making a few dubious assumptions involving rounding and independence, and drawing inspiration
+from the association graph encoding. Since we are trying to find an \emph{injective} mapping from a
+pattern $P = G(p, d_p)$ to a target $T = G(t, d_t)$, there are $t^{\underline{p}} = t \cdot (t -
+1)\cdots(t - p + 1)$ possible assignments of target vertices to pattern vertices.  We expect the
+pattern to have $d_p \cdot \binom{p}{2}$ edges, so we obtain the probability of all of these edges
+being mapped to edges in the target by raising $d_t$ to this power, giving an expected number of
+solutions of \[ \langle Sol \rangle = t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}}
+\textnormal{,} \] which we plot in ?? somewhere.
+
+\section{Induced Isomorphisms}
+
+In \cref{figure:induced} we show the induced behaviour (modified Glasgow to include the complement
+as a supplemental graph).
+
+Pattern order 10, what we expect. Patterns 20 and 30 have large unsatisfiable regions in the middle,
+but these remain hard. Pattern orders 14, 15 and 16 show the transition between the two styles.
+
+Odd unsat with LAD at the top right (not EHPs), trivial for Glasgow: mapping a non-clique into a
+clique. Glasgow looks at non-edges and can see immediately that no mapping is possible.
+
+Note that this is solver-independent.
+
+\begin{figure*}[tb]
+    \input{gen-graph-induced.tex}
+    \caption{Behaviour of algorithms on the induced variant, shown in the style
+    of \cref{figure:non-induced}. The second row shows a bound on the satisfiable region, by
+    considering where a \emph{non-}induced isomorphism may also be a non-induced isomorphism between
+    complement graphs.}\label{figure:induced}
+\end{figure*}
+
+\subsection{Predicting Induced Behaviour}
+
+By repeating the argument for locating the non-induced phase transition and additionally considering
+non-edges, we get an expected number of solutions of \[ \langle Sol \rangle = t^{\underline{p}}
+\cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot {d_{\overline{t}}}^{d_{\overline{p}} \cdot \binom{p}{2}}
+\textnormal{,} \] where $d_{\overline{p}} = 1 - d_p$ and $d_{\overline{t}} = 1 - d_t$. We plot this
+?? somewhere.
+
+?? Graph showing the product of satisfiable with vertical flip, min (unsat) and max (sat) of search
+nodes with vertical flip.
+
+This suggests that we're not able to make use of adjacency and non-adjacency simultaneously.
+
+\section{Other Encodings and Solvers}
+
+\begin{figure*}[t]
+    \input{gen-graph-sat.tex}
+    \caption{Behaviour of other solvers on the induced variant on smaller graphs, shown in the style of
+        \cref{figure:non-induced}. The second row shows the number of search nodes used by the Glasgow
+        algorithm, the third row shows the number of decisions made by the Clasp pseudo-boolean solver,
+        the fourth row shows the number of search nodes used by a clique encoding, and the fifth a mixed
+        integer encoding with Gurobi.}\label{figure:alt}
+\end{figure*}
+
+Is this just a weakness of current subgraph isomorphism solvers? In \cref{figure:alt} we repeat the
+experiments on smaller pattern and target graphs, using different solving techniques. Although these
+techniques are not competitive in absolute terms, we wish to see if the same pattern of behaviour
+occurs.
+
+Direct pseudo-boolean encoding: variables for each assignment, cardinality constraints, edge and
+non-edge constraints. We use the Clasp solver. We see that hard instances remain hard, including
+inside the central region, and that the easy satisfiable instances remain easy. (Although not
+presented here, we also tried an equivalent SAT encoding with the Glucose solver, and saw very
+similar performance.)
 

### 2015-11-15 paper/paper.tex
Tweaks

@@
         \end{scope}
-    \end{tikzpicture}\\[0.5cm]
+    \end{tikzpicture}\\[0.75cm]
     \begin{tikzpicture}[scale=0.5]
@@
     5 and 4 to 3, is both non-induced and induced.
-    Below, the association graph encoding for the induced version. The bold clique of size
-    four corresponds to the same mapping.}
+    Below, the association graph encoding for the induced version: the highlighted clique
+    corresponds to the same mapping.}
     \label{figure:sip}
@@
 Throughout, our graphs are unlabelled, undirected, and do not have any loops (vertices which are
-adjacent to themselves). We write $\operatorname{V}(G)$ for the vertex set of a graph $G$. A
-\emph{non-induced subgraph isomorphism} from a graph $P$ (called the \emph{pattern}) to a graph $T$
-(called the \emph{target}) is an injective mapping from $\operatorname{V}(P)$ to
+adjacent to themselves).  The \emph{order} of a graph is the cardinality of its vertex set. We write
+$\operatorname{V}(G)$ for the vertex set of a graph $G$.  We write $G(v, p)$ for a random graph with
+$v$ vertices, and an edge between each distinct pair of vertices with independent probability $p$.
+
+A \emph{non-induced subgraph isomorphism} from a graph $P$ (called the \emph{pattern}) to a graph
+$T$ (called the \emph{target}) is an injective mapping from $\operatorname{V}(P)$ to
 $\operatorname{V}(T)$ which preserves adjacency---that is, for every adjacent $v$ and $w$ in
@@
 for a non-induced isomorphism, and $i : P \hookrightarrow T$ for an induced isomorphism.
-
-The \emph{order} of a graph is the cardinality of its vertex set.  We write $G(v, p)$ for a random
-graph with $v$ vertices, and an edge between each distinct pair of vertices with independent
-probability $p$.
 
@@
 adjacent to every other in the subset) of size equal to the order of the pattern graph exists in the
-association graph if and only if the problem is satisfiable (and the vertices selected correspond to
-the assignments of the solution).
-
-?? Cite max common subgraph.
+association graph if and only if the problem is satisfiable, and the vertices selected correspond to
+the assignments making the solution \citep{Levi:1973}.
 

### 2015-12-23 paper/paper.tex
IJCAI 16
More
Betterer titles

@@
 
-\usepackage{ijcai13}
+\usepackage{ijcai16}
 \usepackage{times}
@@
 
-\title{Really Hard Instances for the Subgraph Isomorphism Problem}
+\title{Heuristics and Really Hard Instances for Subgraph Isomorphism Problems}
 \author{Ciaran McCreesh\thanks{This work was supported by the Engineering and Physical Sciences
@@
     this region remains hard when using pseudo-boolean and mixed integer encodings, and under
-    reduction to the clique problem.
+    reduction to the clique problem. Finally, we show that existing heuristics for the non-induced
+    variant can be recovered by determining how to maximise the expected number of solutions, and
+    that this technique gives new, better heuristics for the induced variant.
 \end{abstract}
@@
     \caption{The location of the phase transition, as the pattern grows. Shown are points where
-    between 20\% and 80\% of instances are satisfiable.}
+    between 20\% and 80\% of instances are satisfiable. ?? Can we put predictions on here too?}
     \label{figure:phase-transition-bands}
@@
 
-\subsection{Locating the Phase Transition}
+\subsection{Locating the Phase Transition, and Heuristics}
 
@@
 
+\begin{figure*}[tb]
+    \input{gen-graph-induced.tex}
+    \caption{Behaviour of algorithms on the induced variant, shown in the style
+    of \cref{figure:non-induced}. The second row shows a bound on the satisfiable region, by
+    considering where a \emph{non-}induced isomorphism may also be a non-induced isomorphism between
+    complement graphs.}\label{figure:induced}
+\end{figure*}
+
 In \cref{figure:induced} we show the induced behaviour (modified Glasgow to include the complement
@@
 Note that this is solver-independent.
-
-\begin{figure*}[tb]
-    \input{gen-graph-induced.tex}
-    \caption{Behaviour of algorithms on the induced variant, shown in the style
-    of \cref{figure:non-induced}. The second row shows a bound on the satisfiable region, by
-    considering where a \emph{non-}induced isomorphism may also be a non-induced isomorphism between
-    complement graphs.}\label{figure:induced}
-\end{figure*}
 
@@
 This suggests that we're not able to make use of adjacency and non-adjacency simultaneously.
+
+\subsection{New Heuristics}
 

### 2015-12-24 code/random_subgraph_isomorphism.cc
supplemental-induced option

@@ -120,6 +120,7 @@ auto main(int argc, char * argv[]) -> int
             ("help",                                  "Display help information")
             ("timeout",            po::value<int>(),  "Abort after this many seconds")
             ("induced",                               "Induced version")
+            ("supplemental-induced",                  "Use supplemental graphs on the compement")
             ("clique",                                "Use clique algorithm")
             ("density",                               "Generate using density instead of probability")
             ("invert-pattern-order",                  "Use reverse order for pattern degree value ordering")
@@ -174,6 +175,7 @@ auto main(int argc, char * argv[]) -> int
         Params params;
 
         params.induced = options_vars.count("induced");
+        params.supplemental_induced = options_vars.count("supplemental-induced");
 
         params.invert_pattern_order = options_vars.count("invert-pattern-order");
         params.invert_target_order = options_vars.count("invert-target-order");

### 2015-12-24 code/sequential.cc

@@ -17,7 +17,7 @@ namespace
         Satisfiable
     };
 
-    template <unsigned n_words_, int k_, int l_, bool induced_>
+    template <unsigned n_words_, int k_, int l_, bool induced_, bool supplemental_induced_>
     struct SequentialSubgraphIsomorphism
     {
         struct Domain
@@ -65,7 +65,7 @@ namespace
 
         const Params & params;
 
-        static constexpr int max_graphs = 1 + (l_ - 1) * k_ + (induced_ ? 1 : 0);
+        static constexpr int max_graphs = 1 + (((l_ - 1) * k_) * (induced_ && supplemental_induced_ ? 2 : 1)) + (induced_ ? 1 : 0);
         std::array<FixedBitGraph<n_words_>, max_graphs> target_graphs;
         std::array<FixedBitGraph<n_words_>, max_graphs> pattern_graphs;
 
@@ -216,6 +216,106 @@ namespace
 
                 target_graphs.at(1 + ((l_ - 1) * k_)) = target_graphs.at(0);
                 target_graphs.at(1 + ((l_ - 1) * k_)).complement();
+
+                if (supplemental_induced_) {
+                    constexpr unsigned start = 1 + ((l_ - 1) * k_);
+
+                    if (l_ >= 2) {
+                        for (unsigned v = 0 ; v < pattern_size ; ++v) {
+                            auto nv = pattern_graphs.at(start + 0).neighbourhood(v);
+                            for (int c = nv.first_set_bit() ; c != -1 ; c = nv.first_set_bit()) {
+                                nv.unset(c);
+                                auto nc = pattern_graphs.at(start + 0).neighbourhood(c);
+                                for (int w = nc.first_set_bit() ; w != -1 && unsigned(w) <= v ; w = nc.first_set_bit()) {
+                                    nc.unset(w);
+                                    if (k_ >= 3 && pattern_graphs.at(start + 2).adjacent(v, w))
+                                        pattern_graphs.at(start + 3).add_edge(v, w);
+                                    else if (k_ >= 2 && pattern_graphs.at(start + 1).adjacent(v, w))
+                                        pattern_graphs.at(start + 2).add_edge(v, w);
+                                    else if (k_ >= 1)
+                                        pattern_graphs.at(start + 1).add_edge(v, w);
+                                }
+                            }
+                        }
+                    }
+
+                    if (l_ >= 3) {
+                        for (unsigned v = 0 ; v < pattern_size ; ++v) {
[…]

### 2015-12-26 code/sequential.cc
Don't invert alldiff tiebreaking

@@ -70,7 +70,7 @@ namespace
         std::array<FixedBitGraph<n_words_>, max_graphs> pattern_graphs;
 
         std::vector<int> pattern_order, target_order;
-        std::array<int, n_words_ * bits_per_word> pattern_degree_tiebreak;
+        std::array<int, n_words_ * bits_per_word> pattern_degree_tiebreak, all_diff_tiebreak;
 
         unsigned pattern_size, full_pattern_size, target_size;
 
@@ -102,8 +102,10 @@ namespace
                     if (target.adjacent(target_order.at(i), target_order.at(j)))
                         target_graphs.at(0).add_edge(i, j);
 
-            for (unsigned j = 0 ; j < pattern_size ; ++j)
+            for (unsigned j = 0 ; j < pattern_size ; ++j) {
                 pattern_degree_tiebreak.at(j) = pattern_graphs.at(0).degree(j) * (params.invert_pattern_order ? -1 : 1);
+                all_diff_tiebreak.at(j) = pattern_graphs.at(0).degree(j);
+            }
         }
 
         auto build_supplemental_graphs() -> void
@@ -530,7 +532,7 @@ namespace
             std::sort(domains_order.begin(), domains_order.begin() + domains.size(),
                     [&] (int a, int b) {
                     return (domains.at(a).popcount < domains.at(b).popcount) ||
-                    (domains.at(a).popcount == domains.at(b).popcount && pattern_degree_tiebreak.at(domains.at(a).v) > pattern_degree_tiebreak.at(domains.at(b).v));
+                    (domains.at(a).popcount == domains.at(b).popcount && all_diff_tiebreak.at(domains.at(a).v) > all_diff_tiebreak.at(domains.at(b).v));
                     });
 
             // counting all-different

### 2015-12-27 code/random_subgraph_isomorphism.cc
No backjumping option

@@ -125,6 +125,7 @@ auto main(int argc, char * argv[]) -> int
             ("density",                               "Generate using density instead of probability")
             ("invert-pattern-order",                  "Use reverse order for pattern degree value ordering")
             ("invert-target-order",                   "Use reverse order for target degree value ordering")
+            ("no-backjumping",                        "Disable backjumping")
             ;
 
         po::options_description all_options{ "All options" };
@@ -176,6 +177,7 @@ auto main(int argc, char * argv[]) -> int
 
         params.induced = options_vars.count("induced");
         params.supplemental_induced = options_vars.count("supplemental-induced");
+        params.no_backjumping = options_vars.count("no-backjumping");
 
         params.invert_pattern_order = options_vars.count("invert-pattern-order");
         params.invert_target_order = options_vars.count("invert-target-order");

### 2015-12-27 code/sequential.cc

@@ -407,8 +407,9 @@ namespace
                     case Search::Unsatisfiable:  break;
                 }
 
-                if (search_result.second.independent_of(domains, new_domains))
-                    return search_result;
+                if (! params.no_backjumping)
+                    if (search_result.second.independent_of(domains, new_domains))
+                        return search_result;
 
                 shared_failed_variables.add(search_result.second);
             }

### 2016-01-06 paper/paper.tex
More stuff
More
Graph differently

@@
 
+?? Heuristics
+
 \subsection{Definitions}
@@
 satisfiable instances tend to be easier but show greater variation than unsatisfiable instances, and
-there are exceptionally hard satisfiable instances. (The parallel version of the Glasgow solver is
-designed to eliminate these.)
+there are exceptionally hard satisfiable instances \citep{Smith:1997}.
 
@@
 
-In the top row of \cref{figure:non-induced} we show the phase transition for the non-induced
-variant, for patterns of order 10, 20 and 30, a target of order 150, and varying pattern (x-axis)
-and target (y-axis) densities. Inside the orange region, at the bottom right of each plot, every
+What if we alter the edge probabilities for both the pattern graph and the target graph?  In the top
+row of \cref{figure:non-induced} we show the phase transition for the non-induced variant, for
+patterns of order 10, 20 and 30, a target of order 150, and varying pattern (x-axis) and target
+(y-axis) edge probabilities. Inside the orange region, at the bottom right of each plot, every
 instance is unsatisfiable---here we are trying to find a dense pattern in a sparse target. In the
@@
 nodes. Cannot be used to compare performance of algorithms directly, but this does show that each
-solver found the same set of instances difficult.
-
-Satisfiable is easy, until very close to the phase transition. Unsatisfiable is much harder than
-satisfiable. This is solver-independent, although VF2 is terrible.
+solver found the same set of instances difficult.  Satisfiable is easy, until very close to the
+phase transition. Unsatisfiable is much harder than satisfiable. This is solver-independent,
+although VF2 is terrible.
 
 ?? Are unsatisfiable instances on the phase transition much harder than satisfiable ones?
-
-In figure \cref{figure:phase-transition-bands} we show what happens for larger patterns.
-
-\begin{figure}[tb]
-    \input{gen-graph-phase-transition-bands.tex}
-    \caption{The location of the phase transition, as the pattern grows. Shown are points where
-    between 20\% and 80\% of instances are satisfiable. ?? Can we put predictions on here too?}
-    \label{figure:phase-transition-bands}
-\end{figure}
 
@@
 pattern $P = G(p, d_p)$ to a target $T = G(t, d_t)$, there are $t^{\underline{p}} = t \cdot (t -
-1)\cdots(t - p + 1)$ possible assignments of target vertices to pattern vertices.  We expect the
+1) \cdot \ldots \cdot (t - p + 1)$ possible assignments of target vertices to pattern vertices.  We expect the
 pattern to have $d_p \cdot \binom{p}{2}$ edges, so we obtain the probability of all of these edges
@@
 solutions of \[ \langle Sol \rangle = t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}}
-\textnormal{,} \] which we plot in ?? somewhere.
+\textnormal{,} \] which we plot in the second row of \cref{figure:non-induced}.
+
+We can see that this prediction is generally accurate, except that for very low and very high
+pattern densities, we overestimate the satisfiable region slightly. This is because although an
+expected number of solutions much below one implies a high likelihood of unsatisfiability, it is not
+true that a high expected number of solutions implies that any particular instance is likely to be
+satisfiable---a similar behaviour is seen with random constraint satisfaction problems
+\citep{Smith:1994,Smith:1996}.
 
@@
 
+?? Note that it is always best to try to move towards the satisfiable region, even for unsatisfiable
+instances. This contradicts the suggestion by \citet{Walsh:1998} that switching heuristics based
+upon an estimate of the solubility of the problem may offer good performance.
+
 \section{Other Encodings and Solvers}
@@
 
+Dynamic heuristics
+
+Other random models and non-random
+
+Does not give hard instances for graph isomorphism
+
 \bibliographystyle{named}

### 2016-01-10 paper/paper.tex
More words
Notation, etc

@@
 
-?? Heuristics
+?? Also there's heuristics. Figure out how to work this smoothly into the intro.
 
@@
 (although this much RAM was not needed), running Scientific Linux release 6.7.   ?? Software
-versions. Clasp 3.1.3, Gurobi 6.0.5, Glasgow, LAD version 2, VFLib. Software was compiled using GCC
+versions: Clasp 3.1.3, Gurobi 6.0.5, Glasgow, LAD version 2, VFLib. Software was compiled using GCC
 4.9.
@@
 non-edges, we get an expected number of solutions of \[ \langle Sol \rangle = t^{\underline{p}}
-\cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot {d_{\overline{t}}}^{d_{\overline{p}} \cdot \binom{p}{2}}
-\textnormal{,} \] where $d_{\overline{p}} = 1 - d_p$ and $d_{\overline{t}} = 1 - d_t$. We plot this
-?? somewhere.
+    \cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot {(1 - d_{t})}^{(1 - d_{p}) \cdot \binom{p}{2}}
+\textnormal{.} \] We plot this
+on the second row of \cref{figure:induced}---again, our prediction is accurate except for very
+sparse or very dense patterns.
 

### 2016-01-11 paper/paper.tex
Muck around a bit

@@
 
-% \usepackage{showframe}
+\usepackage{showframe}
 
@@
 
-\begin{figure*}[tb]
+\begin{figure*}[p]
     \input{gen-graph-induced.tex}
@@
     \caption{Behaviour of other solvers on the induced variant on smaller graphs, shown in the style of
-        \cref{figure:non-induced}. The second row shows the number of search nodes used by the Glasgow
-        algorithm, the third row shows the number of decisions made by the Clasp pseudo-boolean solver,
-        the fourth row shows the number of search nodes used by a clique encoding, and the fifth a mixed
-        integer encoding with Gurobi.}\label{figure:alt}
+        \cref{figure:non-induced}. The second row shows the number of search nodes used by the
+    Glasgow algorithm, the third row shows the number of decisions made by the Clasp pseudo-boolean
+solver, and the fourth row shows the number of search nodes used by a clique
+encoding.}\label{figure:alt}
 \end{figure*}
@@
 
+We also tried using a mixed integer encoding along similar lines. Gurobi was only able to solve some
+of the trivial satisfiable instances, and was almost never able to prove unsatisfiability within the
+time limit. We do not plot these results.
+
 Clique encoding: association graph, a version of BBMC modified to solve the decision problem. The
 hard instances in the central region remain hard. Additionally, some of the easy satisfiable
-instances become hard. Possible weakness of clique algorithms for this kind of input.
+instances become hard. Possible weakness in variable ordering heuristics.
 
-MIP encoding with Gurobi.
+Some easy instances become hard, here, but hard instances remain hard. In particular, the large
+central region for the larger pattern graphs which is not near the phase transitions is still
+relatively difficult in all encodings. This suggests that these instances may genuinely be hard,
+despite not being near a phase transition, and that this does not indicate a weakness in subgraph
+isomorphism solvers.
 
 ?? \citep{Anton:2009}
-
-?? \citep{Lipets:2009}
 
@@
 
+?? \citep{Lipets:2009} inexact solutions?
+
 \bibliographystyle{named}

### 2016-01-11 paper/paper.tex
More definitionary

@@
 adjacent to themselves).  The \emph{order} of a graph is the cardinality of its vertex set. We write
-$\operatorname{V}(G)$ for the vertex set of a graph $G$.  We write $G(v, p)$ for a random graph with
-$v$ vertices, and an edge between each distinct pair of vertices with independent probability $p$.
+$\operatorname{V}(G)$ for the vertex set of a graph $G$. The \emph{complement} of a graph $G$,
+denoted $\overline{G}$, is the
+graph with the same vertex set as $G$, and with an edge between distinct vertices $v$ and $w$ iff
+$v$ and $w$ are not adjacent in $G$. We write $G(v, p)$ for a ?? ER random graph with $v$ vertices,
+and an edge between each distinct pair of vertices with independent probability $p$.
 
@@
 then $i(v)$ and $i(w)$ must not be adjacent in $T$. We use the notation $i : P \rightarrowtail T$
-for a non-induced isomorphism, and $i : P \hookrightarrow T$ for an induced isomorphism.
+for a non-induced isomorphism, and $i : P \hookrightarrow T$ for an induced isomorphism. Observe
+that an induced isomorphism $i : P \hookrightarrow T$ is a non-induced isomorphism $i : P
+\rightarrowtail T$ which is also a non-induced isomorphism $\overline{i} : \overline{P}
+\rightarrowtail \overline{T}$.
 

### 2016-01-11 paper/paper.tex
Move

@@
 degree) and permuting the vertices.  This is not ideal: such instances are guaranteed to be
-satisfiable, and so existing benchmark suites contain relatively few non-trivial unsatisfiable
-instances. These instances also tend to be computationally fairly easy, with most of the difficulty
-being in dealing with the size of the model.
+satisfiable (?? mention \citep{Anton:2009}), and so existing benchmark suites contain relatively few
+non-trivial unsatisfiable instances. These instances also tend to be computationally fairly easy,
+with most of the difficulty being in dealing with the size of the model.
 
@@
 
-?? \citep{Anton:2009}
-
 \section{Conclusion}

### 2016-01-12 prediction/induced-line.rb
Very clevar

@@ -1,0 +1,59 @@
+#!/usr/bin/ruby
+# vim: set sw=4 sts=4 et tw=80 :
+
+require 'bigdecimal'
+
+ps = ARGV[0].to_i
+ts = ARGV[1].to_i
+flip = ARGV[2].to_i == "true"
+
+0.step(500, 1) do | x |
+    pd = BigDecimal.new(x) / 500;
+
+    results = [0, 0]
+    expsols = [0, 0]
+
+    [0, 1].each do | pass |
+
+        if 0 == pass
+            tdupper = x + 10
+            tdlower = 0
+        else
+            tdupper = 500
+            tdlower = x + 10
+        end
+
+        0.upto(100) do
+            td = BigDecimal.new(tdupper + tdlower) / 1000;
+
+            if flip then pd, td = td, pd end
+
+            cliqueedges = BigDecimal.new(ps * (ps - 1) / 2);
+            nedges = pd * cliqueedges;
+            nnonedges = (1 - pd) * cliqueedges;
+
+            edgesokprob = td ** nedges;
+            nonedgesokprob = (1 - td) ** nnonedges;
+
+            solprob = edgesokprob * nonedgesokprob;
+
+            nstates = BigDecimal.new(1);
+            1.upto(ps) do | n |
+                nstates *= (ts - n + 1);
+            end
+
+            expsols[pass] = nstates * solprob;
+
+            if ((expsols[pass] < 1) ^ (pass == 1))
+                tdlower = (tdupper + tdlower) / 2
+            elsif ((expsols[pass] > 1) ^ (pass == 1))
+                tdupper = (tdupper + tdlower) / 2
+            else
+                break
+            end
+        end
+        results[pass] = (tdupper + tdlower) / 1000.0
+    end
+
+    puts(pd.to_f.to_s + " " + results[0].to_s + " " + expsols[0].to_f.to_s + " " + results[1].to_s + " " + expsols[1].to_f.to_s + " " + (x + 10).to_s)
+end

### 2016-01-12 prediction/noninduced-line.rb

@@ -1,0 +1,44 @@
+#!/usr/bin/ruby
+# vim: set sw=4 sts=4 et tw=80 :
+
+require 'bigdecimal'
+
+ps = ARGV[0].to_i
+ts = ARGV[1].to_i
+
+0.upto(500) do | x |
+    pd = BigDecimal.new(x) / 500;
+
+    tdupper = 500
+    tdlower = 0
+    expsols = 0
+
+    0.upto(100) do | y |
+        td = BigDecimal.new(tdupper + tdlower) / 1000;
+
+        cliqueedges = BigDecimal.new(ps * (ps - 1) / 2);
+        nedges = pd * cliqueedges;
+        nnonedges = (1 - pd) * cliqueedges;
+
+        edgesokprob = td ** nedges;
+
+        solprob = edgesokprob;
+
+        nstates = BigDecimal.new(1);
+        1.upto(ps) do | n |
+            nstates *= (ts - n + 1);
+        end
+
+        expsols = nstates * solprob;
+
+        if (expsols < 1)
+            tdlower = (tdupper + tdlower) / 2
+        elsif (expsols > 1)
+            tdupper = (tdupper + tdlower) / 2
+        else
+            break
+        end
+    end
+
+    puts(pd.to_f.to_s + " " + ((tdupper + tdlower) / 1000.0).to_s + " " + expsols.to_s)
+end

### 2016-01-12 prediction/induced-line.rb
Better predictions

@@ -5,28 +5,42 @@ require 'bigdecimal'
 
 ps = ARGV[0].to_i
 ts = ARGV[1].to_i
-flip = ARGV[2].to_i == "true"
+flip = (ARGV[2] == "true")
 
-0.step(500, 1) do | x |
-    pd = BigDecimal.new(x) / 500;
-
-    results = [0, 0]
+0.step(500, 10) do | x |
+    pdresult = [0, 0]
+    tdresult = [0, 0]
     expsols = [0, 0]
+    pd, td = 0, 0
 
     [0, 1].each do | pass |
 
-        if 0 == pass
-            tdupper = x + 10
-            tdlower = 0
+        if flip then
+            if 0 == pass
+                lower = 250
+                upper = 500
+            else
+                lower = 0
+                upper = 250
+            end
         else
-            tdupper = 500
-            tdlower = x + 10
+            if 0 == pass then
+                upper = x + 10
+                lower = 0
+            else
+                upper = 500
+                lower = x + 10
+            end
         end
 
         0.upto(100) do
-            td = BigDecimal.new(tdupper + tdlower) / 1000;
-
-            if flip then pd, td = td, pd end
+            if flip then
+                td = BigDecimal.new(x) / 500;
+                pd = BigDecimal.new(upper + lower) / 1000;
+            else
+                pd = BigDecimal.new(x) / 500;
+                td = BigDecimal.new(upper + lower) / 1000;
+            end
 
             cliqueedges = BigDecimal.new(ps * (ps - 1) / 2);
             nedges = pd * cliqueedges;
@@ -45,15 +59,23 @@ flip = ARGV[2].to_i == "true"
             expsols[pass] = nstates * solprob;
 
             if ((expsols[pass] < 1) ^ (pass == 1))
-                tdlower = (tdupper + tdlower) / 2
+                lower = (upper + lower) / 2
             elsif ((expsols[pass] > 1) ^ (pass == 1))
-                tdupper = (tdupper + tdlower) / 2
+                upper = (upper + lower) / 2
             else
                 break
             end
         end
-        results[pass] = (tdupper + tdlower) / 1000.0
+
+        if flip then
+            pdresult[pass] = (upper + lower) / 1000.0
+            tdresult[pass] = td
+        else
+            pdresult[pass] = pd
[…]

### 2016-01-13 paper/paper.tex
More writing and drawing

@@
 
-\usepackage{showframe}
+% \usepackage{showframe}
 
@@
 handle problem instances with many hundreds of vertices in the pattern graph, and up to ten thousand
-vertices in the target graph \citep{Cordella:2004,Solnon:2010,Audemard:2014,McCreesh:2015}.
-
-However, these algorithms cannot handle \emph{arbitrary} instances of this size. The experimental
-evaluations of these algorithms were performed using a mix of real-world graphs, graphs that encode
-biochemistry and computer vision problems, and randomly generated graph pairs. Using random
-instances to evaluate algorithm behaviour can be beneficial, because it provides a way of generating
-many instances cheaply, and reduces the risk of over-fitting when tuning design parameters. The
-random instances used in each case came from common datasets (cite ??), which were generated by
-taking a random subgraph of a random graph (using various models, including scale-free and bounded
-degree) and permuting the vertices.  This is not ideal: such instances are guaranteed to be
-satisfiable (?? mention \citep{Anton:2009}), and so existing benchmark suites contain relatively few
-non-trivial unsatisfiable instances. These instances also tend to be computationally fairly easy,
-with most of the difficulty being in dealing with the size of the model.
+vertices in the target graph \citep{Cordella:2004,Solnon:2010,Audemard:2014,McCreesh:2015}, leading
+to successful application in areas such as computer
+vision \citep{Damiand:2011,Solnon:2015}, biochemistry \citep{Giugno:2013}, and pattern recognition
+\citep{Conte:2004}.
 
@@
 
+However, these algorithms cannot handle \emph{arbitrary} instances of this size. The experimental
+evaluations of these algorithms were performed using a mix of real-world graphs, graphs that encode
+biochemistry and computer vision problems, and randomly generated graph pairs. Using random
+instances to evaluate algorithm behaviour can be beneficial, because it provides a way of generating
+many instances cheaply, and reduces the risk of over-fitting when tuning design parameters. The
+random instances used in each case came from common datasets \citep{DeSanto:2003,Zampelli:2010},
+which were generated by taking a random subgraph of a random graph (using various models, including
+Erd\H{o}s-R\'enyi, scale-free, bounded degree, and meshes) and permuting the vertices.  This is not
+ideal: such instances are guaranteed to be satisfiable (?? mention \citep{Anton:2009}), and so
+existing benchmark suites contain relatively few non-trivial unsatisfiable instances. These
+instances also tend to be computationally fairly easy, with most of the difficulty being in dealing
+with the size of the model.
+
 Generating non-trivial unsatisfiable random instances is not straightforward. Using a
@@
 explain really hard.
+
+?? Maybe introduce \cref{figure:phase-transition} here instead of at the start of the next section?
 
@@
 graph with the same vertex set as $G$, and with an edge between distinct vertices $v$ and $w$ iff
-$v$ and $w$ are not adjacent in $G$. We write $G(v, p)$ for a ?? ER random graph with $v$ vertices,
-and an edge between each distinct pair of vertices with independent probability $p$.
+$v$ and $w$ are not adjacent in $G$. We write $G(v, p)$ for an Erd\H{o}s-R\'enyi random graph with
+$v$ vertices, and an edge between each distinct pair of vertices with independent probability $p$.
 
@@
 association graph if and only if the problem is satisfiable, and the vertices selected correspond to
-the assignments making the solution \citep{Levi:1973}.
+the assignments making the solution \citep{Levi:1973}. We illustrate this in
+\cref{figure:sip}. ?? Can we move this to section 4? I thought we'd need this for the sol formulae,
+but it appears not.
 
@@
 looking for solver-independent behaviour.
+
+?? Timeouts
 
@@
         edge probability, both from 0 to 1. Along the top row, we show the proportion of instances which are
-        satisfiable; the white bands shows the phase transitions. The second row is our prediction
-        of what the first row will look like. On the third row, we show the number of search nodes
-        used by the Glasgow algorithm, on the fourth row, the number of search nodes used by the LAD
-        algorithm, and the fifth, VF2; the dark regions indicate ``really hard'' instances.}
+        satisfiable; the white bands shows the phase transitions, and the black lines are our
+        predictions of where the phase transition will occur. On the second row, we show the number
+    of search nodes used by the Glasgow algorithm, on the third row, the number of search nodes used
+by the LAD algorithm, and the fourth, VF2; the dark regions indicate ``really hard'' instances.}
     \label{figure:non-induced}
@@
 white band between the regions shows the location of the phase transition: here, roughly half the
-instances are satisfiable.
+instances are satisfiable. (We discuss the black line below.)
 
@@
 
-\subsection{Locating the Phase Transition, and Heuristics}
+\subsection{Locating the Phase Transition}
 
@@
 solutions of \[ \langle Sol \rangle = t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}}
-\textnormal{,} \] which we plot in the second row of \cref{figure:non-induced}.
+\textnormal{.} \] This formula predicts a very sharp phase transition from $\langle Sol \rangle \ll
+1$ to $\langle Sol \rangle \gg 1$. We plot where this occurs using black lines in the first row of
+\cref{figure:non-induced}.
 
@@
 
+?? Also because of half edges.
+
+\subsection{Variable and Value Ordering Heuristics}
+
+?? Note that the Glasgow heuristics are moving us towards the satisfiable region, or increasing the
+expected number of solutions \citep{Gent:1996:EN}.
+
+?? LAD does SDF with no tiebreaking, no value ordering heuristic.
+
+?? In contrast, VF2 heuristics work by selecting vertices adjacent to vertices already selected.
+
+?? I suppose we should mention kappa \citep{Gent:1996:Kappa}, etc, here. This whole thing is a bit
+subtle: maximising E(N) or Kappa as a principle is encoding-independent, but whether this is done by
+SDF, deg, etc, is encoding-dependent.
+
 \section{Induced Isomorphisms}
 
-\begin{figure*}[p]
+\begin{figure*}[tb]
     \input{gen-graph-induced.tex}
     \caption{Behaviour of algorithms on the induced variant, shown in the style
-    of \cref{figure:non-induced}. The second row shows a bound on the satisfiable region, by
+    of \cref{figure:non-induced}. The second, third and fourth rows show the number of search nodes used by the
+    Glasgow, LAD and VF2 algorithms. The fifth row shows a bound on the satisfiable region, by
     considering where a \emph{non-}induced isomorphism may also be a non-induced isomorphism between
-    complement graphs.}\label{figure:induced}
+    complement graphs. The final row shows when the Glasgow algorithm performs better when given the
+    complements of the pattern and target graphs as inputs.}\label{figure:induced}
 \end{figure*}
@@
 
+?? Explain the flip bound on row 5, and note that the difficult region is roughly near the phase
+transitions in the original graphs.
+
+?? This suggests that we're not able to make use of adjacency and non-adjacency simultaneously. Note
+that this is solver-independent. In the following section we see whether this is an oddity of
+subgraph isomorphism algorithms.
+
 Odd unsat with LAD at the top right (not EHPs), trivial for Glasgow: mapping a non-clique into a
 clique. Glasgow looks at non-edges and can see immediately that no mapping is possible.
-
-Note that this is solver-independent.
 
@@
 \textnormal{.} \] We plot this
-on the second row of \cref{figure:induced}---again, our prediction is accurate except for very
+using black lines on \cref{figure:induced}---again, our prediction is accurate except for very
 sparse or very dense patterns.
 
-?? Graph showing the product of satisfiable with vertical flip, min (unsat) and max (sat) of search
-nodes with vertical flip.
+\subsection{New Heuristics}
 
-This suggests that we're not able to make use of adjacency and non-adjacency simultaneously.
+?? Explain why we might guess degree heuristics just won't work for induced.
 
-\subsection{New Heuristics}
+?? Explain bottom row of \cref{figure:induced}, and give way of calculating it.
 
@@
 upon an estimate of the solubility of the problem may offer good performance.
+
+?? Is this heuristic better in general, on non-random instances?
 
@@
 
+?? The odd stuff in the middle, which isn't near a phase transition, but is hard.
+
 Is this just a weakness of current subgraph isomorphism solvers? In \cref{figure:alt} we repeat the
@@
 non-edge constraints. We use the Clasp solver. We see that hard instances remain hard, including
-inside the central region, and that the easy satisfiable instances remain easy. (Although not
-presented here, we also tried an equivalent SAT encoding with the Glucose solver, and saw very
-similar performance.)
+inside the central region, and that the easy satisfiable instances remain easy.
 
-We also tried using a mixed integer encoding along similar lines. Gurobi was only able to solve some
-of the trivial satisfiable instances, and was almost never able to prove unsatisfiability within the
-time limit. We do not plot these results.
+Although not presented here, we tried an equivalent SAT encoding with direct encoding of the
+cardinality constraints using the Glucose solver, and saw very similar performance.  We also tried
+using a mixed integer encoding along similar lines---Gurobi was only able to solve some of the
+trivial satisfiable instances, and was almost never able to prove unsatisfiability within the time
+limit. We do not plot these results.
 
-Clique encoding: association graph, a version of BBMC modified to solve the decision problem. The
-hard instances in the central region remain hard. Additionally, some of the easy satisfiable
-instances become hard. Possible weakness in variable ordering heuristics.
+Clique encoding: association graph, a version of BBMC \citep{SanSegundo:2011} modified to solve the
+decision problem. The hard instances in the central region remain hard. Additionally, some of the
+easy satisfiable instances become hard. Possible weakness in variable ordering heuristics.
 

### 2016-01-13 code/solve_subgraph_isomorphism.cc
Code for file-based experiments

@@ -1,0 +1,284 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "sequential.hh"
+#include "clique.hh"
+
+#include <boost/program_options.hpp>
+
+#include <iostream>
+#include <fstream>
+#include <exception>
+#include <cstdlib>
+#include <chrono>
+#include <thread>
+#include <mutex>
+#include <condition_variable>
+#include <algorithm>
+
+namespace po = boost::program_options;
+
+using std::chrono::steady_clock;
+using std::chrono::duration_cast;
+using std::chrono::milliseconds;
+
+namespace
+{
+    class GraphFileError :
+        public std::exception
+    {
+        private:
+            std::string _what;
+
+        public:
+            GraphFileError(const std::string & filename, const std::string & message) throw ();
+
+            auto what() const throw () -> const char *;
+    };
+
+    GraphFileError::GraphFileError(const std::string & filename, const std::string & message) throw () :
+        _what("Error reading graph file '" + filename + "': " + message)
+    {
+    }
+
+    auto GraphFileError::what() const throw () -> const char *
+    {
+        return _what.c_str();
+    }
+
+    namespace
+    {
+        auto read_word(std::ifstream & infile) -> int
+        {
+            int x;
+            infile >> x;
+            return x;
+        }
+    }
+
+    auto read_lad(const std::string & filename) -> Graph
+    {
+        Graph result(0);
+
+        std::ifstream infile{ filename };
+        if (! infile)
+            throw GraphFileError{ filename, "unable to open file" };
+
+        result.resize(read_word(infile));
+        if (! infile)
+            throw GraphFileError{ filename, "error reading size" };
+
+        for (int r = 0 ; r < result.size() ; ++r) {
+            int c_end = read_word(infile);
+            if (! infile)
+                throw GraphFileError{ filename, "error reading edges count" };
+
+            for (int c = 0 ; c < c_end ; ++c) {
+                int e = read_word(infile);
+
+                if (e < 0 || e >= result.size())
+                    throw GraphFileError{ filename, "edge index out of bounds" };
+
+                result.add_edge(r, e);
+            }
+        }
+
+        std::string rest;
+        if (infile >> rest)
+            throw GraphFileError{ filename, "EOF not reached, next text is \"" + rest + "\"" };
+        if (! infile.eof())
+            throw GraphFileError{ filename, "EOF not reached" };
+
+        return result;
+    }
+
+}
+
[…]

### 2016-01-13 paper/paper.tex
State space
More outline

@@
 
+?? Actually, it's not necessarily obvious that these are the right parameters to use... For example,
+mean degree is a meaningful parameter for random graphs... Also, the formula shows it's not exactly
+the ``relative'' order, but some other function of the two orders.
+
 For certain choices of parameters for induced isomorphisms, there are two phase transitions, going
@@
 $v$ vertices, and an edge between each distinct pair of vertices with independent probability $p$.
+
+?? I'm sure I figured out how to make overlines typeset better. Remember to look this up.
 
@@
 
-Some stuff about the complexity peak in the Glasgow solver. This looks remarkably similar to random
+?? Some stuff about the complexity peak in the Glasgow solver. This looks remarkably similar to random
 3SAT problems---compare, for example, Figure 1 of \citet{LeytonBrown:2014}. In particular,
@@
 patterns of order 10, 20 and 30, a target of order 150, and varying pattern (x-axis) and target
-(y-axis) edge probabilities. Inside the orange region, at the bottom right of each plot, every
-instance is unsatisfiable---here we are trying to find a dense pattern in a sparse target. In the
-purple region, at the top left, every instance is satisfiable---we are looking for a sparse pattern
-in a dense target (which is easy, since we only have to preserve adjacency, not non-adjacency). The
-white band between the regions shows the location of the phase transition: here, roughly half the
-instances are satisfiable. (We discuss the black line below.)
+(y-axis) edge probabilities. Each axis runs over 51 edge probabilities, from 0 to 1 in steps of
+0.05; for each of these 2601 points, we generate ten random instances. The colour denotes (?? wrong
+word?) the proportion of these instances which were found to be satisfiable.  Inside the orange
+region, at the bottom right of each plot, every instance is unsatisfiable---here we are trying to
+find a dense pattern in a sparse target. In the purple region, at the top left, every instance is
+satisfiable---we are looking for a sparse pattern in a dense target (which is easy, since we only
+have to preserve adjacency, not non-adjacency). The white band between the regions shows the
+location of the phase transition: here, roughly half the instances are satisfiable. (We discuss the
+black line below.)
 
-On subsequent rows, we show the average difficulty of different algorithms.  We measure search
-nodes. Cannot be used to compare performance of algorithms directly, but this does show that each
-solver found the same set of instances difficult.  Satisfiable is easy, until very close to the
-phase transition. Unsatisfiable is much harder than satisfiable. This is solver-independent,
-although VF2 is terrible.
+On subsequent rows, we show the average number of search nodes used by the different algorithms. In
+general, satisfiable instances are easy, until very close to the phase transition. As we hit the
+phase transition and move into the unsatisfiable region, we see a complexity increase. Finally, as
+we pass the phase transition and move deeper into the unsatisfiable region, instances become easier
+again. This behaviour is largely solver-independent, although VF2 has a larger hard region than
+Glasgow or LAD. Thus, although we have moved away from the single control parameter typically
+studied, ?? I guess we want to say it's not that surprising?
 
@@
 To predict the location of the phase transition, we can calculate the expected number of solutions
-by making a few dubious assumptions involving rounding and independence, and drawing inspiration
-from the association graph encoding. Since we are trying to find an \emph{injective} mapping from a
-pattern $P = G(p, d_p)$ to a target $T = G(t, d_t)$, there are $t^{\underline{p}} = t \cdot (t -
-1) \cdot \ldots \cdot (t - p + 1)$ possible assignments of target vertices to pattern vertices.  We expect the
-pattern to have $d_p \cdot \binom{p}{2}$ edges, so we obtain the probability of all of these edges
-being mapped to edges in the target by raising $d_t$ to this power, giving an expected number of
-solutions of \[ \langle Sol \rangle = t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}}
-\textnormal{.} \] This formula predicts a very sharp phase transition from $\langle Sol \rangle \ll
-1$ to $\langle Sol \rangle \gg 1$. We plot where this occurs using black lines in the first row of
-\cref{figure:non-induced}.
+by making a few dubious assumptions involving rounding and independence. Since we are trying to find
+an \emph{injective} mapping from a pattern $P = G(p, d_p)$ to a target $T = G(t, d_t)$, there are
+$t^{\underline{p}} = t \cdot (t - 1) \cdot \ldots \cdot (t - p + 1)$ possible assignments of target
+vertices to pattern vertices.  We expect the pattern to have $d_p \cdot \binom{p}{2}$ edges, so we
+obtain the probability of all of these edges being mapped to edges in the target by raising $d_t$ to
+this power, giving an expected number of solutions of \[ \langle Sol \rangle = t^{\underline{p}}
+\cdot {d_t}^{d_p \cdot \binom{p}{2}} \textnormal{.} \] This formula predicts a very sharp phase
+transition from $\langle Sol \rangle \ll 1$ to $\langle Sol \rangle \gg 1$. We plot where this
+occurs using black lines in the first row of \cref{figure:non-induced}.
 
@@
 
+?? Note that the size of the state space is defined differently to in binary CSPs: we're using the
+all diff constraint to determine the size of the state space, and are not treating it as a
+constraint. In other words, our generate-and-test approach is alldiff-aware. There are probably five
+people in the world who care about this and would have seen it as very interesting twenty years ago.
+
 \subsection{Variable and Value Ordering Heuristics}
 
-?? Note that the Glasgow heuristics are moving us towards the satisfiable region, or increasing the
-expected number of solutions \citep{Gent:1996:EN}.
-
-?? LAD does SDF with no tiebreaking, no value ordering heuristic.
+?? Point out how the value ordering heuristics, which were developed empirically, can be recovered
+just by using this formula.  Note that the Glasgow heuristics are moving us towards the satisfiable
+region, or increasing the expected number of solutions \citep{Gent:1996:EN}. LAD does SDF with no
+tiebreaking, no value ordering heuristic.
 
@@
 
-?? I suppose we should mention kappa \citep{Gent:1996:Kappa}, etc, here. This whole thing is a bit
-subtle: maximising E(N) or Kappa as a principle is encoding-independent, but whether this is done by
-SDF, deg, etc, is encoding-dependent.
+?? I suppose we should mention kappa \citep{Gent:1996:Kappa}, etc, here (or possibly in the
+conclusion?). This whole thing is a bit subtle: maximising E(N) or Kappa as a principle is
+encoding-independent, but whether this is done by SDF, deg, etc, is encoding-dependent. This seems
+to be known by these same five people, but maybe isn't in the literature anywhere.
 
@@
 
-In \cref{figure:induced} we show the induced behaviour (modified Glasgow to include the complement
-as a supplemental graph).
+In \cref{figure:induced} we repeat our experiments, finding induced isomorphisms. With a pattern of
+order 10, we get two independent phase transitions (?? describe properly). For each of the solvers,
+the satisfiable instances before either phase transition are generally easy, instances near the
+phase transition are hard, and the large central unsatisfiable region is again easy.
 
-Pattern order 10, what we expect. Patterns 20 and 30 have large unsatisfiable regions in the middle,
-but these remain hard. Pattern orders 14, 15 and 16 show the transition between the two styles.
+?? For larger patterns of order 20 and 30, we have large unsatisfiable regions in the middle,
+but these remain hard.
+
+Pattern orders 14, 15 and 16 show the transition between the two styles.
 
@@
 
-?? This suggests that we're not able to make use of adjacency and non-adjacency simultaneously. Note
-that this is solver-independent. In the following section we see whether this is an oddity of
-subgraph isomorphism algorithms.
+?? Note that the large unexpectedly hard region in the middle is solver-independent. Perhaps
+subgraph isomorphism algorithms are not able to make use of adjacency and non-adjacency
+simultaneously, and this is a weakness in how all three are designed? In the following section we
+debunk this wonderful hypothesis.
 
-Odd unsat with LAD at the top right (not EHPs), trivial for Glasgow: mapping a non-clique into a
-clique. Glasgow looks at non-edges and can see immediately that no mapping is possible.
+?? Does kappa predict the hard region in the middle?
+
+?? Maybe note a few solver-dependent hard instances.  There's odd unsat with LAD at the top right
+(not EHPs), trivial for Glasgow: mapping a non-clique into a clique. Glasgow looks at non-edges and
+can see immediately that no mapping is possible, but LAD doesn't do degree filtering on complements.
 
@@
 
-?? Explain why we might guess degree heuristics just won't work for induced.
+?? Explain why we might guess degree heuristics just won't work for induced: for everything we say
+about degree, the opposite holds too for the complement constraints.
 
-?? Explain bottom row of \cref{figure:induced}, and give way of calculating it.
+?? Explain bottom row of \cref{figure:induced}, and give way of calculating it. The key point is
+that depending upon where we are, changes to one of the two terms has more effect in the equation.
+
+?? Possibly send the data to Lars, and get him to run his portfolios stuff on it, and then make fun
+of machine learning?
 
@@
 upon an estimate of the solubility of the problem may offer good performance.
+
+\subsection{Non-Random Instances}
 
@@
     Glasgow algorithm, the third row shows the number of decisions made by the Clasp pseudo-boolean
-solver, and the fourth row shows the number of search nodes used by a clique
+solver, and the fourth row shows the number of search nodes used by BBMC on the clique
 encoding.}\label{figure:alt}
@@
 
-?? The odd stuff in the middle, which isn't near a phase transition, but is hard.
+?? We're looking at the odd stuff in the middle, which isn't near a phase transition, but is hard.
 
 Is this just a weakness of current subgraph isomorphism solvers? In \cref{figure:alt} we repeat the
-experiments on smaller pattern and target graphs, using different solving techniques. Although these
-techniques are not competitive in absolute terms, we wish to see if the same pattern of behaviour
-occurs.
+induced experiments on smaller pattern and target graphs, using different solving techniques.
+Although these techniques are not competitive in absolute terms, we wish to see if the same pattern
+of behaviour occurs.
 
@@
 non-edge constraints. We use the Clasp solver. We see that hard instances remain hard, including
-inside the central region, and that the easy satisfiable instances remain easy.
+inside the central region, and that the easy satisfiable instances remain easy. ?? Nag James to
+write this bit, and possibly kidnap his baby if he isn't cooperative.
 
@@
 
-Clique encoding: association graph, a version of BBMC \citep{SanSegundo:2011} modified to solve the
+?? Clique encoding: association graph, a version of BBMC \citep{SanSegundo:2011} modified to solve the
 decision problem. The hard instances in the central region remain hard. Additionally, some of the
-easy satisfiable instances become hard. Possible weakness in variable ordering heuristics.
+easy satisfiable instances become hard. Possible weakness in variable ordering heuristics. The
+interesting thing is here that clique knows nothing about adjacency or non-adjacency.
 
-Some easy instances become hard, here, but hard instances remain hard. In particular, the large
+?? Some easy instances become hard, here, but hard instances remain hard. In particular, the large
 central region for the larger pattern graphs which is not near the phase transitions is still
@@
 
-Dynamic heuristics
+?? We found hard instances, some where we'd expect, some in unexpected places. These unexpected
+instances, by and large, remained hard under reductions. We looked at how existing non-induced
+heuristics could be recovered from Sol, and how this gives us better heuristics for the induced
+variant.
 
-Other random models and non-random
+?? Consider dynamic heuristics.
 
-Does not give hard instances for graph isomorphism
+?? Other random models and non-random.
+
+?? Does not give hard instances for graph isomorphism.
 
 ?? \citep{Lipets:2009} inexact solutions?
+
+\section*{Acknowledgements}
+
+?? I suppose Craig could be given an inkling of undeserved gratitude if we have space left.
 

### 2016-01-14 paper/paper.tex
More writing about other solvers

@@
 
-?? We're looking at the odd stuff in the middle, which isn't near a phase transition, but is hard.
-
-Is this just a weakness of current subgraph isomorphism solvers? In \cref{figure:alt} we repeat the
+The region in the parameter space where both pattern and target have medium density is far from a
+phase transition, but nevertheless contains instances that are hard for current subgraph isomorphism
+solvers. We would like to know whether this is due to a weakness in current solvers, or whether
+instances in this region are inherently difficult to solve. In \cref{figure:alt} we repeat the
 induced experiments on smaller pattern and target graphs, using different solving techniques.
@@
 
-Direct pseudo-boolean encoding: variables for each assignment, cardinality constraints, edge and
-non-edge constraints. We use the Clasp solver. We see that hard instances remain hard, including
-inside the central region, and that the easy satisfiable instances remain easy. ?? Nag James to
-write this bit, and possibly kidnap his baby if he isn't cooperative.
+We create a pseudo-boolean (PB) encoding as follows. For each pattern vertex $v$ and each target
+vertex $w$, create a binary variable which takes the value 1 if and only if that $v$ is mapped to
+$w$.  Constraints are added to ensure that each pattern vertex maps to exactly one target vertex,
+that each target vertex is mapped to by at most one pattern vertex, and that adjacent edges are
+mapped to adjacent edges. Because we are solving the induced problem, we also add constraints to
+ensure that non-adjacent edges are mapped to non-adjacent edges. We used the Clasp
+solver~\citep{gekakaosscsc11a} to solve the pseudo-boolean instances. The instances that are hard
+for the Glasgow solver remain hard for the PB solver, including instances inside the central region,
+and the easy satisfiable instances remain easy.
 
-Although not presented here, we tried an equivalent SAT encoding with direct encoding of the
-cardinality constraints using the Glucose solver, and saw very similar performance.  We also tried
-using a mixed integer encoding along similar lines---Gurobi was only able to solve some of the
-trivial satisfiable instances, and was almost never able to prove unsatisfiability within the time
-limit. We do not plot these results.
+We implemented an equivalent SAT encoding with direct encoding of the cardinality constraints. Using
+the Glucose solver, we saw very similar performance to the PB encoding. We also implemented a
+similar integer program encoding; the Gurobi solver was only able to solve some of the trivial
+satisfiable instances, and was almost never able to prove unsatisfiability within the time limit. We
+do not plot the results for either of these solvers.
 

### 2016-01-14 paper/paper.tex
More outline
Typso

@@
 \author{Ciaran McCreesh\thanks{This work was supported by the Engineering and Physical Sciences
-    Research Council [grant number EP/K503058/1]} \and Patrick Prossser \and James Trimble \\
+    Research Council [grant number EP/K503058/1]} \and Patrick Prosser \and James Trimble \\
 University of Glasgow, Glasgow, Scotland \\
@@
 that depending upon where we are, changes to one of the two terms has more effect in the equation.
+Note how sharp and well-defined this is.
 

### 2016-01-14 paper/paper.tex
Betterer writing

@@
 We create a pseudo-boolean (PB) encoding as follows. For each pattern vertex $v$ and each target
-vertex $w$, create a binary variable which takes the value 1 if and only if that $v$ is mapped to
+vertex $w$, we have a binary variable which takes the value 1 if and only if that $v$ is mapped to
 $w$.  Constraints are added to ensure that each pattern vertex maps to exactly one target vertex,
-that each target vertex is mapped to by at most one pattern vertex, and that adjacent edges are
-mapped to adjacent edges. Because we are solving the induced problem, we also add constraints to
-ensure that non-adjacent edges are mapped to non-adjacent edges. We used the Clasp
-solver~\citep{gekakaosscsc11a} to solve the pseudo-boolean instances. The instances that are hard
-for the Glasgow solver remain hard for the PB solver, including instances inside the central region,
-and the easy satisfiable instances remain easy.
+that each target vertex is mapped to by at most one pattern vertex, that adjacent vertices are
+mapped to adjacent vertices, and that non-adjacent vertices are mapped to non-adjacent vertices. We
+used the Clasp solver~\citep{gekakaosscsc11a} to solve the pseudo-boolean instances. The instances
+that are hard for the Glasgow solver remain hard for the PB solver, including instances inside the
+central region, and the easy satisfiable instances remain easy.
 
@@
 
-?? Clique encoding: association graph, a version of BBMC \citep{SanSegundo:2011} modified to solve the
-decision problem. The hard instances in the central region remain hard. Additionally, some of the
-easy satisfiable instances become hard. Possible weakness in variable ordering heuristics. The
-interesting thing is here that clique knows nothing about adjacency or non-adjacency.
+We also reduced the problem to the clique problem via the association graph encoding show in
+\cref{figure:sip}, and solved the instances using a version of BBMC \citep{SanSegundo:2011} modified
+to solve the decision problem. Again, the instances in the central region remain hard, and
+additionally, some of the easy satisfiable instances become hard.
 
-?? Some easy instances become hard, here, but hard instances remain hard. In particular, the large
-central region for the larger pattern graphs which is not near the phase transitions is still
-relatively difficult in all encodings. This suggests that these instances may genuinely be hard,
-despite not being near a phase transition, and that this does not indicate a weakness in subgraph
-isomorphism solvers.
+These results suggest that the central region may be genuinely hard, despite not being near a phase
+transition.  The clique encoding in particular rules out the hypothesis that subgraph isomorphism
+solvers only find this region hard due to not reasoning simultaneously about adjacency and
+non-adjacency, since the association graph encoding constraints consider compatibility rather than
+adjacency and non-adjacency.
 

### 2016-01-16 paper/paper.tex
Constrainedness, etc

@@
 
+% cref style
+\crefname{figure}{Figure}{Figures}
+\Crefname{figure}{Figure}{Figures}
+
+% http://tex.stackexchange.com/questions/22100/the-bar-and-overline-commands
+\newcommand{\shortoverline}[1]{\mkern 1.5mu\overline{\mkern-1.5mu#1\mkern-1.5mu}\mkern 1.5mu}
+
 \title{Heuristics and Really Hard Instances for Subgraph Isomorphism Problems}
@@
     We show how to generate ``really hard'' random instances for variants of the subgraph
-    isomorphism problem. For the non-induced variant of the problem, we predict and observe a phase
+    isomorphism problem. For the non-induced variant, we predict and observe a phase
     transition between satisfiable and unsatisfiable instances, with a corresponding complexity peak
-    seen in three different solvers. For the induced variant of the problem, the behaviour is more
-    complicated, with two phase transitions and an unexpectedly hard region in between. We show that
-    this region remains hard when using pseudo-boolean and mixed integer encodings, and under
-    reduction to the clique problem. Finally, we show that existing heuristics for the non-induced
-    variant can be recovered by determining how to maximise the expected number of solutions, and
-    that this technique gives new, better heuristics for the induced variant.
+    in three different solvers. For the induced variant, much richer behaviour is observed, and we
+    see evidence that constrainedness is a better measure of difficulty than proximity to a phase
+    transition. Finally, we show that existing heuristics for the non-induced variant can be
+    recovered by determining how to maximise the expected number of solutions, and that this
+    technique gives new, better heuristics for the induced variant.
 \end{abstract}
@@
         \end{scope}
-    \end{tikzpicture}\\[0.75cm]
-    \begin{tikzpicture}[scale=0.5]
-        \begin{scope}
-            \foreach \n in {1, ..., 6}{
-                \node [draw, circle, fill=uofgcobalt, inner sep=1.5pt] (N1\n) at (\n, 0) {\n};
-            }
-
-            \draw [decorate, decoration={brace, raise=0.2cm}, very thick] (N11.north west) -- (N16.north
-                east) node [midway, above=0.3cm] { $1 \mapsto$ };
-
-            \foreach \n in {1, ..., 6}{
-                \node [draw, circle, fill=uofgpumpkin, inner sep=1.5pt] (N2\n) at (0, -\n) {\n};
-            }
-
-            \draw [decorate, decoration={brace, raise=0.2cm, mirror}, very thick] (N21.north west) --
-                (N26.south west) node [midway, left=0.3cm] { $2 \mapsto$ };
-
-            \foreach \n in {1, ..., 6}{
-                \node [draw, circle, fill=uofglawn, inner sep=1.5pt] (N3\n) at (7, -\n) {\n};
-            }
-
-            \draw [decorate, decoration={brace, raise=0.2cm}, very thick] (N31.north east) -- (N36.south
-                east) node [midway, right=0.3cm] { $3 \mapsto$ };
-
-            \foreach \n in {1, ..., 6}{
-                \node [draw, circle, fill=uofgthistle, inner sep=1.5pt] (N4\n) at (\n, -7) {\n};
-            }
-
-            \draw [decorate, decoration={brace, raise=0.2cm, mirror}, very thick] (N41.south west) --
-                (N46.south east) node [midway, below=0.3cm] { $4 \mapsto$ };
-
-            \begin{scope}[on background layer]
-                \draw [color=uofgsandstone!50] ($(N11)!0.8!(N11.south)$) -- ($(N22)!0.8!(N22.east)$);
-                \draw [color=uofgsandstone!50] ($(N11)!0.8!(N11.south)$) -- ($(N23)!0.8!(N23.east)$);
-                \draw [color=uofgsandstone!50] ($(N11)!0.8!(N11.south)$) -- ($(N24)!0.8!(N24.east)$);
-                \draw [color=uofgsandstone!50] ($(N11)!0.8!(N11.south)$) -- ($(N25)!0.8!(N25.east)$);
-                \draw [color=uofgsandstone!50] ($(N11)!0.8!(N11.south)$) -- ($(N36)!0.8!(N36.west)$);
-                \draw [color=uofgsandstone!50] ($(N11)!0.8!(N11.south)$) -- ($(N46)!0.8!(N46.north)$);
-                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N12)!0.8!(N12.south)$);
-                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N32)!0.8!(N32.west)$);
-                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N42)!0.8!(N42.north)$);
-                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N13)!0.8!(N13.south)$);
-                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N33)!0.8!(N33.west)$);
-                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N43)!0.8!(N43.north)$);
-                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N14)!0.8!(N14.south)$);
-                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N34)!0.8!(N34.west)$);
-                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N44)!0.8!(N44.north)$);
-                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N15)!0.8!(N15.south)$);
-                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N35)!0.8!(N35.west)$);
-                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N45)!0.8!(N45.north)$);
-                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N22)!0.8!(N22.east)$);
-                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N42)!0.8!(N42.north)$);
-                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N23)!0.8!(N23.east)$);
-                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N43)!0.8!(N43.north)$);
-                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N24)!0.8!(N24.east)$);
-                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N44)!0.8!(N44.north)$);
-                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N25)!0.8!(N25.east)$);
-                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N45)!0.8!(N45.north)$);
-                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N16)!0.8!(N16.south)$);
-                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N22)!0.8!(N22.east)$);
-                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N32)!0.8!(N32.west)$);
-                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N23)!0.8!(N23.east)$);
-                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N33)!0.8!(N33.west)$);
-                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N24)!0.8!(N24.east)$);
-                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N34)!0.8!(N34.west)$);
-                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N25)!0.8!(N25.east)$);
-                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N35)!0.8!(N35.west)$);
-                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N16)!0.8!(N16.south)$);
-                \draw [color=uofgsandstone!50] ($(N12)!0.8!(N12.south)$) -- ($(N33)!0.8!(N33.west)$);
-                \draw [color=uofgsandstone!50] ($(N12)!0.8!(N12.south)$) -- ($(N24)!0.8!(N24.east)$);
-                \draw [color=uofgsandstone!50] ($(N12)!0.8!(N12.south)$) -- ($(N45)!0.8!(N45.north)$);
-                \draw [color=uofgsandstone!50] ($(N22)!0.8!(N22.east)$) -- ($(N14)!0.8!(N14.south)$);
-                \draw [color=uofgsandstone!50] ($(N22)!0.8!(N22.east)$) -- ($(N34)!0.8!(N34.west)$);
-                \draw [color=uofgsandstone!50] ($(N22)!0.8!(N22.east)$) -- ($(N44)!0.8!(N44.north)$);
-                \draw [color=uofgsandstone!50] ($(N22)!0.8!(N22.east)$) -- ($(N16)!0.8!(N16.south)$);
-                \draw [color=uofgsandstone!50] ($(N22)!0.8!(N22.east)$) -- ($(N36)!0.8!(N36.west)$);
-                \draw [color=uofgsandstone!50] ($(N22)!0.8!(N22.east)$) -- ($(N46)!0.8!(N46.north)$);
-                \draw [color=uofgsandstone!50] ($(N32)!0.8!(N32.west)$) -- ($(N13)!0.8!(N13.south)$);
-                \draw [color=uofgsandstone!50] ($(N32)!0.8!(N32.west)$) -- ($(N24)!0.8!(N24.east)$);
-                \draw [color=uofgsandstone!50] ($(N32)!0.8!(N32.west)$) -- ($(N44)!0.8!(N44.north)$);
-                \draw [color=uofgsandstone!50] ($(N32)!0.8!(N32.west)$) -- ($(N15)!0.8!(N15.south)$);
-                \draw [color=uofgsandstone!50] ($(N32)!0.8!(N32.west)$) -- ($(N26)!0.8!(N26.east)$);
-                \draw [color=uofgsandstone!50] ($(N32)!0.8!(N32.west)$) -- ($(N46)!0.8!(N46.north)$);
-                \draw [color=uofgsandstone!50] ($(N42)!0.8!(N42.north)$) -- ($(N13)!0.8!(N13.south)$);
-                \draw [color=uofgsandstone!50] ($(N42)!0.8!(N42.north)$) -- ($(N24)!0.8!(N24.east)$);
-                \draw [color=uofgsandstone!50] ($(N42)!0.8!(N42.north)$) -- ($(N34)!0.8!(N34.west)$);
-                \draw [color=uofgsandstone!50] ($(N42)!0.8!(N42.north)$) -- ($(N15)!0.8!(N15.south)$);
-                \draw [color=uofgsandstone!50] ($(N42)!0.8!(N42.north)$) -- ($(N26)!0.8!(N26.east)$);
-                \draw [color=uofgsandstone!50] ($(N42)!0.8!(N42.north)$) -- ($(N36)!0.8!(N36.west)$);
-                \draw [color=uofgsandstone!50] ($(N13)!0.8!(N13.south)$) -- ($(N34)!0.8!(N34.west)$);
-                \draw [color=uofgsandstone!50] ($(N13)!0.8!(N13.south)$) -- ($(N44)!0.8!(N44.north)$);
-                \draw [color=uofgsandstone!50] ($(N13)!0.8!(N13.south)$) -- ($(N25)!0.8!(N25.east)$);
-                \draw [color=uofgsandstone!50] ($(N13)!0.8!(N13.south)$) -- ($(N26)!0.8!(N26.east)$);
-                \draw [color=uofgsandstone!50] ($(N23)!0.8!(N23.east)$) -- ($(N15)!0.8!(N15.south)$);
-                \draw [color=uofgsandstone!50] ($(N23)!0.8!(N23.east)$) -- ($(N35)!0.8!(N35.west)$);
-                \draw [color=uofgsandstone!50] ($(N23)!0.8!(N23.east)$) -- ($(N45)!0.8!(N45.north)$);
-                \draw [color=uofgsandstone!50] ($(N23)!0.8!(N23.east)$) -- ($(N16)!0.8!(N16.south)$);
-                \draw [color=uofgsandstone!50] ($(N23)!0.8!(N23.east)$) -- ($(N36)!0.8!(N36.west)$);
-                \draw [color=uofgsandstone!50] ($(N23)!0.8!(N23.east)$) -- ($(N46)!0.8!(N46.north)$);
-                \draw [color=uofgsandstone!50] ($(N33)!0.8!(N33.west)$) -- ($(N14)!0.8!(N14.south)$);
-                \draw [color=uofgsandstone!50] ($(N33)!0.8!(N33.west)$) -- ($(N25)!0.8!(N25.east)$);
-                \draw [color=uofgsandstone!50] ($(N33)!0.8!(N33.west)$) -- ($(N45)!0.8!(N45.north)$);
-                \draw [color=uofgsandstone!50] ($(N33)!0.8!(N33.west)$) -- ($(N26)!0.8!(N26.east)$);
-                \draw [color=uofgsandstone!50] ($(N33)!0.8!(N33.west)$) -- ($(N46)!0.8!(N46.north)$);
-                \draw [color=uofgsandstone!50] ($(N43)!0.8!(N43.north)$) -- ($(N14)!0.8!(N14.south)$);
-                \draw [color=uofgsandstone!50] ($(N43)!0.8!(N43.north)$) -- ($(N25)!0.8!(N25.east)$);
-                \draw [color=uofgsandstone!50] ($(N43)!0.8!(N43.north)$) -- ($(N36)!0.8!(N36.west)$);
-                \draw [color=uofgsandstone!50] ($(N14)!0.8!(N14.south)$) -- ($(N35)!0.8!(N35.west)$);
-                \draw [color=uofgsandstone!50] ($(N14)!0.8!(N14.south)$) -- ($(N45)!0.8!(N45.north)$);
-                \draw [color=uofgsandstone!50] ($(N14)!0.8!(N14.south)$) -- ($(N26)!0.8!(N26.east)$);
-                \draw [color=uofgsandstone!50] ($(N24)!0.8!(N24.east)$) -- ($(N16)!0.8!(N16.south)$);
-                \draw [color=uofgsandstone!50] ($(N24)!0.8!(N24.east)$) -- ($(N36)!0.8!(N36.west)$);
-                \draw [color=uofgsandstone!50] ($(N24)!0.8!(N24.east)$) -- ($(N46)!0.8!(N46.north)$);
-                \draw [color=uofgsandstone!50] ($(N34)!0.8!(N34.west)$) -- ($(N15)!0.8!(N15.south)$);
-                \draw [color=uofgsandstone!50] ($(N34)!0.8!(N34.west)$) -- ($(N26)!0.8!(N26.east)$);
-                \draw [color=uofgsandstone!50] ($(N34)!0.8!(N34.west)$) -- ($(N46)!0.8!(N46.north)$);
-                \draw [color=uofgsandstone!50] ($(N44)!0.8!(N44.north)$) -- ($(N15)!0.8!(N15.south)$);
-                \draw [color=uofgsandstone!50] ($(N44)!0.8!(N44.north)$) -- ($(N26)!0.8!(N26.east)$);
-                \draw [color=uofgsandstone!50] ($(N44)!0.8!(N44.north)$) -- ($(N36)!0.8!(N36.west)$);
-                \draw [color=uofgsandstone!50] ($(N15)!0.8!(N15.south)$) -- ($(N26)!0.8!(N26.east)$);
-                \draw [color=uofgsandstone!50] ($(N25)!0.8!(N25.east)$) -- ($(N16)!0.8!(N16.south)$);
-                \draw [color=uofgsandstone!50] ($(N25)!0.8!(N25.east)$) -- ($(N36)!0.8!(N36.west)$);
-                \draw [color=uofgsandstone!50] ($(N25)!0.8!(N25.east)$) -- ($(N46)!0.8!(N46.north)$);
-                \draw [color=uofgsandstone!50] ($(N35)!0.8!(N35.west)$) -- ($(N46)!0.8!(N46.north)$);
-                \draw [color=uofgsandstone!50] ($(N45)!0.8!(N45.north)$) -- ($(N26)!0.8!(N26.east)$);
-                \draw [color=uofgsandstone!50] ($(N45)!0.8!(N45.north)$) -- ($(N36)!0.8!(N36.west)$);
-                \draw [ultra thick] ($(N12)!0.8!(N12.south)$) -- ($(N43)!0.8!(N43.north)$);
-                \draw [ultra thick] ($(N12)!0.8!(N12.south)$) -- ($(N35)!0.8!(N35.west)$);
-                \draw [ultra thick] ($(N12)!0.8!(N12.south)$) -- ($(N26)!0.8!(N26.east)$);
-                \draw [ultra thick] ($(N43)!0.8!(N43.north)$) -- ($(N35)!0.8!(N35.west)$);
-                \draw [ultra thick] ($(N43)!0.8!(N43.north)$) -- ($(N26)!0.8!(N26.east)$);
-                \draw [ultra thick] ($(N35)!0.8!(N35.west)$) -- ($(N26)!0.8!(N26.east)$);
-            \end{scope}
-        \end{scope}
     \end{tikzpicture}
 
-    \caption{On top, there is a non-induced isomorphism from the pattern graph, in the center, to the target
+    \caption{There is a non-induced isomorphism from the pattern graph, in the center, to the target
     on the right, mapping vertex 1 to 1, 2 to 3, 3 to 5 and 4 to 6. This is not an induced
     isomorphism, since there is an edge between 1 and 5 in the target but not between 1 and 3 in the
-    pattern. The mapping from the pattern to the (same) target on the left, sending 1 to 2, 2 to 6, 3 to
-    5 and 4 to 3, is both non-induced and induced.
-    Below, the association graph encoding for the induced version: the highlighted clique
-    corresponds to the same mapping.}
+    pattern. The mapping from the pattern to the (same) target on the left, sending 1 to 2, 2 to 6,
+    3 to 5 and 4 to 3, is both non-induced and induced.}
     \label{figure:sip}
@@
 which were generated by taking a random subgraph of a random graph (using various models, including
-Erd\H{o}s-R\'enyi, scale-free, bounded degree, and meshes) and permuting the vertices.  This is not
-ideal: such instances are guaranteed to be satisfiable (?? mention \citep{Anton:2009}), and so
-existing benchmark suites contain relatively few non-trivial unsatisfiable instances. These
-instances also tend to be computationally fairly easy, with most of the difficulty being in dealing
-with the size of the model.
-
-Generating non-trivial unsatisfiable random instances is not straightforward. Using a
-pattern graph from one of the random suites with the ``wrong'' target graph tends to give either a
-trivially unsatisfiable instance, or a satisfiable instance. (In particular, it is \emph{not} the
-case that a relatively small random graph is unlikely to appear in a larger random graph.)
+Erd\H{o}s-R\'enyi, scale-free, bounded degree, and meshes) and permuting the vertices. Such
+instances are guaranteed to be satisfiable---\citet{Anton:2009} exploited this property to create
+large sets of random satisfiable SAT instances.  However, since this has been the only approach used
+to generate random subgraph isomorphism instances, existing benchmark suites contain relatively few
+non-trivial unsatisfiable instances, and the satisfiable instances tend to be computationally fairly
+easy, with most of the difficulty being in dealing with the size of the model.  This problem cannot
+be addressed by using a pattern graph from one of the random suites with the ``wrong'' target graph:
+this tends to give either a trivially unsatisfiable instance, or a satisfiable instance. (In
+particular, it is \emph{not} the case that a relatively small random graph is unlikely to appear in
+a larger random graph.)
 
@@
 
-?? Maybe introduce \cref{figure:phase-transition} here instead of at the start of the next section?
-
-For subgraph isomorphism we have three relevant control parameters: we can independently alter the
+For subgraph isomorphism we identify relevant control parameters: we can independently alter the
 edge probability of the pattern graph, the edge probability of the target graph, and the relative
@@
 satisfiability problems: there is a phase transition from satisfiable to unsatisfiable (and we can
-predict its location), and we see a complexity peak occur (with three different solvers) near this
+predict its location), and we see a complexity peak occur with three different solvers near this
 phase transition.
-
-?? Actually, it's not necessarily obvious that these are the right parameters to use... For example,
-mean degree is a meaningful parameter for random graphs... Also, the formula shows it's not exactly
-the ``relative'' order, but some other function of the two orders.
 
@@
 unsatisfiable region to be really hard---this may be unexpected, since this region is not near a
-phase transition and is apparently over-constrained. To show that this is not a simple weakness of
-current subgraph isomorphism algorithms, we verify that this region is also hard when using use a
-pseudo-boolean encoding, a mixed integer model, and under reduction to the clique problem.
+phase transition. To show that this is not a simple weakness of current subgraph isomorphism
+algorithms, we verify that this region is also hard when using use a pseudo-boolean encoding, and
+under reduction to the clique problem. Interestingly, the constrainedness measure
+\citep{Gent:1996:Kappa} \emph{does} predict this difficult region---these instances provide (?? the
+first??) evidence in favour of constrainedness, rather than proximity to a phase transition, being
+an accurate predictor of difficulty.
 
-?? Also there's heuristics. Figure out how to work this smoothly into the intro.
+?? Finally, we discuss the relationship between heuristics and the expected number of solutions,
+recovering existing heuristics for the non-induced variant, and introducing new heuristics for the
+induced variant. I have failed miserably at integrating this paragraph smoothly.
 
@@
 $\operatorname{V}(G)$ for the vertex set of a graph $G$. The \emph{complement} of a graph $G$,
-denoted $\overline{G}$, is the
+denoted $\shortoverline{G}$, is the
 graph with the same vertex set as $G$, and with an edge between distinct vertices $v$ and $w$ iff
@@
 $v$ vertices, and an edge between each distinct pair of vertices with independent probability $p$.
-
-?? I'm sure I figured out how to make overlines typeset better. Remember to look this up.
 
@@
 that an induced isomorphism $i : P \hookrightarrow T$ is a non-induced isomorphism $i : P
-\rightarrowtail T$ which is also a non-induced isomorphism $\overline{i} : \overline{P}
-\rightarrowtail \overline{T}$.
-
-The \emph{association graph encoding} of a subgraph isomorphism problem is constructed by creating a
-new graph with a vertex for each pair $(p, t)$ of vertices from the pattern and target graphs
-respectively. There is an edge between vertex $(p_1, t_1)$ and vertex $(p_2, t_2)$ if mapping $p_1$
-to $t_1$ and $p_2$ to $t_2$ simultaneously is permitted. For the non-induced problem, permitted
-means either $p_1$ is not adjacent to $p_2$, or both $p_1$ and $p_2$, and $t_1$ and $t_2$, are
-adjacent; for the induced problem, it means that $p_1$ is adjacent to $p_2$ if and only if $t_1$ is
-adjacent to $t_2$. Either way, it can be seen that a clique (a subset of vertices, each of which is
-adjacent to every other in the subset) of size equal to the order of the pattern graph exists in the
-association graph if and only if the problem is satisfiable, and the vertices selected correspond to
-the assignments making the solution \citep{Levi:1973}. We illustrate this in
-\cref{figure:sip}. ?? Can we move this to section 4? I thought we'd need this for the sol formulae,
-but it appears not.
+\rightarrowtail T$ which is also a non-induced isomorphism $\shortoverline{i} : \shortoverline{P}
+\rightarrowtail \shortoverline{T}$.
 
@@
 ?? Stuff about runtimes and recursive calls. We are not aiming to compare solvers; rather, we are
-looking for solver-independent behaviour.
+looking for solver-independent behaviour. We need a quick note on the models used here too.
 
@@
 \begin{figure}[tb]
+    \setlength{\abovecaptionskip}{6pt}
     \input{gen-graph-phase-transition.tex}
@@
         complexity peak with the Glasgow solver in the non-induced variant. Each point represents
-        one instance. The lines show mean search effort, and mean proportion satisfiable, using a
-    larger sample size.}
+        one instance. The lines show mean search effort, and mean proportion satisfiable.}
     \label{figure:phase-transition}
@@
 
+?? Are unsatisfiable instances on the phase transition much harder than satisfiable ones?
+
 \begin{figure}[tb]
+    \setlength{\abovecaptionskip}{0pt}
     \input{gen-graph-non-induced.tex}
@@
 Glasgow or LAD. Thus, although we have moved away from the single control parameter typically
-studied, ?? I guess we want to say it's not that surprising?
-
-?? Are unsatisfiable instances on the phase transition much harder than satisfiable ones?
+studied, we observe similar behaviour to other \NP-complete problems.
 
@@
 all diff constraint to determine the size of the state space, and are not treating it as a
-constraint. In other words, our generate-and-test approach is alldiff-aware. There are probably five
-people in the world who care about this and would have seen it as very interesting twenty years ago.
+constraint. In other words, our generate-and-test approach is alldiff-aware. This matters for kappa.
 
@@
 
-?? Point out how the value ordering heuristics, which were developed empirically, can be recovered
-just by using this formula.  Note that the Glasgow heuristics are moving us towards the satisfiable
-region, or increasing the expected number of solutions \citep{Gent:1996:EN}. LAD does SDF with no
-tiebreaking, no value ordering heuristic.
+When designing variable and value ordering heuristics for backtracking search algorithms, there are
+various general principles which appear to be effective---one of these is to try to maximise the
+expected number of solutions inside any subproblem considered during search \citep{Gent:1996:EN}.
+This is usually done by cheaper surrogates, rather than direct calculation. When branching, both LAD
+and Glasgow will pick a variable with fewest remaining values in its domain: doing this will
+generally reduce the first part of the $\langle Sol \rangle$ equation by as little as possible. When
+two or domains are of equal size, LAD simply breaks ties lexicographically. Glasgow, however, will
+pick a variable corresponding to a pattern vertex of highest degree. This strategy was determined
+empirically, but could have been derived from the $\langle Sol \rangle$ formula: picking a
+pattern vertex of high degree will make the remaining pattern subgraph sparser, which will decrease
+the exponent in the second half of the formula, maximising the overall value. LAD does not apply a
+value ordering heuristic, but Glasgow does: it prefers target vertices of lowest degree. Again, this
+was determined empirically, but it has the effect of increasing $\langle Sol \rangle$ by increasing
+the remaining target density.
 
-?? In contrast, VF2 heuristics work by selecting vertices adjacent to vertices already selected.
-
-?? I suppose we should mention kappa \citep{Gent:1996:Kappa}, etc, here (or possibly in the
-conclusion?). This whole thing is a bit subtle: maximising E(N) or Kappa as a principle is
-encoding-independent, but whether this is done by SDF, deg, etc, is encoding-dependent. This seems
-to be known by these same five people, but maybe isn't in the literature anywhere.
+The VF2 heuristics, in contrast, are based around preserving connectivity. This perhaps explains
+some of VF2's weakness on graphs that are not extremely sparse.
 
@@
 \begin{figure*}[tb]
+    \setlength{\abovecaptionskip}{0pt}
     \input{gen-graph-induced.tex}
@@
     complement graphs. The final row shows when the Glasgow algorithm performs better when given the
-    complements of the pattern and target graphs as inputs.}\label{figure:induced}
+    complements of the pattern and target graphs as inputs. ?? Also plot constrainedness
+here.}\label{figure:induced}
 \end{figure*}
@@
 
+\subsection{Predicting Induced Behaviour}
+
+By repeating the argument for locating the non-induced phase transition and additionally considering
+non-edges, we get an expected number of solutions of \[ \langle Sol \rangle = t^{\underline{p}}
+    \cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot {(1 - d_{t})}^{(1 - d_{p}) \cdot \binom{p}{2}}
+\textnormal{.} \] We plot this
+using black lines on \cref{figure:induced}---again, our prediction is accurate except for very
+sparse or very dense patterns.
+
+\subsection{Why is the Central Region Hard?}
+
 ?? Explain the flip bound on row 5, and note that the difficult region is roughly near the phase
@@
 
-?? Does kappa predict the hard region in the middle?
+?? Constrainedness predicts the hard region in the middle \citep{Gent:1996:Kappa}! Add a two line
+description of kappa, the slightly modified way we calculated it, and plot it on the graph.
 
@@
 can see immediately that no mapping is possible, but LAD doesn't do degree filtering on complements.
-
-\subsection{Predicting Induced Behaviour}
-
-By repeating the argument for locating the non-induced phase transition and additionally considering
-non-edges, we get an expected number of solutions of \[ \langle Sol \rangle = t^{\underline{p}}
-    \cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot {(1 - d_{t})}^{(1 - d_{p}) \cdot \binom{p}{2}}
-\textnormal{.} \] We plot this
-using black lines on \cref{figure:induced}---again, our prediction is accurate except for very
-sparse or very dense patterns.
 
@@
 
-?? Note that it is always best to try to move towards the satisfiable region, even for unsatisfiable
-instances. This contradicts the suggestion by \citet{Walsh:1998} that switching heuristics based
-upon an estimate of the solubility of the problem may offer good performance.
+?? Note that when the pattern is small, it is always best to try to move towards the satisfiable
+region, even for unsatisfiable instances. It's not clear what's going on when the pattern is large.
+Note the suggestion by \citet{Walsh:1998} that switching heuristics based upon an estimate of the
+solubility of the problem may offer good performance.
 
-\subsection{Non-Random Instances}
-
-?? Is this heuristic better in general, on non-random instances?
+?? Is this heuristic better in general, on non-random instances? Unfortunately most of the instances
+in the standard benchmark set are too sparse for it to kick in. However, it provides a justification
+for reusing the non-induced heuristics.
 
@@
 \begin{figure*}[t]
+    \setlength{\abovecaptionskip}{0pt}
     \input{gen-graph-sat.tex}
@@
 solvers. We would like to know whether this is due to a weakness in current solvers, or whether
-instances in this region are inherently difficult to solve. In \cref{figure:alt} we repeat the
-induced experiments on smaller pattern and target graphs, using different solving techniques.
-Although these techniques are not competitive in absolute terms, we wish to see if the same pattern
-of behaviour occurs.
+instances in this region are inherently difficult to solve, as constrainedness suggests. In
+\cref{figure:alt} we repeat the induced experiments on smaller pattern and target graphs, using
+different solving techniques.  Although these techniques are not competitive in absolute terms, we
+wish to see if the same pattern of behaviour occurs.
 
@@
 
-We also reduced the problem to the clique problem via the association graph encoding show in
-\cref{figure:sip}, and solved the instances using a version of BBMC \citep{SanSegundo:2011} modified
-to solve the decision problem. Again, the instances in the central region remain hard, and
-additionally, some of the easy satisfiable instances become hard.
+The \emph{association graph encoding} of a subgraph isomorphism problem is constructed by creating a
+new graph with a vertex for each pair $(p, t)$ of vertices from the pattern and target graphs
+respectively. There is an edge between vertex $(p_1, t_1)$ and vertex $(p_2, t_2)$ if mapping $p_1$
+to $t_1$ and $p_2$ to $t_2$ simultaneously is permitted, i.e.\ $p_1$ is adjacent to $p_2$ if and
+only if $t_1$ is adjacent to $t_2$. A clique of size equal to the order of the pattern graph exists
+in the association graph if and only if the problem is satisfiable \citep{Levi:1973}.  We used this
+encoding with a version of the BBMC clique algorithm \citep{SanSegundo:2011} modified to solve the
+decision problem. Again, the instances in the central region remain hard, and additionally, some of
+the easy satisfiable instances become hard.
 
 These results suggest that the central region may be genuinely hard, despite not being near a phase
-transition.  The clique encoding in particular rules out the hypothesis that subgraph isomorphism
-solvers only find this region hard due to not reasoning simultaneously about adjacency and
-non-adjacency, since the association graph encoding constraints consider compatibility rather than
-adjacency and non-adjacency.
+transition, and that constrainedness is the better measure of difficulty. The clique encoding in
+particular rules out the hypothesis that subgraph isomorphism solvers only find this region hard due
+to not reasoning simultaneously about adjacency and non-adjacency, since the association graph
+encoding constraints consider compatibility rather than adjacency and non-adjacency.
+
+?? Lends weight to kappa
 

### 2016-01-17 paper/paper.tex
kappa, etc

@@
     \input{gen-graph-non-induced.tex}
-    \caption{Behaviour of algorithms on the non-induced variant. Each point is the average of ten
-        runs. For each plot, the x-axis is the pattern edge probability and the y-axis is the target
-        edge probability, both from 0 to 1. Along the top row, we show the proportion of instances which are
-        satisfiable; the white bands shows the phase transitions, and the black lines are our
-        predictions of where the phase transition will occur. On the second row, we show the number
-    of search nodes used by the Glasgow algorithm, on the third row, the number of search nodes used
-by the LAD algorithm, and the fourth, VF2; the dark regions indicate ``really hard'' instances.}
+    \caption{Behaviour of algorithms on the non-induced variant. For each plot, the x-axis is the
+        pattern edge probability and the y-axis is the target edge probability, both from 0 to 1.
+        Along the top row, we show the proportion of instances which are satisfiable; the white
+        bands shows the phase transitions, and the black lines are our predictions of where the
+        phase transition will occur. On the second row, we show the number of search nodes used by the
+        Glasgow algorithm, on the third row, the number of search nodes used by the LAD algorithm, and the
+        fourth, VF2; the dark regions indicate ``really hard'' instances.}
     \label{figure:non-induced}
@@
 
-\begin{figure*}[tb]
-    \setlength{\abovecaptionskip}{0pt}
+\begin{figure*}[p]
     \input{gen-graph-induced.tex}
@@
     considering where a \emph{non-}induced isomorphism may also be a non-induced isomorphism between
-    complement graphs. The final row shows when the Glasgow algorithm performs better when given the
-    complements of the pattern and target graphs as inputs. ?? Also plot constrainedness
-here.}\label{figure:induced}
+    complement graphs. The sixth row plots constrainedness: the darkest region is where
+    constrainedness is exactly one, and the lighter regions show where the problem is either over-
+    or under-constrained. The final row shows when the Glasgow algorithm performs better when given
+    the complements of the pattern and target graphs as
+inputs.}\label{figure:induced}
 \end{figure*}
@@
 ?? Constrainedness predicts the hard region in the middle \citep{Gent:1996:Kappa}! Add a two line
-description of kappa, the slightly modified way we calculated it, and plot it on the graph.
+description of kappa, the slightly modified way we calculated it. We plot this on the sixth row of
+\cref{figure:induced}.
 

### 2016-01-17 prediction/induced-kappa.rb

@@ -1,0 +1,35 @@
+#!/usr/bin/ruby
+# vim: set sw=4 sts=4 et tw=80 :
+
+require 'bigdecimal'
+
+ps = ARGV[0].to_i
+ts = ARGV[1].to_i
+
+0.upto(50) do | y |
+    0.upto(50) do | x |
+        pd = BigDecimal.new(x) / 50;
+        td = BigDecimal.new(y) / 50;
+
+        cliqueedges = BigDecimal.new(ps * (ps - 1) / 2);
+        nedges = pd * cliqueedges;
+        nnonedges = (1 - pd) * cliqueedges;
+
+        edgesokprob = td ** nedges;
+        nonedgesokprob = (1 - td) ** nnonedges;
+
+        solprob = edgesokprob * nonedgesokprob;
+
+        nstates = BigDecimal.new(1);
+        1.upto(ps) do | n |
+            nstates *= (ts - n + 1);
+        end
+
+        expsols = nstates * solprob;
+
+        kappa = if 0 == expsols then 10 else 1 - ((BigMath.log(expsols, 50)) / BigMath.log(nstates, 50)) end
+
+        print kappa.to_f.to_s + " "
+    end
+    puts
+end

### 2016-01-18 paper/paper.tex
Tweaks

@@
 
-For subgraph isomorphism we identify relevant control parameters: we can independently alter the
-edge probability of the pattern graph, the edge probability of the target graph, and the relative
-orders (number of vertices) of the pattern and target graphs.  For non-induced isomorphisms, with
-the correct choice of parameters we see results very similar to those observed with boolean
-satisfiability problems: there is a phase transition from satisfiable to unsatisfiable (and we can
-predict its location), and we see a complexity peak occur with three different solvers near this
-phase transition.
+For subgraph isomorphism we identify three relevant control parameters: we can independently alter
+the edge probability of the pattern graph, the edge probability of the target graph, and the
+relative orders (number of vertices) of the pattern and target graphs.  For non-induced
+isomorphisms, with the correct choice of parameters we see results very similar to those observed
+with boolean satisfiability problems: there is a phase transition from satisfiable to unsatisfiable
+(and we can predict its location), and we see a complexity peak occur with three different solvers
+near this phase transition.
 
@@
 really hard to solve. However, each of the three solvers we tested also finds the central
-unsatisfiable region to be really hard---this may be unexpected, since this region is not near a
+unsatisfiable region to be hard---this may be unexpected, since this region is not near a
 phase transition. To show that this is not a simple weakness of current subgraph isomorphism

### 2016-01-19 paper/paper.tex
Fixed a few typos

@@
 phase transition. To show that this is not a simple weakness of current subgraph isomorphism
-algorithms, we verify that this region is also hard when using use a pseudo-boolean encoding, and
+algorithms, we verify that this region is also hard when using a pseudo-boolean encoding, and
 under reduction to the clique problem. Interestingly, the constrainedness measure
@@
 denoted $\shortoverline{G}$, is the
-graph with the same vertex set as $G$, and with an edge between distinct vertices $v$ and $w$ iff
+graph with the same vertex set as $G$, and with an edge between distinct vertices $v$ and $w$ if and
+only if
 $v$ and $w$ are not adjacent in $G$. We write $G(v, p)$ for an Erd\H{o}s-R\'enyi random graph with
@@
 generally reduce the first part of the $\langle Sol \rangle$ equation by as little as possible. When
-two or domains are of equal size, LAD simply breaks ties lexicographically. Glasgow, however, will
+two or more domains are of equal size, LAD simply breaks ties lexicographically. Glasgow, however, will
 pick a variable corresponding to a pattern vertex of highest degree. This strategy was determined

### 2016-01-20 paper/paper.tex
Patrick's comments

@@
 instances are guaranteed to be satisfiable---\citet{Anton:2009} exploited this property to create
-large sets of random satisfiable SAT instances.  However, since this has been the only approach used
-to generate random subgraph isomorphism instances, existing benchmark suites contain relatively few
-non-trivial unsatisfiable instances, and the satisfiable instances tend to be computationally fairly
-easy, with most of the difficulty being in dealing with the size of the model.  This problem cannot
-be addressed by using a pattern graph from one of the random suites with the ``wrong'' target graph:
-this tends to give either a trivially unsatisfiable instance, or a satisfiable instance. (In
-particular, it is \emph{not} the case that a relatively small random graph is unlikely to appear in
-a larger random graph.)
+large sets of random satisfiable boolean satisfiability instances.  However, since this has been the
+only approach used to generate random subgraph isomorphism instances, existing benchmark suites
+contain relatively few non-trivial unsatisfiable instances, and the satisfiable instances tend to be
+computationally fairly easy, with most of the difficulty being in dealing with the size of the
+model.  This problem cannot be addressed by using a pattern graph from one of the random suites with
+the ``wrong'' target graph: this tends to give either a trivially unsatisfiable instance, or a
+satisfiable instance. (In particular, it is \emph{not} the case that a relatively small random graph
+is unlikely to appear in a larger random graph.)
 
@@
 What if we alter the edge probabilities for both the pattern graph and the target graph?  In the top
-row of \cref{figure:non-induced} we show the phase transition for the non-induced variant, for
-patterns of order 10, 20 and 30, a target of order 150, and varying pattern (x-axis) and target
-(y-axis) edge probabilities. Each axis runs over 51 edge probabilities, from 0 to 1 in steps of
-0.05; for each of these 2601 points, we generate ten random instances. The colour denotes (?? wrong
-word?) the proportion of these instances which were found to be satisfiable.  Inside the orange
-region, at the bottom right of each plot, every instance is unsatisfiable---here we are trying to
-find a dense pattern in a sparse target. In the purple region, at the top left, every instance is
-satisfiable---we are looking for a sparse pattern in a dense target (which is easy, since we only
-have to preserve adjacency, not non-adjacency). The white band between the regions shows the
-location of the phase transition: here, roughly half the instances are satisfiable. (We discuss the
-black line below.)
+row of \cref{figure:non-induced} we show the satisfiability phase transition for the non-induced
+variant, for patterns of order 10, 20 and 30, a target of order 150, and varying pattern (x-axis)
+and target (y-axis) edge probabilities. Each axis runs over 51 edge probabilities, from 0 to 1 in
+steps of 0.05; for each of these 2601 points, we generate ten random instances. The colour denotes
+(?? wrong word?) the proportion of these instances which were found to be satisfiable.  Inside the
+orange region, at the bottom right of each plot, every instance is unsatisfiable---here we are
+trying to find a dense pattern in a sparse target. In the purple region, at the top left, every
+instance is satisfiable---we are looking for a sparse pattern in a dense target (which is easy,
+since we only have to preserve adjacency, not non-adjacency). The white band between the regions
+shows the location of the phase transition: here, roughly half the instances are satisfiable. (We
+discuss the black line below.)
 
@@
 phase transition and move into the unsatisfiable region, we see a complexity increase. Finally, as
-we pass the phase transition and move deeper into the unsatisfiable region, instances become easier
-again. This behaviour is largely solver-independent, although VF2 has a larger hard region than
-Glasgow or LAD. Thus, although we have moved away from the single control parameter typically
-studied, we observe similar behaviour to other \NP-complete problems.
+we pass through the phase transition and move deeper into the unsatisfiable region, instances become
+easier again. This behaviour is largely solver-independent, although VF2 has a larger hard region
+than Glasgow or LAD. Thus, although we have moved away from a single control parameter, we still
+observe the easy-hard-easy pattern seen in many other \NP-complete problems.
 

### 2016-01-22 prediction/noninduced-kappa.rb
Kappa for non-induced too

@@ -1,0 +1,33 @@
+#!/usr/bin/ruby
+# vim: set sw=4 sts=4 et tw=80 :
+
+require 'bigdecimal'
+
+ps = ARGV[0].to_i
+ts = ARGV[1].to_i
+
+0.upto(50) do | y |
+    0.upto(50) do | x |
+        pd = BigDecimal.new(x) / 50;
+        td = BigDecimal.new(y) / 50;
+
+        cliqueedges = BigDecimal.new(ps * (ps - 1) / 2);
+        nedges = pd * cliqueedges;
+
+        edgesokprob = td ** nedges;
+
+        solprob = edgesokprob;
+
+        nstates = BigDecimal.new(1);
+        1.upto(ps) do | n |
+            nstates *= (ts - n + 1);
+        end
+
+        expsols = nstates * solprob;
+
+        kappa = if 0 == expsols then 10 else 1 - ((BigMath.log(expsols, 50)) / BigMath.log(nstates, 50)) end
+
+        print kappa.to_f.to_s + " "
+    end
+    puts
+end

### 2016-01-22 lad/lad_random_subgraph_isomorphism.cc
Can complement for LAD too

@@ -9,7 +9,7 @@
 
 namespace po = boost::program_options;
 
-auto create_random_graph(int size, double density, int seed) -> std::vector<std::vector<uint8_t>>
+auto create_random_graph(int size, double density, int seed, bool complement) -> std::vector<std::vector<uint8_t>>
 {
     std::mt19937 rand;
     rand.seed(seed);
@@ -20,8 +20,11 @@ auto create_random_graph(int size, double density, int seed) -> std::vector<std:
     for (int e = 0 ; e < size ; ++e) {
         for (int f = e + 1 ; f < size ; ++f) {
             if (dist(rand) <= density) {
-                result[e][f] = 1;
-                result[f][e] = 1;
+                result[e][f] = complement ? 0 : 1;
+                result[f][e] = complement ? 0 : 1;
+            } else {
+                result[e][f] = complement ? 1 : 0;
+                result[f][e] = complement ? 1 : 0;
             }
         }
     }
@@ -47,6 +50,7 @@ auto main(int argc, char * argv[]) -> int
             ("size",    po::value<int>(),     "Number of vertices")
             ("density", po::value<double>(),  "Density")
             ("seed",    po::value<int>(),     "Seed")
+            ("complement",                    "Take the complement")
             ;
 
         all_options.add(display_options);
@@ -80,7 +84,8 @@ auto main(int argc, char * argv[]) -> int
         }
 
         /* Create graphs */
-        auto graph = create_random_graph(options_vars["size"].as<int>(), options_vars["density"].as<double>(), options_vars["seed"].as<int>());
+        auto graph = create_random_graph(options_vars["size"].as<int>(), options_vars["density"].as<double>(), options_vars["seed"].as<int>(),
+                options_vars.count("complement"));
 
         std::cout << graph.size() << std::endl;
 

### 2016-01-26 paper/paper.tex
Little tweaks

@@
 \title{Heuristics and Really Hard Instances for Subgraph Isomorphism Problems}
-\author{Ciaran McCreesh\thanks{This work was supported by the Engineering and Physical Sciences
-    Research Council [grant number EP/K503058/1]} \and Patrick Prosser \and James Trimble \\
-University of Glasgow, Glasgow, Scotland \\
-c.mccreesh.1@research.gla.ac.uk}
+
+% \author{Ciaran McCreesh\thanks{This work was supported by the Engineering and Physical Sciences
+%     Research Council [grant number EP/K503058/1]} \and Patrick Prosser \and James Trimble \\
+% University of Glasgow, Glasgow, Scotland \\
+% c.mccreesh.1@research.gla.ac.uk}
+
+\author{First Author \thanks{This work was supported by a small pile of money, but not as much as we
+    would have liked.} \and Second Author
+    \and Third Author \\
+Some Instituion, Somewhere, Some Country \\
+anonymous@example.com}
 
@@
 \begin{abstract}
-    We show how to generate ``really hard'' random instances for variants of the subgraph
-    isomorphism problem. For the non-induced variant, we predict and observe a phase
-    transition between satisfiable and unsatisfiable instances, with a corresponding complexity peak
-    in three different solvers. For the induced variant, much richer behaviour is observed, and we
-    see evidence that constrainedness is a better measure of difficulty than proximity to a phase
-    transition. Finally, we show that existing heuristics for the non-induced variant can be
-    recovered by determining how to maximise the expected number of solutions, and that this
-    technique gives new, better heuristics for the induced variant.
+    We show how to generate ``really hard'' random instances for subgraph isomorphism problems. For
+    the non-induced variant, we predict and observe a phase transition between satisfiable and
+    unsatisfiable instances, with a corresponding complexity peak seen in three different solvers. For
+    the induced variant, much richer behaviour is observed, and constrainedness gives
+    a better measure of difficulty than proximity to a phase transition. We also discuss variable
+    and value ordering heuristics, and their relationship to the expected number of solutions.
 \end{abstract}
@@
 
-Our experiments were performed on systems with Intel Xeon E5-4650 v2 (Q1'14) CPUs and 768GBytes RAM
+Our experiments were performed on systems with Intel Xeon E5-4650 v2 CPUs and 768GBytes RAM
 (although this much RAM was not needed), running Scientific Linux release 6.7.   ?? Software
@@
 
-\section*{Acknowledgements}
-
-?? I suppose Craig could be given an inkling of undeserved gratitude if we have space left.
+% \section*{Acknowledgements}
+% 
+% Thanks to Kitty Meeks. Also, I suppose Craig could be given an inkling of undeserved gratitude if
+% we have space left.
 

### 2016-01-27 paper/paper.tex
Muck around some more

@@
 
-\author{First Author \thanks{This work was supported by a small pile of money, but not as much as we
-    would have liked.} \and Second Author
+\author{First Author \thanks{This work was supported by a small pile of money and a large
+    electricity bill.} \and Second Author
     \and Third Author \\
@@
 
-    \caption{There is a non-induced isomorphism from the pattern graph, in the center, to the target
+    \caption{There is a non-induced isomorphism from the pattern graph, in the centre, to the target
     on the right, mapping vertex 1 to 1, 2 to 3, 3 to 5 and 4 to 6. This is not an induced
@@
 builds upon the phase transition phenomena observed in satisfiability and graph colouring problems
-first described by \citet{Cheeseman:1991} and \citet{Mitchell:1992}---?? cite later stuff and
-explain really hard.
-
-For subgraph isomorphism we identify three relevant control parameters: we can independently alter
-the edge probability of the pattern graph, the edge probability of the target graph, and the
-relative orders (number of vertices) of the pattern and target graphs.  For non-induced
-isomorphisms, with the correct choice of parameters we see results very similar to those observed
-with boolean satisfiability problems: there is a phase transition from satisfiable to unsatisfiable
-(and we can predict its location), and we see a complexity peak occur with three different solvers
-near this phase transition.
+first described by \citet{Cheeseman:1991} and \citet{Mitchell:1992}.  For subgraph isomorphism we
+identify three relevant control parameters: we can independently alter the edge probability of the
+pattern graph, the edge probability of the target graph, and the relative orders (number of
+vertices) of the pattern and target graphs.  For non-induced isomorphisms, with the correct choice
+of parameters we see results very similar to those observed with boolean satisfiability problems:
+there is a phase transition from satisfiable to unsatisfiable (and we can predict its location), and
+we see a complexity peak occur with three different solvers near this phase transition.
 
@@
 really hard to solve. However, each of the three solvers we tested also finds the central
-unsatisfiable region to be hard---this may be unexpected, since this region is not near a
-phase transition. To show that this is not a simple weakness of current subgraph isomorphism
-algorithms, we verify that this region is also hard when using a pseudo-boolean encoding, and
-under reduction to the clique problem. Interestingly, the constrainedness measure
-\citep{Gent:1996:Kappa} \emph{does} predict this difficult region---these instances provide (?? the
-first??) evidence in favour of constrainedness, rather than proximity to a phase transition, being
-an accurate predictor of difficulty.
+unsatisfiable region to be hard---this may be unexpected, since this region is not near a phase
+transition. To show that this is not a simple weakness of current subgraph isomorphism algorithms,
+we verify that this region is also hard when using a pseudo-boolean encoding, and under reduction to
+the clique problem. Interestingly, the constrainedness measure \citep{Gent:1996:Kappa} \emph{does}
+predict this difficult region---these instances provide evidence in favour of constrainedness,
+rather than proximity to a phase transition, being an accurate predictor of difficulty.
 
@@
 Our experiments were performed on systems with Intel Xeon E5-4650 v2 CPUs and 768GBytes RAM
-(although this much RAM was not needed), running Scientific Linux release 6.7.   ?? Software
-versions: Clasp 3.1.3, Gurobi 6.0.5, Glasgow, LAD version 2, VFLib. Software was compiled using GCC
-4.9.
+(although this much RAM was not needed), running Scientific Linux release 6.7. We selected three
+subgraph isomorphism solvers: the Glasgow solver \citep{McCreesh:2015}, LAD \citep{Solnon:2010}, and
+VF2 \citep{Cordella:2004}; each was compiled using GCC 4.9. These solvers all use backtracking
+search to build up an assignment of target vertices (values) to pattern vertices (variables), with
+different kinds of inference and ordering heuristics.
 
-?? Stuff about runtimes and recursive calls. We are not aiming to compare solvers; rather, we are
-looking for solver-independent behaviour. We need a quick note on the models used here too.
-
-?? Timeouts
+In each case we measure the number of recursive calls (guessed assignments) made, not runtimes. We
+are not aiming to compare absolute performance between solvers; rather, we are looking for
+solver-independent patterns of difficulty. We used a timeout of 1,000 seconds, which was enough for
+the Glasgow solver to solve nearly all our instances (whose orders were selected with this timeout
+in mind), although we may slightly overestimate the proportion of unsatisfiable instances for
+extremely sparse or dense pattern graphs. The LAD and VF2 solvers experienced many more failures
+with this timeout, so our picture of just how hard the hardest instances are with these solvers is
+less detailed.
 
@@
 always find a match) to entirely unsatisfiable instances (a maximum clique in this order and edge
-probability of target graph will usually have between 9 and 12 vertices). Indeed,
-\cref{figure:phase-transition} shows that this is the case: the line (and the points show a subset
-of the samples) ?? describe this a bit more, find the 0.5 SAT mark.
+probability of target graph will usually have between 9 and 12 vertices). The dark line in
+\cref{figure:phase-transition} shows that this is the case. For densities of 0.67 or greater, no
+instance is satisfiable; with densities of 0.44 or less, every instance is satisfiable; and with a
+density of 0.55, roughly half the instances are satisfiable.
 
-?? Some stuff about the complexity peak in the Glasgow solver. This looks remarkably similar to random
+The light line plots mean search effort using the Glasgow solver: for sparse patterns, the problem
+is trivial, for dense patterns proving unsatisfiability is not particularly difficult, and we see a
+complexity peak around the point where half the instances are satisfiable.  We also plot the search
+cost of individual instances, as points. The behaviour we observe looks remarkably similar to random
 3SAT problems---compare, for example, Figure 1 of \citet{LeytonBrown:2014}. In particular,
@@
 there are exceptionally hard satisfiable instances \citep{Smith:1997}.
-
-?? Are unsatisfiable instances on the phase transition much harder than satisfiable ones?
 
@@
 steps of 0.05; for each of these 2601 points, we generate ten random instances. The colour denotes
-(?? wrong word?) the proportion of these instances which were found to be satisfiable.  Inside the
-orange region, at the bottom right of each plot, every instance is unsatisfiable---here we are
-trying to find a dense pattern in a sparse target. In the purple region, at the top left, every
-instance is satisfiable---we are looking for a sparse pattern in a dense target (which is easy,
-since we only have to preserve adjacency, not non-adjacency). The white band between the regions
-shows the location of the phase transition: here, roughly half the instances are satisfiable. (We
-discuss the black line below.)
+the proportion of these instances which were found to be satisfiable.  Inside the orange region, at
+the bottom right of each plot, every instance is unsatisfiable---here we are trying to find a dense
+pattern in a sparse target. In the purple region, at the top left, every instance is
+satisfiable---we are looking for a sparse pattern in a dense target (which is easy, since we only
+have to preserve adjacency, not non-adjacency). The white band between the regions shows the
+location of the phase transition: here, roughly half the instances are satisfiable. (We discuss the
+black line below.)
 
@@
 
-To predict the location of the phase transition, we can calculate the expected number of solutions
-by making a few dubious assumptions involving rounding and independence. Since we are trying to find
-an \emph{injective} mapping from a pattern $P = G(p, d_p)$ to a target $T = G(t, d_t)$, there are
-$t^{\underline{p}} = t \cdot (t - 1) \cdot \ldots \cdot (t - p + 1)$ possible assignments of target
-vertices to pattern vertices.  We expect the pattern to have $d_p \cdot \binom{p}{2}$ edges, so we
-obtain the probability of all of these edges being mapped to edges in the target by raising $d_t$ to
-this power, giving an expected number of solutions of \[ \langle Sol \rangle = t^{\underline{p}}
-\cdot {d_t}^{d_p \cdot \binom{p}{2}} \textnormal{.} \] This formula predicts a very sharp phase
-transition from $\langle Sol \rangle \ll 1$ to $\langle Sol \rangle \gg 1$. We plot where this
-occurs using black lines in the first row of \cref{figure:non-induced}.
+We can approximately predict the location of the phase transition by calculating (with
+simplifications regarding rounding and independence) the expected number of solutions for given
+parameters. Since we are trying to find an \emph{injective} mapping from a pattern $P = G(p, d_p)$
+to a target $T = G(t, d_t)$, there are $t^{\underline{p}} = t \cdot (t - 1) \cdot \ldots \cdot (t -
+p + 1)$ possible assignments of target vertices to pattern vertices.  We expect the pattern to have
+$d_p \cdot \binom{p}{2}$ edges, so we obtain the probability of all of these edges being mapped to
+edges in the target by raising $d_t$ to this power, giving an expected number of solutions of \[
+\langle Sol \rangle = t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \textnormal{.} \] This
+formula predicts a very sharp phase transition from $\langle Sol \rangle \ll 1$ to $\langle Sol
+\rangle \gg 1$. We plot where this occurs using black lines in the first row of
+\cref{figure:non-induced}.
 
-We can see that this prediction is generally accurate, except that for very low and very high
-pattern densities, we overestimate the satisfiable region slightly. This is because although an
-expected number of solutions much below one implies a high likelihood of unsatisfiability, it is not
-true that a high expected number of solutions implies that any particular instance is likely to be
+This prediction is generally reasonably accurate, except that for very low and very high pattern
+densities, we overestimate the satisfiable region. This is due to variance: although an expected
+number of solutions much below one implies a high likelihood of unsatisfiability, it is not true
+that a high expected number of solutions implies that any particular instance is likely to be
 satisfiable---a similar behaviour is seen with random constraint satisfaction problems
 \citep{Smith:1994,Smith:1996}.
-
-?? Also because of half edges.
-
-?? Note that the size of the state space is defined differently to in binary CSPs: we're using the
-all diff constraint to determine the size of the state space, and are not treating it as a
-constraint. In other words, our generate-and-test approach is alldiff-aware. This matters for kappa.
 
@@
 two or more domains are of equal size, LAD simply breaks ties lexicographically. Glasgow, however, will
-pick a variable corresponding to a pattern vertex of highest degree. This strategy was determined
-empirically, but could have been derived from the $\langle Sol \rangle$ formula: picking a
-pattern vertex of high degree will make the remaining pattern subgraph sparser, which will decrease
-the exponent in the second half of the formula, maximising the overall value. LAD does not apply a
-value ordering heuristic, but Glasgow does: it prefers target vertices of lowest degree. Again, this
-was determined empirically, but it has the effect of increasing $\langle Sol \rangle$ by increasing
-the remaining target density.
+pick a variable corresponding to a pattern vertex of highest degree. This strategy was apparently
+determined empirically, but could have been derived from the $\langle Sol \rangle$ formula: picking
+a pattern vertex of high degree will make the remaining pattern subgraph sparser, which will
+decrease the exponent in the second half of the formula, maximising the overall value. LAD does not
+apply a value ordering heuristic, but Glasgow does: it prefers target vertices of lowest degree.
+Again, this appears to have been determined empirically, but it has the effect of increasing
+$\langle Sol \rangle$ by increasing the remaining target density.
 
@@
 
-In \cref{figure:induced} we repeat our experiments, finding induced isomorphisms. With a pattern of
-order 10, we get two independent phase transitions (?? describe properly). For each of the solvers,
-the satisfiable instances before either phase transition are generally easy, instances near the
-phase transition are hard, and the large central unsatisfiable region is again easy.
+In the first four rows of \cref{figure:induced} we repeat our experiments, finding induced
+isomorphisms. With a pattern of order 10, we get two independent phase transitions: the bottom right
+half of the plots resemble the non-induced results, and the top left half is close to a mirror
+image. The central satisfiable region, which is away from either phase transition, is
+computationally easy, but instances near the phase transition are hard.
 
-?? For larger patterns of order 20 and 30, we have large unsatisfiable regions in the middle,
-but these remain hard.
+For larger patterns of order 20 and 30, we have a large unsatisfiable region in the middle. Despite
+not being near either phase transition, instances in the centre remain computationally challenging.
+We also plot patterns of orders 14, 15 and 16, to show the transition between the two behaviours.
 
-Pattern orders 14, 15 and 16 show the transition between the two styles.
+\subsection{Predictions and Heuristics}
 
-\subsection{Predicting Induced Behaviour}
+For small patterns, simply considering the mirror image of the non-induced prediction provides a
+reasonable approximation of the location of the phase transition, but as we show on the fifth row of
+the figure, this breaks down for larger patterns. For a better prediction, we repeat the argument
+for locating the non-induced phase transition and additionally considering non-edges, to get an
+expected number of solutions of \[ \langle Sol \rangle = t^{\underline{p}} \cdot {d_t}^{d_p \cdot
+\binom{p}{2}} \cdot {(1 - d_{t})}^{(1 - d_{p}) \cdot \binom{p}{2}} \textnormal{.} \] We plot
+this using black lines on \cref{figure:induced}---again, our prediction is accurate except for
+very sparse or very dense patterns.
 
-By repeating the argument for locating the non-induced phase transition and additionally considering
-non-edges, we get an expected number of solutions of \[ \langle Sol \rangle = t^{\underline{p}}
-    \cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot {(1 - d_{t})}^{(1 - d_{p}) \cdot \binom{p}{2}}
-\textnormal{.} \] We plot this
-using black lines on \cref{figure:induced}---again, our prediction is accurate except for very
-sparse or very dense patterns.
+We might guess that degree-based heuristics would just not work for the induced problem: for any
+claim about the degree, the opposite will hold for the complement constraints. However, empirically,
+this is not the case: on the final row of \cref{figure:induced}, we show whether it is better to use
+the original pattern and target as the input to the Glasgow algorithm, or to take the complements.
+(The only steps performed by the Glagsow algorithm which differs under taking the complements are
+the degree-based heuristics.  LAD and VF2 are not symmetric in this way: LAD performs a filtering
+step using degree information, but does not consider the complement degree, and VF2 uses
+connectivity in the pattern graph.)
+
+?? Explain bottom row of \cref{figure:induced}, and give way of calculating it. The key point is
+that depending upon where we are, changes to one of the two terms has more effect in the equation.
+Note how sharp and well-defined this is.
+
+?? Note that when the pattern is small, it is always best to try to move towards the satisfiable
+region, even for unsatisfiable instances. It's not clear what's going on when the pattern is large.
+Note the suggestion by \citet{Walsh:1998} that switching heuristics based upon an estimate of the
+solubility of the problem may offer good performance.
+
+?? Is this heuristic better in general, on non-random instances? Unfortunately most of the instances
+in the standard benchmark set are too sparse for it to kick in. However, it provides a justification
+for reusing the non-induced heuristics.
 
@@
 
+?? Note that the size of the state space is defined differently to in binary CSPs: we're using the
+all diff constraint to determine the size of the state space, and are not treating it as a
+constraint. In other words, our generate-and-test approach is alldiff-aware. This matters for kappa.
+
 ?? Maybe note a few solver-dependent hard instances.  There's odd unsat with LAD at the top right
@@
 can see immediately that no mapping is possible, but LAD doesn't do degree filtering on complements.
-
-\subsection{New Heuristics}
-
-?? Explain why we might guess degree heuristics just won't work for induced: for everything we say
-about degree, the opposite holds too for the complement constraints.
-
-?? Explain bottom row of \cref{figure:induced}, and give way of calculating it. The key point is
-that depending upon where we are, changes to one of the two terms has more effect in the equation.
-Note how sharp and well-defined this is.
-
-?? Possibly send the data to Lars, and get him to run his portfolios stuff on it, and then make fun
-of machine learning?
-
-?? Note that when the pattern is small, it is always best to try to move towards the satisfiable
-region, even for unsatisfiable instances. It's not clear what's going on when the pattern is large.
-Note the suggestion by \citet{Walsh:1998} that switching heuristics based upon an estimate of the
-solubility of the problem may offer good performance.
-
-?? Is this heuristic better in general, on non-random instances? Unfortunately most of the instances
-in the standard benchmark set are too sparse for it to kick in. However, it provides a justification
-for reusing the non-induced heuristics.
-
-\section{Other Encodings and Solvers}
 
@@
 mapped to adjacent vertices, and that non-adjacent vertices are mapped to non-adjacent vertices. We
-used the Clasp solver~\citep{gekakaosscsc11a} to solve the pseudo-boolean instances. The instances
-that are hard for the Glasgow solver remain hard for the PB solver, including instances inside the
-central region, and the easy satisfiable instances remain easy.
+used the Clasp solver~\citep{gekakaosscsc11a} version 3.1.3 to solve the pseudo-boolean instances.
+The instances that are hard for the Glasgow solver remain hard for the PB solver, including
+instances inside the central region, and the easy satisfiable instances remain easy.
 

### 2016-01-31 paper/paper.tex
Title consistency
More stuff

@@
 to successful application in areas such as computer
-vision \citep{Damiand:2011,Solnon:2015}, biochemistry \citep{Giugno:2013}, and pattern recognition
-\citep{Conte:2004}.
+vision \citep{Damiand:2011,Solnon:2015}, biochemistry \citep{Giugno:2013,Carletti:2015}, and pattern
+recognition \citep{Conte:2004}.
 
@@
 computationally fairly easy, with most of the difficulty being in dealing with the size of the
-model.  This problem cannot be addressed by using a pattern graph from one of the random suites with
-the ``wrong'' target graph: this tends to give either a trivially unsatisfiable instance, or a
+model.  ?? \citet{Battiti:2007}
+
+This problem cannot be addressed by using a pattern graph from one of the random suites with the
+``wrong'' target graph: this tends to give either a trivially unsatisfiable instance, or a
 satisfiable instance. (In particular, it is \emph{not} the case that a relatively small random graph
@@
 
-\section{Induced Isomorphisms}
+\section{Induced Subgraph Isomorphisms}
 
@@
     or under-constrained. The final row shows when the Glasgow algorithm performs better when given
-    the complements of the pattern and target graphs as
-inputs.}\label{figure:induced}
+    the complements of the pattern and target graphs as inputs---the solid lines show the location
+of the phase transition, and the dotted lines are $t_d=0.5$ and the $p_d=t_d$ diagonal.}\label{figure:induced}
 \end{figure*}
@@
 
-For small patterns, simply considering the mirror image of the non-induced prediction provides a
-reasonable approximation of the location of the phase transition, but as we show on the fifth row of
-the figure, this breaks down for larger patterns. For a better prediction, we repeat the argument
-for locating the non-induced phase transition and additionally considering non-edges, to get an
-expected number of solutions of \[ \langle Sol \rangle = t^{\underline{p}} \cdot {d_t}^{d_p \cdot
-\binom{p}{2}} \cdot {(1 - d_{t})}^{(1 - d_{p}) \cdot \binom{p}{2}} \textnormal{.} \] We plot
-this using black lines on \cref{figure:induced}---again, our prediction is accurate except for
-very sparse or very dense patterns.
+For small patterns, simply considering the $p_d=t_d$ diagonal mirror image of the non-induced
+prediction provides a reasonable approximation of the location of the phase transition (recall the
+complement property), but as we show on the fifth row of the figure, this breaks down for larger
+patterns. For a better prediction, we repeat the argument for locating the non-induced phase
+transition and additionally considering non-edges, to get an expected number of solutions of \[
+\langle Sol \rangle = t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot {(1 -
+d_{t})}^{(1 - d_{p}) \cdot \binom{p}{2}} \textnormal{.} \] We plot this using black lines on the top
+row of \cref{figure:induced}---again, our prediction is accurate except for very sparse or very
+dense patterns.
 
@@
 
-?? Explain bottom row of \cref{figure:induced}, and give way of calculating it. The key point is
-that depending upon where we are, changes to one of the two terms has more effect in the equation.
-Note how sharp and well-defined this is.
+For patterns of order 10, it is always better to try to move towards the satisfiable region: if we
+are in the bottom right diagonal half, we are best retaining the original heuristics (which move us
+towards the top right), and if we are in the top left we should use the complement instead. This
+goes against a suggestion by \citet{Walsh:1998} that switching heuristics based upon an estimate of
+the solubility of the problem may offer good performance.
 
-?? Note that when the pattern is small, it is always best to try to move towards the satisfiable
-region, even for unsatisfiable instances. It's not clear what's going on when the pattern is large.
-Note the suggestion by \citet{Walsh:1998} that switching heuristics based upon an estimate of the
-solubility of the problem may offer good performance.
+For larger patterns, more complex behaviour emerges. If we are the intersection of the bottom half
+and the bottom right diagonal of the search space, we should always retain the original heuristic,
+and if we are in the intersection of the top half and the top left diagonal, we should always use
+the complements. This behaviour can be predicted by taking the partial derivatives of $\langle Sol
+\rangle$ in the $-p_d$ and $t_d$ directions.  However, when inside the remaining two eights of the
+parameter space, the partial derivatives of $\langle Sol \rangle$ disagree on which heuristic to
+use, and using directional derivatives is not enough to resolve the problem. A close observation of
+the data suggests that the actual location of the phase transition may be involved (and perhaps
+\citeauthor{Walsh:1998}'s suggestion applies only in these conditions). In any case, $\langle Sol
+\rangle$ is insufficient to explain the observed behaviour.
 
-?? Is this heuristic better in general, on non-random instances? Unfortunately most of the instances
-in the standard benchmark set are too sparse for it to kick in. However, it provides a justification
-for reusing the non-induced heuristics.
+In practice, this is unlikely to be a problem: most real-world instances are extremely sparse and
+are usually easy, which perhaps explains the continuing popularity of VF2's connectivity-based
+heuristics \citep{Carletti:2015}. In this situation, these experiments justify re-using the
+non-induced heuristics on induced problems.
 
@@
 
-?? Explain the flip bound on row 5, and note that the difficult region is roughly near the phase
-transitions in the original graphs.
-
-?? Note that the large unexpectedly hard region in the middle is solver-independent. Perhaps
-subgraph isomorphism algorithms are not able to make use of adjacency and non-adjacency
-simultaneously, and this is a weakness in how all three are designed? In the following section we
-debunk this wonderful hypothesis.
-
-?? Constrainedness predicts the hard region in the middle \citep{Gent:1996:Kappa}! Add a two line
-description of kappa, the slightly modified way we calculated it. We plot this on the sixth row of
-\cref{figure:induced}.
-
-?? Note that the size of the state space is defined differently to in binary CSPs: we're using the
-all diff constraint to determine the size of the state space, and are not treating it as a
-constraint. In other words, our generate-and-test approach is alldiff-aware. This matters for kappa.
-
-?? Maybe note a few solver-dependent hard instances.  There's odd unsat with LAD at the top right
-(not EHPs), trivial for Glasgow: mapping a non-clique into a clique. Glasgow looks at non-edges and
-can see immediately that no mapping is possible, but LAD doesn't do degree filtering on complements.
+The region in the parameter space where both pattern and target have medium density is far from a
+phase transition, but nevertheless contains instances that are hard for all three of our subgraph
+isomorphism solvers. We would like to know whether this is due to a weakness in current solvers
+(perhaps our solvers cannot region about adjacency and non-adjacency simultaneously?), or whether
+instances in this region are inherently difficult to solve.  Thus we repeat the induced experiments
+on smaller pattern and target graphs, using different solving techniques.  Although these techniques
+are not competitive in absolute terms, we wish to see if the same pattern of behaviour occurs. The
+results are plotted in \cref{figure:alt}.
 
@@
 
-The region in the parameter space where both pattern and target have medium density is far from a
-phase transition, but nevertheless contains instances that are hard for current subgraph isomorphism
-solvers. We would like to know whether this is due to a weakness in current solvers, or whether
-instances in this region are inherently difficult to solve, as constrainedness suggests. In
-\cref{figure:alt} we repeat the induced experiments on smaller pattern and target graphs, using
-different solving techniques.  Although these techniques are not competitive in absolute terms, we
-wish to see if the same pattern of behaviour occurs.
-
-We create a pseudo-boolean (PB) encoding as follows. For each pattern vertex $v$ and each target
+The pseudo-boolean (PB) encoding is as follows. For each pattern vertex $v$ and each target
 vertex $w$, we have a binary variable which takes the value 1 if and only if that $v$ is mapped to
@@
 
-We implemented an equivalent SAT encoding with direct encoding of the cardinality constraints. Using
-the Glucose solver, we saw very similar performance to the PB encoding. We also implemented a
+We also implemented an equivalent SAT encoding with direct encoding of the cardinality constraints.
+Using the Glucose solver, we saw very similar performance to the PB encoding. We also implemented a
 similar integer program encoding; the Gurobi solver was only able to solve some of the trivial
@@
 These results suggest that the central region may be genuinely hard, despite not being near a phase
-transition, and that constrainedness is the better measure of difficulty. The clique encoding in
-particular rules out the hypothesis that subgraph isomorphism solvers only find this region hard due
-to not reasoning simultaneously about adjacency and non-adjacency, since the association graph
-encoding constraints consider compatibility rather than adjacency and non-adjacency.
+transition. The clique encoding in particular rules out the hypothesis that subgraph isomorphism
+solvers only find this region hard due to not reasoning simultaneously about adjacency and
+non-adjacency, since the association graph encoding constraints consider compatibility rather than
+adjacency and non-adjacency.
 
-?? Lends weight to kappa
+Constrainedness, denoted $\kappa$, is an alternative measure of difficulty designed to generalise
+hardness parameters across different combinatorial problems \citep{Gent:1996:Kappa}. A problem with
+$\kappa < 1$ is said to be underconstrained, and is likely to be satisfiable; a problem with $\kappa
+> 1$ is overconstrained, and is likely to be unsatisfiable. Empirically, problems with $\kappa$
+close to 1 are hard, and problems where $\kappa$ is very small or very large are usually easy. By
+handling injectivity as a restriction on the size of the state space rather than as a constraint, we
+derive \[ \kappa = 1 - \frac{\log t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot {(1 -
+d_{t})}^{(1 - d_{p}) \cdot \binom{p}{2}}}{\log t^{\underline{p}}}\] for induced isomorphisms, which
+we plot on the sixth row of \cref{figure:induced}. We see that constrainedness predicts that the
+central region will still be relatively difficult for larger pattern graphs: although the problem is
+overconstrained, it is less overconstrained than in the regions the Glasgow and LAD solvers found easy.
+Thus it seems that rather than just being a unification of existing generalised heuristic
+techniques, constrainedness also gives a better predictor of difficulty than proximity to a phase
+transition.  Unfortunately, constrainedness does not help us with heuristics: minimising
+constrainedness gives the same predictions as maximising the expected number of solutions.
 
@@
 
-?? \citep{Lipets:2009} inexact solutions?
-
 % \section*{Acknowledgements}

### 2016-01-31 prediction/induced-line-actual.rb

@@ -1,0 +1,98 @@
+#!/usr/bin/ruby
+# vim: set sw=4 sts=4 et tw=80 :
+
+require 'bigdecimal'
+
+ps = ARGV[0].to_i
+ts = ARGV[1].to_i
+flip = (ARGV[2] == "true")
+
+measured = []
+
+File.open(ARGV[3], "r") do | f |
+    f.each_line do | line |
+       measured << line.split
+    end
+end
+
+0.step(500, 10) do | x |
+    pdresult = [0, 0]
+    tdresult = [0, 0]
+    pdround = [0, 0]
+    tdround = [0, 0]
+    expsols = [0, 0]
+    pd, td = 0, 0
+
+    [0, 1].each do | pass |
+
+        if flip then
+            if 0 == pass
+                lower = 250
+                upper = 500
+            else
+                lower = 0
+                upper = 250
+            end
+        else
+            if 0 == pass then
+                upper = x + 10
+                lower = 0
+            else
+                upper = 500
+                lower = x + 10
+            end
+        end
+
+        0.upto(100) do
+            if flip then
+                td = BigDecimal.new(x) / 500;
+                pd = BigDecimal.new(upper + lower) / 1000;
+            else
+                pd = BigDecimal.new(x) / 500;
+                td = BigDecimal.new(upper + lower) / 1000;
+            end
+
+            cliqueedges = BigDecimal.new(ps * (ps - 1) / 2);
+            nedges = pd * cliqueedges;
+            nnonedges = (1 - pd) * cliqueedges;
+
+            edgesokprob = td ** nedges;
+            nonedgesokprob = (1 - td) ** nnonedges;
+
+            solprob = edgesokprob * nonedgesokprob;
+
+            nstates = BigDecimal.new(1);
+            1.upto(ps) do | n |
+                nstates *= (ts - n + 1);
+            end
+
+            expsols[pass] = nstates * solprob;
+
+            if ((expsols[pass] < 1) ^ (pass == 1))
+                lower = (upper + lower) / 2
+            elsif ((expsols[pass] > 1) ^ (pass == 1))
+                upper = (upper + lower) / 2
+            else
+                break
+            end
+        end
+
+        if flip then
+            pdresult[pass] = (upper + lower) / 1000.0
+            tdresult[pass] = td
+        else
+            pdresult[pass] = pd
[…]

### 2016-02-01 paper/paper.tex
Space, etc

@@
 requires that the mapping preserve non-adjacency, so there are no ``extra edges'' in the copy of the
-pattern that we find. We illustrate both variants in \cref{figure:sip}.
+pattern that we find. % We illustrate both variants in \cref{figure:sip}.
 Despite these problems being \NP-complete, modern practical subgraph isomorphism algorithms can
@@
 
-\begin{figure}[b]
-    \centering
-    \begin{tikzpicture}[scale=0.5]
-        \begin{scope}
-            \node [draw, circle, fill=uofgcobalt, inner sep=1.5pt] (Na) at (1,  0) {1};
-            \node [draw, circle, fill=uofgpumpkin, inner sep=1.5pt] (Nb) at (1, -1.5) {2};
-            \node [draw, circle, fill=uofglawn, inner sep=1.5pt] (Nc) at (0, -3) {3};
-            \node [draw, circle, fill=uofgthistle, inner sep=1.5pt] (Nd) at (2, -3) {4};
-
-            \draw [ultra thick] (Na) -- (Nb);
-            \draw [ultra thick] (Nb) -- (Nc);
-            \draw [ultra thick] (Nc) -- (Nd);
-            \draw [ultra thick] (Nb) -- (Nd);
-
-            \node [draw, circle, fill=uofgcobalt, inner sep=1.5pt] (N1) at (5.5,  0) {1};
-            \node [draw, circle, fill=white, inner sep=1.5pt] (N2) at (7.5,  0) {2};
-            \node [draw, circle, fill=uofgpumpkin, inner sep=1.5pt] (N3) at (5.5, -1.5) {3};
-            \node [draw, circle, fill=white, inner sep=1.5pt] (N4) at (7.5, -1.5) {4};
-            \node [draw, circle, fill=uofglawn, inner sep=1.5pt] (N5) at (5.5, -3) {5};
-            \node [draw, circle, fill=uofgthistle, inner sep=1.5pt] (N6) at (7.5, -3) {6};
-
-            \draw [thick, color=uofgsandstone!50] (N1) -- (N2);
-            \draw [ultra thick] (N1) -- (N3);
-            \draw [thick, color=uofgsandstone!50] (N1) -- (N4);
-            \draw [thick, color=uofgsandstone!50] (N2) -- (N4);
-            \draw [ultra thick] (N3) -- (N5);
-            \draw [ultra thick] (N3) -- (N6);
-            \draw [thick, color=uofgsandstone!50] (N4) -- (N6);
-            \draw [ultra thick] (N5) -- (N6);
-            \draw [thick, color=uofgsandstone!50] (N2) to [in=45, out=315] (N6);
-            \draw [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);
-
-            \node [draw, circle, fill=white, inner sep=1.5pt] (M1) at (-5.5,  0) {1};
-            \node [draw, circle, fill=uofgcobalt, inner sep=1.5pt] (M2) at (-3.5,  0) {2};
-            \node [draw, circle, fill=uofgthistle, inner sep=1.5pt] (M3) at (-5.5, -1.5) {3};
-            \node [draw, circle, fill=white, inner sep=1.5pt] (M4) at (-3.5, -1.5) {4};
-            \node [draw, circle, fill=uofglawn, inner sep=1.5pt] (M5) at (-5.5, -3) {5};
-            \node [draw, circle, fill=uofgpumpkin, inner sep=1.5pt] (M6) at (-3.5, -3) {6};
-
-            \draw [thick, color=uofgsandstone!50] (M1) -- (M2);
-            \draw [thick, color=uofgsandstone!50] (M1) -- (M3);
-            \draw [thick, color=uofgsandstone!50] (M1) -- (M4);
-            \draw [thick, color=uofgsandstone!50] (M2) -- (M4);
-            \draw [ultra thick] (M3) -- (M5);
-            \draw [ultra thick] (M3) -- (M6);
-            \draw [thick, color=uofgsandstone!50] (M4) -- (M6);
-            \draw [ultra thick] (M5) -- (M6);
-            \draw [ultra thick] (M2) to [in=45, out=315] (M6);
-            \draw [thick, color=uofgsandstone!50] (M1) to [in=135, out=225] (M5);
-
-            \node [anchor=center, font=\Large] (A1) at (-1.5, -1.5) { $\hookleftarrow$ };
-            \node [anchor=center, font=\Large] (A2) at ( 3.5, -1.5) { $\rightarrowtail$ };
-        \end{scope}
-    \end{tikzpicture}
-
-    \caption{There is a non-induced isomorphism from the pattern graph, in the centre, to the target
-    on the right, mapping vertex 1 to 1, 2 to 3, 3 to 5 and 4 to 6. This is not an induced
-    isomorphism, since there is an edge between 1 and 5 in the target but not between 1 and 3 in the
-    pattern. The mapping from the pattern to the (same) target on the left, sending 1 to 2, 2 to 6,
-    3 to 5 and 4 to 3, is both non-induced and induced.}
-    \label{figure:sip}
-\end{figure}
+%%% \begin{figure}[b]
+%%%     \centering
+%%%     \begin{tikzpicture}[scale=0.5]
+%%%         \begin{scope}
+%%%             \node [draw, circle, fill=uofgcobalt, inner sep=1.5pt] (Na) at (1,  0) {1};
+%%%             \node [draw, circle, fill=uofgpumpkin, inner sep=1.5pt] (Nb) at (1, -1.5) {2};
+%%%             \node [draw, circle, fill=uofglawn, inner sep=1.5pt] (Nc) at (0, -3) {3};
+%%%             \node [draw, circle, fill=uofgthistle, inner sep=1.5pt] (Nd) at (2, -3) {4};
+%%% 
+%%%             \draw [ultra thick] (Na) -- (Nb);
+%%%             \draw [ultra thick] (Nb) -- (Nc);
+%%%             \draw [ultra thick] (Nc) -- (Nd);
+%%%             \draw [ultra thick] (Nb) -- (Nd);
+%%% 
+%%%             \node [draw, circle, fill=uofgcobalt, inner sep=1.5pt] (N1) at (5.5,  0) {1};
+%%%             \node [draw, circle, fill=white, inner sep=1.5pt] (N2) at (7.5,  0) {2};
+%%%             \node [draw, circle, fill=uofgpumpkin, inner sep=1.5pt] (N3) at (5.5, -1.5) {3};
+%%%             \node [draw, circle, fill=white, inner sep=1.5pt] (N4) at (7.5, -1.5) {4};
+%%%             \node [draw, circle, fill=uofglawn, inner sep=1.5pt] (N5) at (5.5, -3) {5};
+%%%             \node [draw, circle, fill=uofgthistle, inner sep=1.5pt] (N6) at (7.5, -3) {6};
+%%% 
+%%%             \draw [thick, color=uofgsandstone!50] (N1) -- (N2);
+%%%             \draw [ultra thick] (N1) -- (N3);
+%%%             \draw [thick, color=uofgsandstone!50] (N1) -- (N4);
+%%%             \draw [thick, color=uofgsandstone!50] (N2) -- (N4);
+%%%             \draw [ultra thick] (N3) -- (N5);
+%%%             \draw [ultra thick] (N3) -- (N6);
+%%%             \draw [thick, color=uofgsandstone!50] (N4) -- (N6);
+%%%             \draw [ultra thick] (N5) -- (N6);
+%%%             \draw [thick, color=uofgsandstone!50] (N2) to [in=45, out=315] (N6);
+%%%             \draw [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);
+%%% 
+%%%             \node [draw, circle, fill=white, inner sep=1.5pt] (M1) at (-5.5,  0) {1};
+%%%             \node [draw, circle, fill=uofgcobalt, inner sep=1.5pt] (M2) at (-3.5,  0) {2};
+%%%             \node [draw, circle, fill=uofgthistle, inner sep=1.5pt] (M3) at (-5.5, -1.5) {3};
+%%%             \node [draw, circle, fill=white, inner sep=1.5pt] (M4) at (-3.5, -1.5) {4};
+%%%             \node [draw, circle, fill=uofglawn, inner sep=1.5pt] (M5) at (-5.5, -3) {5};
+%%%             \node [draw, circle, fill=uofgpumpkin, inner sep=1.5pt] (M6) at (-3.5, -3) {6};
+%%% 
+%%%             \draw [thick, color=uofgsandstone!50] (M1) -- (M2);
+%%%             \draw [thick, color=uofgsandstone!50] (M1) -- (M3);
+%%%             \draw [thick, color=uofgsandstone!50] (M1) -- (M4);
+%%%             \draw [thick, color=uofgsandstone!50] (M2) -- (M4);
+%%%             \draw [ultra thick] (M3) -- (M5);
+%%%             \draw [ultra thick] (M3) -- (M6);
+%%%             \draw [thick, color=uofgsandstone!50] (M4) -- (M6);
+%%%             \draw [ultra thick] (M5) -- (M6);
+%%%             \draw [ultra thick] (M2) to [in=45, out=315] (M6);
+%%%             \draw [thick, color=uofgsandstone!50] (M1) to [in=135, out=225] (M5);
+%%% 
+%%%             \node [anchor=center, font=\Large] (A1) at (-1.5, -1.5) { $\hookleftarrow$ };
+%%%             \node [anchor=center, font=\Large] (A2) at ( 3.5, -1.5) { $\rightarrowtail$ };
+%%%         \end{scope}
+%%%     \end{tikzpicture}
+%%% 
+%%%     \caption{There is a non-induced isomorphism from the pattern graph, in the centre, to the target
+%%%     on the right, mapping vertex 1 to 1, 2 to 3, 3 to 5 and 4 to 6. This is not an induced
+%%%     isomorphism, since there is an edge between 1 and 5 in the target but not between 1 and 3 in the
+%%%     pattern. The mapping from the pattern to the (same) target on the left, sending 1 to 2, 2 to 6,
+%%%     3 to 5 and 4 to 3, is both non-induced and induced.}
+%%%     \label{figure:sip}
+%%% \end{figure}
 
@@
 handling injectivity as a restriction on the size of the state space rather than as a constraint, we
-derive \[ \kappa = 1 - \frac{\log t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot {(1 -
-d_{t})}^{(1 - d_{p}) \cdot \binom{p}{2}}}{\log t^{\underline{p}}}\] for induced isomorphisms, which
+derive \[ \kappa = 1 - \frac{\log \left( t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot {(1 -
+d_{t})}^{(1 - d_{p}) \cdot \binom{p}{2}} \right)}{\log t^{\underline{p}}}\] for induced isomorphisms, which
 we plot on the sixth row of \cref{figure:induced}. We see that constrainedness predicts that the

### 2016-02-01 paper/paper.tex
Muck around more

--- 2016-01-31 paper/paper.tex
@@
     the induced variant, much richer behaviour is observed, and constrainedness gives
-    a better measure of difficulty than proximity to a phase transition. We also discuss variable
+    a better measure of difficulty than does proximity to a phase transition. We also discuss variable
     and value ordering heuristics, and their relationship to the expected number of solutions.
@@
 computationally fairly easy, with most of the difficulty being in dealing with the size of the
-model.  ?? \citet{Battiti:2007}
+model. This has lead to bias in algorithm design, to the extent that some proposed techniques will
+\emph{only} work on satisfiable instances \citet{Battiti:2007}.
 
-This problem cannot be addressed by using a pattern graph from one of the random suites with the
-``wrong'' target graph: this tends to give either a trivially unsatisfiable instance, or a
-satisfiable instance. (In particular, it is \emph{not} the case that a relatively small random graph
-is unlikely to appear in a larger random graph.)
+The lack of unsatisfiable instances cannot be addressed by using a pattern graph from one of the
+random suites with the ``wrong'' target graph: this tends to give either a trivially unsatisfiable
+instance, or a satisfiable instance. (In particular, it is \emph{not} the case that a relatively
+small random graph is unlikely to appear in a larger random graph.)
 
@@
 
-\begin{figure*}[p]
+\begin{figure*}[t]
+    \setlength{\abovecaptionskip}{0pt}
     \input{gen-graph-induced.tex}
@@
     of \cref{figure:non-induced}. The second, third and fourth rows show the number of search nodes used by the
-    Glasgow, LAD and VF2 algorithms. The fifth row shows a bound on the satisfiable region, by
-    considering where a \emph{non-}induced isomorphism may also be a non-induced isomorphism between
-    complement graphs. The sixth row plots constrainedness: the darkest region is where
-    constrainedness is exactly one, and the lighter regions show where the problem is either over-
+    Glasgow, LAD and VF2 algorithms. The fifth row plots constrainedness: the darkest region is where
+    $\kappa = 1$, and the lighter regions show where the problem is either over-
     or under-constrained. The final row shows when the Glasgow algorithm performs better when given
@@
 
-For small patterns, simply considering the $p_d=t_d$ diagonal mirror image of the non-induced
-prediction provides a reasonable approximation of the location of the phase transition (recall the
-complement property), but as we show on the fifth row of the figure, this breaks down for larger
-patterns. For a better prediction, we repeat the argument for locating the non-induced phase
-transition and additionally considering non-edges, to get an expected number of solutions of \[
-\langle Sol \rangle = t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot {(1 -
-d_{t})}^{(1 - d_{p}) \cdot \binom{p}{2}} \textnormal{.} \] We plot this using black lines on the top
-row of \cref{figure:induced}---again, our prediction is accurate except for very sparse or very
-dense patterns.
+To predict the location of the induced phase transition, we repeat the argument for locating the
+non-induced phase transition and additionally considering non-edges, to get an expected number of
+solutions of \[ \langle Sol \rangle = t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot
+{(1 - d_{t})}^{(1 - d_{p}) \cdot \binom{p}{2}} \textnormal{.} \] We plot this using black lines on
+the top row of \cref{figure:induced}---again, our prediction is accurate except for very sparse or
+very dense patterns.
 
@@
 handling injectivity as a restriction on the size of the state space rather than as a constraint, we
-derive \[ \kappa = 1 - \frac{\log t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot {(1 -
-d_{t})}^{(1 - d_{p}) \cdot \binom{p}{2}}}{\log t^{\underline{p}}}\] for induced isomorphisms, which
-we plot on the sixth row of \cref{figure:induced}. We see that constrainedness predicts that the
+derive \[ \kappa = 1 - \frac{\log \left( t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot {(1 -
+d_{t})}^{(1 - d_{p}) \cdot \binom{p}{2}} \right)}{\log t^{\underline{p}}}\] for induced isomorphisms, which
+we plot on the fifth row of \cref{figure:induced}. We see that constrainedness predicts that the
 central region will still be relatively difficult for larger pattern graphs: although the problem is

### 2016-02-01 paper/paper.tex
Use less space

@@
 
-The lack of unsatisfiable instances cannot be addressed by using a pattern graph from one of the
-random suites with the ``wrong'' target graph: this tends to give either a trivially unsatisfiable
-instance, or a satisfiable instance. (In particular, it is \emph{not} the case that a relatively
-small random graph is unlikely to appear in a larger random graph.)
+%%% The lack of unsatisfiable instances cannot be addressed by using a pattern graph from one of the
+%%% random suites with the ``wrong'' target graph: this tends to give either a trivially unsatisfiable
+%%% instance, or a satisfiable instance. (In particular, it is \emph{not} the case that a relatively
+%%% small random graph is unlikely to appear in a larger random graph.)
 
@@
 of parameters we see results very similar to those observed with boolean satisfiability problems:
-there is a phase transition from satisfiable to unsatisfiable (and we can predict its location), and
-we see a complexity peak occur with three different solvers near this phase transition.
+there is a phase transition (whose location we can predict) from satisfiable to unsatisfiable,
+we see a solver-independent complexity peak occur near this phase transition, and understanding this
+behaviour helps us to select variable and value ordering heuristics.
 
@@
 really hard to solve. However, each of the three solvers we tested also finds the central
-unsatisfiable region to be hard---this may be unexpected, since this region is not near a phase
-transition. To show that this is not a simple weakness of current subgraph isomorphism algorithms,
-we verify that this region is also hard when using a pseudo-boolean encoding, and under reduction to
-the clique problem. Interestingly, the constrainedness measure \citep{Gent:1996:Kappa} \emph{does}
-predict this difficult region---these instances provide evidence in favour of constrainedness,
-rather than proximity to a phase transition, being an accurate predictor of difficulty.
-
-?? Finally, we discuss the relationship between heuristics and the expected number of solutions,
-recovering existing heuristics for the non-induced variant, and introducing new heuristics for the
-induced variant. I have failed miserably at integrating this paragraph smoothly.
+unsatisfiable region to be hard, despite it not being near a phase transition. To show that this is
+not a simple weakness of current subgraph isomorphism algorithms, we verify that this region is also
+hard when using a pseudo-boolean encoding, and under reduction to the clique problem. Interestingly,
+the constrainedness measure proposed by \citet{Gent:1996:Kappa} \emph{does} predict this difficult
+region---these instances provide evidence in favour of constrainedness, rather than proximity to a
+phase transition, being an accurate predictor of difficulty.
 
@@
 3SAT problems---compare, for example, Figure 1 of \citet{LeytonBrown:2014}. In particular,
-satisfiable instances tend to be easier but show greater variation than unsatisfiable instances, and
-there are exceptionally hard satisfiable instances \citep{Smith:1997}.
+satisfiable instances tend to be easier, but show greater variation than unsatisfiable instances. % , and
+% there are exceptionally hard satisfiable instances \citep{Smith:1997}.
 
@@
 general, satisfiable instances are easy, until very close to the phase transition. As we hit the
-phase transition and move into the unsatisfiable region, we see a complexity increase. Finally, as
+phase transition and move into the unsatisfiable region, we see complexity increase. Finally, as
 we pass through the phase transition and move deeper into the unsatisfiable region, instances become
@@
 satisfiable---a similar behaviour is seen with random constraint satisfaction problems
-\citep{Smith:1994,Smith:1996}.
+\citep{Smith:1996}.
+
+% \citep{Smith:1994}
 
@@
 This is usually done by cheaper surrogates, rather than direct calculation. When branching, both LAD
-and Glasgow will pick a variable with fewest remaining values in its domain: doing this will
-generally reduce the first part of the $\langle Sol \rangle$ equation by as little as possible. When
-two or more domains are of equal size, LAD simply breaks ties lexicographically. Glasgow, however, will
-pick a variable corresponding to a pattern vertex of highest degree. This strategy was apparently
+and Glasgow pick a variable with fewest remaining values in its domain: doing this will generally
+reduce the first part of the $\langle Sol \rangle$ equation by as little as possible. When two or
+more domains are of equal size, LAD simply breaks ties lexicographically, whereas Glasgow will pick
+a variable corresponding to a pattern vertex of highest degree. This strategy was apparently
 determined empirically, but could have been derived from the $\langle Sol \rangle$ formula: picking
@@
 Again, this appears to have been determined empirically, but it has the effect of increasing
-$\langle Sol \rangle$ by increasing the remaining target density.
-
-The VF2 heuristics, in contrast, are based around preserving connectivity. This perhaps explains
-some of VF2's weakness on graphs that are not extremely sparse.
+$\langle Sol \rangle$ by increasing the remaining target density. The VF2 heuristics, in contrast,
+are based around preserving connectivity, which gives very little discrimination except on the
+sparsest of inputs.
 
@@
 
-For larger patterns, more complex behaviour emerges. If we are the intersection of the bottom half
+For larger patterns, more complex behaviour emerges. If we are in the intersection of the bottom half
 and the bottom right diagonal of the search space, we should always retain the original heuristic,
@@
 the complements. This behaviour can be predicted by taking the partial derivatives of $\langle Sol
-\rangle$ in the $-p_d$ and $t_d$ directions.  However, when inside the remaining two eights of the
+\rangle$ in the $-p_d$ and $t_d$ directions.  However, when inside the remaining two eighths of the
 parameter space, the partial derivatives of $\langle Sol \rangle$ disagree on which heuristic to
@@
 The region in the parameter space where both pattern and target have medium density is far from a
-phase transition, but nevertheless contains instances that are hard for all three of our subgraph
-isomorphism solvers. We would like to know whether this is due to a weakness in current solvers
-(perhaps our solvers cannot region about adjacency and non-adjacency simultaneously?), or whether
-instances in this region are inherently difficult to solve.  Thus we repeat the induced experiments
-on smaller pattern and target graphs, using different solving techniques.  Although these techniques
-are not competitive in absolute terms, we wish to see if the same pattern of behaviour occurs. The
-results are plotted in \cref{figure:alt}.
+phase transition, but nevertheless contains instances that are hard for all three solvers. We would
+like to know whether this is due to a weakness in current solvers (perhaps our solvers cannot region
+about adjacency and non-adjacency simultaneously?), or whether instances in this region are
+inherently difficult to solve.  Thus we repeat the induced experiments on smaller pattern and target
+graphs, using different solving techniques.  Although these techniques are not competitive in
+absolute terms, we wish to see if the same pattern of behaviour occurs. The results are plotted in
+\cref{figure:alt}.
 

### 2016-02-02 paper/paper.tex
Can we fit the association picture in too?

@@
         \end{scope}
+    \end{tikzpicture}\\[0.75cm]
+    \begin{tikzpicture}[scale=0.5]
+        \begin{scope}
+            \foreach \n in {1, ..., 6}{
+                \node [draw, circle, fill=uofgcobalt, inner sep=1.5pt] (N1\n) at (\n, 0) {\n};
+            }
+
+            \draw [decorate, decoration={brace, raise=0.2cm}, very thick] (N11.north west) -- (N16.north
+                east) node [midway, above=0.3cm] { $1 \mapsto$ };
+
+            \foreach \n in {1, ..., 6}{
+                \node [draw, circle, fill=uofgpumpkin, inner sep=1.5pt] (N2\n) at (0, -\n) {\n};
+            }
+
+            \draw [decorate, decoration={brace, raise=0.2cm, mirror}, very thick] (N21.north west) --
+                (N26.south west) node [midway, left=0.3cm] { $2 \mapsto$ };
+
+            \foreach \n in {1, ..., 6}{
+                \node [draw, circle, fill=uofglawn, inner sep=1.5pt] (N3\n) at (7, -\n) {\n};
+            }
+
+            \draw [decorate, decoration={brace, raise=0.2cm}, very thick] (N31.north east) -- (N36.south
+                east) node [midway, right=0.3cm] { $3 \mapsto$ };
+
+            \foreach \n in {1, ..., 6}{
+                \node [draw, circle, fill=uofgthistle, inner sep=1.5pt] (N4\n) at (\n, -7) {\n};
+            }
+
+            \draw [decorate, decoration={brace, raise=0.2cm, mirror}, very thick] (N41.south west) --
+                (N46.south east) node [midway, below=0.3cm] { $4 \mapsto$ };
+
+            \begin{scope}[on background layer]
+                \draw [color=uofgsandstone!50] ($(N11)!0.8!(N11.south)$) -- ($(N22)!0.8!(N22.east)$);
+                \draw [color=uofgsandstone!50] ($(N11)!0.8!(N11.south)$) -- ($(N23)!0.8!(N23.east)$);
+                \draw [color=uofgsandstone!50] ($(N11)!0.8!(N11.south)$) -- ($(N24)!0.8!(N24.east)$);
+                \draw [color=uofgsandstone!50] ($(N11)!0.8!(N11.south)$) -- ($(N25)!0.8!(N25.east)$);
+                \draw [color=uofgsandstone!50] ($(N11)!0.8!(N11.south)$) -- ($(N36)!0.8!(N36.west)$);
+                \draw [color=uofgsandstone!50] ($(N11)!0.8!(N11.south)$) -- ($(N46)!0.8!(N46.north)$);
+                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N12)!0.8!(N12.south)$);
+                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N32)!0.8!(N32.west)$);
+                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N42)!0.8!(N42.north)$);
+                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N13)!0.8!(N13.south)$);
+                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N33)!0.8!(N33.west)$);
+                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N43)!0.8!(N43.north)$);
+                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N14)!0.8!(N14.south)$);
+                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N34)!0.8!(N34.west)$);
+                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N44)!0.8!(N44.north)$);
+                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N15)!0.8!(N15.south)$);
+                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N35)!0.8!(N35.west)$);
+                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N45)!0.8!(N45.north)$);
+                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N22)!0.8!(N22.east)$);
+                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N42)!0.8!(N42.north)$);
+                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N23)!0.8!(N23.east)$);
+                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N43)!0.8!(N43.north)$);
+                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N24)!0.8!(N24.east)$);
+                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N44)!0.8!(N44.north)$);
+                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N25)!0.8!(N25.east)$);
+                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N45)!0.8!(N45.north)$);
+                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N16)!0.8!(N16.south)$);
+                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N22)!0.8!(N22.east)$);
+                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N32)!0.8!(N32.west)$);
+                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N23)!0.8!(N23.east)$);
+                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N33)!0.8!(N33.west)$);
+                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N24)!0.8!(N24.east)$);
+                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N34)!0.8!(N34.west)$);
+                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N25)!0.8!(N25.east)$);
+                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N35)!0.8!(N35.west)$);
+                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N16)!0.8!(N16.south)$);
+                \draw [color=uofgsandstone!50] ($(N12)!0.8!(N12.south)$) -- ($(N33)!0.8!(N33.west)$);
+                \draw [color=uofgsandstone!50] ($(N12)!0.8!(N12.south)$) -- ($(N24)!0.8!(N24.east)$);
+                \draw [color=uofgsandstone!50] ($(N12)!0.8!(N12.south)$) -- ($(N45)!0.8!(N45.north)$);
+                \draw [color=uofgsandstone!50] ($(N22)!0.8!(N22.east)$) -- ($(N14)!0.8!(N14.south)$);
+                \draw [color=uofgsandstone!50] ($(N22)!0.8!(N22.east)$) -- ($(N34)!0.8!(N34.west)$);
+                \draw [color=uofgsandstone!50] ($(N22)!0.8!(N22.east)$) -- ($(N44)!0.8!(N44.north)$);
+                \draw [color=uofgsandstone!50] ($(N22)!0.8!(N22.east)$) -- ($(N16)!0.8!(N16.south)$);
+                \draw [color=uofgsandstone!50] ($(N22)!0.8!(N22.east)$) -- ($(N36)!0.8!(N36.west)$);
+                \draw [color=uofgsandstone!50] ($(N22)!0.8!(N22.east)$) -- ($(N46)!0.8!(N46.north)$);
+                \draw [color=uofgsandstone!50] ($(N32)!0.8!(N32.west)$) -- ($(N13)!0.8!(N13.south)$);
+                \draw [color=uofgsandstone!50] ($(N32)!0.8!(N32.west)$) -- ($(N24)!0.8!(N24.east)$);
+                \draw [color=uofgsandstone!50] ($(N32)!0.8!(N32.west)$) -- ($(N44)!0.8!(N44.north)$);
+                \draw [color=uofgsandstone!50] ($(N32)!0.8!(N32.west)$) -- ($(N15)!0.8!(N15.south)$);
+                \draw [color=uofgsandstone!50] ($(N32)!0.8!(N32.west)$) -- ($(N26)!0.8!(N26.east)$);
+                \draw [color=uofgsandstone!50] ($(N32)!0.8!(N32.west)$) -- ($(N46)!0.8!(N46.north)$);
+                \draw [color=uofgsandstone!50] ($(N42)!0.8!(N42.north)$) -- ($(N13)!0.8!(N13.south)$);
+                \draw [color=uofgsandstone!50] ($(N42)!0.8!(N42.north)$) -- ($(N24)!0.8!(N24.east)$);
+                \draw [color=uofgsandstone!50] ($(N42)!0.8!(N42.north)$) -- ($(N34)!0.8!(N34.west)$);
+                \draw [color=uofgsandstone!50] ($(N42)!0.8!(N42.north)$) -- ($(N15)!0.8!(N15.south)$);
+                \draw [color=uofgsandstone!50] ($(N42)!0.8!(N42.north)$) -- ($(N26)!0.8!(N26.east)$);
+                \draw [color=uofgsandstone!50] ($(N42)!0.8!(N42.north)$) -- ($(N36)!0.8!(N36.west)$);
+                \draw [color=uofgsandstone!50] ($(N13)!0.8!(N13.south)$) -- ($(N34)!0.8!(N34.west)$);
+                \draw [color=uofgsandstone!50] ($(N13)!0.8!(N13.south)$) -- ($(N44)!0.8!(N44.north)$);
+                \draw [color=uofgsandstone!50] ($(N13)!0.8!(N13.south)$) -- ($(N25)!0.8!(N25.east)$);
+                \draw [color=uofgsandstone!50] ($(N13)!0.8!(N13.south)$) -- ($(N26)!0.8!(N26.east)$);
+                \draw [color=uofgsandstone!50] ($(N23)!0.8!(N23.east)$) -- ($(N15)!0.8!(N15.south)$);
+                \draw [color=uofgsandstone!50] ($(N23)!0.8!(N23.east)$) -- ($(N35)!0.8!(N35.west)$);
+                \draw [color=uofgsandstone!50] ($(N23)!0.8!(N23.east)$) -- ($(N45)!0.8!(N45.north)$);
+                \draw [color=uofgsandstone!50] ($(N23)!0.8!(N23.east)$) -- ($(N16)!0.8!(N16.south)$);
+                \draw [color=uofgsandstone!50] ($(N23)!0.8!(N23.east)$) -- ($(N36)!0.8!(N36.west)$);
+                \draw [color=uofgsandstone!50] ($(N23)!0.8!(N23.east)$) -- ($(N46)!0.8!(N46.north)$);
+                \draw [color=uofgsandstone!50] ($(N33)!0.8!(N33.west)$) -- ($(N14)!0.8!(N14.south)$);
+                \draw [color=uofgsandstone!50] ($(N33)!0.8!(N33.west)$) -- ($(N25)!0.8!(N25.east)$);
+                \draw [color=uofgsandstone!50] ($(N33)!0.8!(N33.west)$) -- ($(N45)!0.8!(N45.north)$);
+                \draw [color=uofgsandstone!50] ($(N33)!0.8!(N33.west)$) -- ($(N26)!0.8!(N26.east)$);
+                \draw [color=uofgsandstone!50] ($(N33)!0.8!(N33.west)$) -- ($(N46)!0.8!(N46.north)$);
+                \draw [color=uofgsandstone!50] ($(N43)!0.8!(N43.north)$) -- ($(N14)!0.8!(N14.south)$);
+                \draw [color=uofgsandstone!50] ($(N43)!0.8!(N43.north)$) -- ($(N25)!0.8!(N25.east)$);
+                \draw [color=uofgsandstone!50] ($(N43)!0.8!(N43.north)$) -- ($(N36)!0.8!(N36.west)$);
+                \draw [color=uofgsandstone!50] ($(N14)!0.8!(N14.south)$) -- ($(N35)!0.8!(N35.west)$);
+                \draw [color=uofgsandstone!50] ($(N14)!0.8!(N14.south)$) -- ($(N45)!0.8!(N45.north)$);
+                \draw [color=uofgsandstone!50] ($(N14)!0.8!(N14.south)$) -- ($(N26)!0.8!(N26.east)$);
+                \draw [color=uofgsandstone!50] ($(N24)!0.8!(N24.east)$) -- ($(N16)!0.8!(N16.south)$);
+                \draw [color=uofgsandstone!50] ($(N24)!0.8!(N24.east)$) -- ($(N36)!0.8!(N36.west)$);
+                \draw [color=uofgsandstone!50] ($(N24)!0.8!(N24.east)$) -- ($(N46)!0.8!(N46.north)$);
+                \draw [color=uofgsandstone!50] ($(N34)!0.8!(N34.west)$) -- ($(N15)!0.8!(N15.south)$);
+                \draw [color=uofgsandstone!50] ($(N34)!0.8!(N34.west)$) -- ($(N26)!0.8!(N26.east)$);
+                \draw [color=uofgsandstone!50] ($(N34)!0.8!(N34.west)$) -- ($(N46)!0.8!(N46.north)$);
+                \draw [color=uofgsandstone!50] ($(N44)!0.8!(N44.north)$) -- ($(N15)!0.8!(N15.south)$);
+                \draw [color=uofgsandstone!50] ($(N44)!0.8!(N44.north)$) -- ($(N26)!0.8!(N26.east)$);
+                \draw [color=uofgsandstone!50] ($(N44)!0.8!(N44.north)$) -- ($(N36)!0.8!(N36.west)$);
+                \draw [color=uofgsandstone!50] ($(N15)!0.8!(N15.south)$) -- ($(N26)!0.8!(N26.east)$);
+                \draw [color=uofgsandstone!50] ($(N25)!0.8!(N25.east)$) -- ($(N16)!0.8!(N16.south)$);
+                \draw [color=uofgsandstone!50] ($(N25)!0.8!(N25.east)$) -- ($(N36)!0.8!(N36.west)$);
+                \draw [color=uofgsandstone!50] ($(N25)!0.8!(N25.east)$) -- ($(N46)!0.8!(N46.north)$);
+                \draw [color=uofgsandstone!50] ($(N35)!0.8!(N35.west)$) -- ($(N46)!0.8!(N46.north)$);
+                \draw [color=uofgsandstone!50] ($(N45)!0.8!(N45.north)$) -- ($(N26)!0.8!(N26.east)$);
+                \draw [color=uofgsandstone!50] ($(N45)!0.8!(N45.north)$) -- ($(N36)!0.8!(N36.west)$);
+                \draw [ultra thick] ($(N12)!0.8!(N12.south)$) -- ($(N43)!0.8!(N43.north)$);
+                \draw [ultra thick] ($(N12)!0.8!(N12.south)$) -- ($(N35)!0.8!(N35.west)$);
+                \draw [ultra thick] ($(N12)!0.8!(N12.south)$) -- ($(N26)!0.8!(N26.east)$);
+                \draw [ultra thick] ($(N43)!0.8!(N43.north)$) -- ($(N35)!0.8!(N35.west)$);
+                \draw [ultra thick] ($(N43)!0.8!(N43.north)$) -- ($(N26)!0.8!(N26.east)$);
+                \draw [ultra thick] ($(N35)!0.8!(N35.west)$) -- ($(N26)!0.8!(N26.east)$);
+            \end{scope}
+        \end{scope}
     \end{tikzpicture}
 
-    \caption{There is a non-induced isomorphism from the pattern graph, in the centre, to the target
+    \caption{On top, there is a non-induced isomorphism from the pattern graph, in the center, to the target
     on the right, mapping vertex 1 to 1, 2 to 3, 3 to 5 and 4 to 6. This is not an induced
     isomorphism, since there is an edge between 1 and 5 in the target but not between 1 and 3 in the
-    pattern. The mapping from the pattern to the (same) target on the left, sending 1 to 2, 2 to 6,
-    3 to 5 and 4 to 3, is both non-induced and induced.}
+    pattern. The mapping from the pattern to the (same) target on the left, sending 1 to 2, 2 to 6, 3 to
+    5 and 4 to 3, is both non-induced and induced.
+    Below, the association graph encoding for the induced version: the highlighted clique
+    corresponds to the same mapping.}
     \label{figure:sip}
@@
 only if $t_1$ is adjacent to $t_2$. A clique of size equal to the order of the pattern graph exists
-in the association graph if and only if the problem is satisfiable \citep{Levi:1973}.  We used this
-encoding with a version of the BBMC clique algorithm \citep{SanSegundo:2011} modified to solve the
-decision problem. Again, the instances in the central region remain hard, and additionally, some of
-the easy satisfiable instances become hard.
+in the association graph if and only if the problem is satisfiable \citep{Levi:1973}.  We illustrate
+this in \cref{figure:sip}. We used this encoding with a version of the BBMC clique algorithm
+\citep{SanSegundo:2011} modified to solve the decision problem. Again, the instances in the central
+region remain hard, and additionally, some of the easy satisfiable instances become hard.
 

### 2016-02-02 paper/paper.tex
James's changes

@@
 only if
-$v$ and $w$ are not adjacent in $G$. We write $G(v, p)$ for an Erd\H{o}s-R\'enyi random graph with
-$v$ vertices, and an edge between each distinct pair of vertices with independent probability $p$.
+$v$ and $w$ are not adjacent in $G$. We write $G(n, p)$ for an Erd\H{o}s-R\'enyi random graph with
+$n$ vertices, and an edge between each distinct pair of vertices with independent probability $p$.
 
@@
     \caption{With a fixed a pattern graph order of 20, a target graph order of 150, a target edge
-        probability of 0.4, and varying pattern edge probability, we observe a phase transition and
+        probability of 0.40, and varying pattern edge probability, we observe a phase transition and
         complexity peak with the Glasgow solver in the non-induced variant. Each point represents
-        one instance. The lines show mean search effort, and mean proportion satisfiable.}
+        one instance. The lines show mean search effort and mean proportion satisfiable.}
     \label{figure:phase-transition}
@@
         bands shows the phase transitions, and the black lines are our predictions of where the
-        phase transition will occur. On the second row, we show the number of search nodes used by the
-        Glasgow algorithm, on the third row, the number of search nodes used by the LAD algorithm, and the
-        fourth, VF2; the dark regions indicate ``really hard'' instances.}
+        phase transition will occur. On the final three rows, we show the number of search nodes used by the
+        Glasgow, LAD and VF2 solvers; the dark regions indicate ``really hard'' instances.}
     \label{figure:non-induced}
@@
 row of \cref{figure:non-induced} we show the satisfiability phase transition for the non-induced
-variant, for patterns of order 10, 20 and 30, a target of order 150, and varying pattern (x-axis)
+variant, for patterns of order 10, 20 and 30, targets of order 150, and varying pattern (x-axis)
 and target (y-axis) edge probabilities. Each axis runs over 51 edge probabilities, from 0 to 1 in
-steps of 0.05; for each of these 2601 points, we generate ten random instances. The colour denotes
+steps of 0.02; for each of these 2601 points, we generate ten random instances. The colour denotes
 the proportion of these instances which were found to be satisfiable.  Inside the orange region, at
@@
 
-When designing variable and value ordering heuristics for backtracking search algorithms, there are
-various general principles which appear to be effective---one of these is to try to maximise the
-expected number of solutions inside any subproblem considered during search \citep{Gent:1996:EN}.
-This is usually done by cheaper surrogates, rather than direct calculation. When branching, both LAD
-and Glasgow pick a variable with fewest remaining values in its domain: doing this will generally
-reduce the first part of the $\langle Sol \rangle$ equation by as little as possible. When two or
-more domains are of equal size, LAD simply breaks ties lexicographically, whereas Glasgow will pick
-a variable corresponding to a pattern vertex of highest degree. This strategy was apparently
+Various general principles have been considered when designing variable and value ordering
+heuristics for backtracking search algorithms---one of these is to try to maximise the expected
+number of solutions inside any subproblem considered during search \citep{Gent:1996:EN}.  This is
+usually done by cheaper surrogates, rather than direct calculation. When branching, both LAD and
+Glasgow pick a variable with fewest remaining values in its domain: doing this will generally reduce
+the first part of the $\langle Sol \rangle$ equation by as little as possible. When two or more
+domains are of equal size, LAD simply breaks ties lexicographically, whereas Glasgow will pick a
+variable corresponding to a pattern vertex of highest degree. This strategy was apparently
 determined empirically, but could have been derived from the $\langle Sol \rangle$ formula: picking
@@
 are in the bottom right diagonal half, we are best retaining the original heuristics (which move us
-towards the top right), and if we are in the top left we should use the complement instead. This
+towards the top left), and if we are in the top left we should use the complement instead. This
 goes against a suggestion by \citet{Walsh:1998} that switching heuristics based upon an estimate of
@@
 phase transition, but nevertheless contains instances that are hard for all three solvers. We would
-like to know whether this is due to a weakness in current solvers (perhaps our solvers cannot region
+like to know whether this is due to a weakness in current solvers (perhaps our solvers cannot reason
 about adjacency and non-adjacency simultaneously?), or whether instances in this region are
@@
 The pseudo-boolean (PB) encoding is as follows. For each pattern vertex $v$ and each target
-vertex $w$, we have a binary variable which takes the value 1 if and only if that $v$ is mapped to
+vertex $w$, we have a binary variable which takes the value 1 if and only if $v$ is mapped to
 $w$.  Constraints are added to ensure that each pattern vertex maps to exactly one target vertex,

### 2016-02-02 paper/paper.tex
Conclusion and spacing

@@
     \centering
-    \begin{tikzpicture}[scale=0.5]
+    \begin{tikzpicture}[scale=0.45,every node/.style={font=\footnotesize}]
         \begin{scope}
-            \node [draw, circle, fill=uofgcobalt, inner sep=1.5pt] (Na) at (1,  0) {1};
-            \node [draw, circle, fill=uofgpumpkin, inner sep=1.5pt] (Nb) at (1, -1.5) {2};
-            \node [draw, circle, fill=uofglawn, inner sep=1.5pt] (Nc) at (0, -3) {3};
-            \node [draw, circle, fill=uofgthistle, inner sep=1.5pt] (Nd) at (2, -3) {4};
+            \node [draw, circle, fill=uofgcobalt, inner sep=1pt] (Na) at (1,  0) {1};
+            \node [draw, circle, fill=uofgpumpkin, inner sep=1pt] (Nb) at (1, -1.5) {2};
+            \node [draw, circle, fill=uofglawn, inner sep=1pt] (Nc) at (0, -3) {3};
+            \node [draw, circle, fill=uofgthistle, inner sep=1pt] (Nd) at (2, -3) {4};
 
@@
 
-            \node [draw, circle, fill=uofgcobalt, inner sep=1.5pt] (N1) at (5.5,  0) {1};
-            \node [draw, circle, fill=white, inner sep=1.5pt] (N2) at (7.5,  0) {2};
-            \node [draw, circle, fill=uofgpumpkin, inner sep=1.5pt] (N3) at (5.5, -1.5) {3};
-            \node [draw, circle, fill=white, inner sep=1.5pt] (N4) at (7.5, -1.5) {4};
-            \node [draw, circle, fill=uofglawn, inner sep=1.5pt] (N5) at (5.5, -3) {5};
-            \node [draw, circle, fill=uofgthistle, inner sep=1.5pt] (N6) at (7.5, -3) {6};
+            \node [draw, circle, fill=uofgcobalt, inner sep=1pt] (N1) at (5.5,  0) {1};
+            \node [draw, circle, fill=white, inner sep=1pt] (N2) at (7.5,  0) {2};
+            \node [draw, circle, fill=uofgpumpkin, inner sep=1pt] (N3) at (5.5, -1.5) {3};
+            \node [draw, circle, fill=white, inner sep=1pt] (N4) at (7.5, -1.5) {4};
+            \node [draw, circle, fill=uofglawn, inner sep=1pt] (N5) at (5.5, -3) {5};
+            \node [draw, circle, fill=uofgthistle, inner sep=1pt] (N6) at (7.5, -3) {6};
 
@@
 
-            \node [draw, circle, fill=white, inner sep=1.5pt] (M1) at (-5.5,  0) {1};
-            \node [draw, circle, fill=uofgcobalt, inner sep=1.5pt] (M2) at (-3.5,  0) {2};
-            \node [draw, circle, fill=uofgthistle, inner sep=1.5pt] (M3) at (-5.5, -1.5) {3};
-            \node [draw, circle, fill=white, inner sep=1.5pt] (M4) at (-3.5, -1.5) {4};
-            \node [draw, circle, fill=uofglawn, inner sep=1.5pt] (M5) at (-5.5, -3) {5};
-            \node [draw, circle, fill=uofgpumpkin, inner sep=1.5pt] (M6) at (-3.5, -3) {6};
+            \node [draw, circle, fill=white, inner sep=1pt] (M1) at (-5.5,  0) {1};
+            \node [draw, circle, fill=uofgcobalt, inner sep=1pt] (M2) at (-3.5,  0) {2};
+            \node [draw, circle, fill=uofgthistle, inner sep=1pt] (M3) at (-5.5, -1.5) {3};
+            \node [draw, circle, fill=white, inner sep=1pt] (M4) at (-3.5, -1.5) {4};
+            \node [draw, circle, fill=uofglawn, inner sep=1pt] (M5) at (-5.5, -3) {5};
+            \node [draw, circle, fill=uofgpumpkin, inner sep=1pt] (M6) at (-3.5, -3) {6};
 
@@
         \end{scope}
-    \end{tikzpicture}\\[0.75cm]
-    \begin{tikzpicture}[scale=0.5]
+    \end{tikzpicture}\\[0.25cm]
+    \begin{tikzpicture}[scale=0.45,every node/.style={font=\footnotesize}]
         \begin{scope}
             \foreach \n in {1, ..., 6}{
-                \node [draw, circle, fill=uofgcobalt, inner sep=1.5pt] (N1\n) at (\n, 0) {\n};
+                \node [draw, circle, fill=uofgcobalt, inner sep=1pt] (N1\n) at (\n, 0) {\n};
             }
@@
             \foreach \n in {1, ..., 6}{
-                \node [draw, circle, fill=uofgpumpkin, inner sep=1.5pt] (N2\n) at (0, -\n) {\n};
+                \node [draw, circle, fill=uofgpumpkin, inner sep=1pt] (N2\n) at (0, -\n) {\n};
             }
@@
             \foreach \n in {1, ..., 6}{
-                \node [draw, circle, fill=uofglawn, inner sep=1.5pt] (N3\n) at (7, -\n) {\n};
+                \node [draw, circle, fill=uofglawn, inner sep=1pt] (N3\n) at (7, -\n) {\n};
             }
@@
             \foreach \n in {1, ..., 6}{
-                \node [draw, circle, fill=uofgthistle, inner sep=1.5pt] (N4\n) at (\n, -7) {\n};
+                \node [draw, circle, fill=uofgthistle, inner sep=1pt] (N4\n) at (\n, -7) {\n};
             }
@@
 random instances used in each case came from common datasets \citep{DeSanto:2003,Zampelli:2010},
-which were generated by taking a random subgraph of a random graph (using various models, including
-Erd\H{o}s-R\'enyi, scale-free, bounded degree, and meshes) and permuting the vertices. Such
-instances are guaranteed to be satisfiable---\citet{Anton:2009} exploited this property to create
-large sets of random satisfiable boolean satisfiability instances.  However, since this has been the
-only approach used to generate random subgraph isomorphism instances, existing benchmark suites
-contain relatively few non-trivial unsatisfiable instances, and the satisfiable instances tend to be
-computationally fairly easy, with most of the difficulty being in dealing with the size of the
-model. This has lead to bias in algorithm design, to the extent that some proposed techniques will
-\emph{only} work on satisfiable instances \citet{Battiti:2007}.
+which were generated by taking a random subgraph of a random (Erd\H{o}s-R\'enyi, scale-free, bounded
+degree, or mesh) graph and permuting the vertices. Such instances are guaranteed to be
+satisfiable---\citet{Anton:2009} exploited this property to create large sets of random satisfiable
+boolean satisfiability instances.  However, since this has been the only approach used to generate
+random subgraph isomorphism instances, existing benchmark suites contain relatively few non-trivial
+unsatisfiable instances, and the satisfiable instances tend to be computationally fairly easy, with
+most of the difficulty being in dealing with the size of the model. This has lead to bias in
+algorithm design, to the extent that some proposed techniques will \emph{only} work on satisfiable
+instances \citep{Battiti:2007}.
 
@@
 
-Throughout, our graphs are unlabelled, undirected, and do not have any loops (vertices which are
-adjacent to themselves).  The \emph{order} of a graph is the cardinality of its vertex set. We write
-$\operatorname{V}(G)$ for the vertex set of a graph $G$. The \emph{complement} of a graph $G$,
-denoted $\shortoverline{G}$, is the
-graph with the same vertex set as $G$, and with an edge between distinct vertices $v$ and $w$ if and
-only if
-$v$ and $w$ are not adjacent in $G$. We write $G(n, p)$ for an Erd\H{o}s-R\'enyi random graph with
-$n$ vertices, and an edge between each distinct pair of vertices with independent probability $p$.
+Throughout, our graphs are unlabelled, undirected, and do not have any loops.  The \emph{order} of a
+graph is the cardinality of its vertex set. We write $\operatorname{V}(G)$ for the vertex set of a
+graph $G$. The \emph{complement} of a graph $G$, denoted $\shortoverline{G}$, is the graph with the
+same vertex set as $G$, and with an edge between distinct vertices $v$ and $w$ if and only if $v$
+and $w$ are not adjacent in $G$. We write $G(n, p)$ for an Erd\H{o}s-R\'enyi random graph with $n$
+vertices, and an edge between each distinct pair of vertices with independent probability $p$.
 
@@
 
-Our experiments were performed on systems with Intel Xeon E5-4650 v2 CPUs and 768GBytes RAM
-(although this much RAM was not needed), running Scientific Linux release 6.7. We selected three
-subgraph isomorphism solvers: the Glasgow solver \citep{McCreesh:2015}, LAD \citep{Solnon:2010}, and
-VF2 \citep{Cordella:2004}; each was compiled using GCC 4.9. These solvers all use backtracking
-search to build up an assignment of target vertices (values) to pattern vertices (variables), with
-different kinds of inference and ordering heuristics.
+Our experiments were performed on systems with Intel Xeon E5-4650 v2 CPUs, running Scientific Linux
+release 6.7. We selected three subgraph isomorphism solvers: the Glasgow solver
+\citep{McCreesh:2015}, LAD \citep{Solnon:2010}, and VF2 \citep{Cordella:2004}; each was compiled
+using GCC 4.9. Each uses backtracking search to build up an assignment of target vertices (values)
+to pattern vertices (variables), with different kinds of inference and ordering heuristics.
 
@@
 3SAT problems---compare, for example, Figure 1 of \citet{LeytonBrown:2014}. In particular,
-satisfiable instances tend to be easier, but show greater variation than unsatisfiable instances. % , and
-% there are exceptionally hard satisfiable instances \citep{Smith:1997}.
+satisfiable instances tend to be easier, but show greater variation than unsatisfiable instances, and
+there are exceptionally hard satisfiable instances \citep{Smith:1997}.
 
@@
 \citeauthor{Walsh:1998}'s suggestion applies only in these conditions). In any case, $\langle Sol
-\rangle$ is insufficient to explain the observed behaviour.
+\rangle$ is insufficient to explain the observed behaviour in these two eighths of the parameter space.
 
@@
 
-We also implemented an equivalent SAT encoding with direct encoding of the cardinality constraints.
-Using the Glucose solver, we saw very similar performance to the PB encoding. We also implemented a
-similar integer program encoding; the Gurobi solver was only able to solve some of the trivial
-satisfiable instances, and was almost never able to prove unsatisfiability within the time limit. We
-do not plot the results for either of these solvers.
+We also implemented (but do not plot) an equivalent SAT encoding with direct encoding of the
+cardinality constraints.  Using the Glucose solver, we saw very similar performance to the PB
+encoding. We also implemented a similar integer program encoding; the Gurobi solver was only able to
+solve some of the trivial satisfiable instances, and was almost never able to prove unsatisfiability
+within the time limit.
 
-The \emph{association graph encoding} of a subgraph isomorphism problem is constructed by creating a
-new graph with a vertex for each pair $(p, t)$ of vertices from the pattern and target graphs
-respectively. There is an edge between vertex $(p_1, t_1)$ and vertex $(p_2, t_2)$ if mapping $p_1$
-to $t_1$ and $p_2$ to $t_2$ simultaneously is permitted, i.e.\ $p_1$ is adjacent to $p_2$ if and
-only if $t_1$ is adjacent to $t_2$. A clique of size equal to the order of the pattern graph exists
-in the association graph if and only if the problem is satisfiable \citep{Levi:1973}.  We illustrate
-this in \cref{figure:sip}. We used this encoding with a version of the BBMC clique algorithm
-\citep{SanSegundo:2011} modified to solve the decision problem. Again, the instances in the central
-region remain hard, and additionally, some of the easy satisfiable instances become hard.
+The \emph{association graph encoding} of a subgraph isomorphism problem (illustrated in
+\cref{figure:sip}) is constructed by creating a new graph with a vertex for each pair $(p, t)$ of
+vertices from the pattern and target graphs respectively. There is an edge between vertex $(p_1,
+t_1)$ and vertex $(p_2, t_2)$ if mapping $p_1$ to $t_1$ and $p_2$ to $t_2$ simultaneously is
+permitted, i.e.\ $p_1$ is adjacent to $p_2$ if and only if $t_1$ is adjacent to $t_2$. A clique of
+size equal to the order of the pattern graph exists in the association graph if and only if the
+problem is satisfiable \citep{Levi:1973}. We used this encoding with a decision version of the BBMC
+clique algorithm \citep{SanSegundo:2011}. Again, the instances in the central region remain hard,
+and additionally, some of the easy satisfiable instances become hard.
 
@@
 
-?? We found hard instances, some where we'd expect, some in unexpected places. These unexpected
-instances, by and large, remained hard under reductions. We looked at how existing non-induced
-heuristics could be recovered from Sol, and how this gives us better heuristics for the induced
-variant.
+We have shown how to generate small but hard instances for the non-induced and induced subgraph
+isomorphism problems, which will help offset the bias in existing datasets. For non-induced
+isomorphisms, behaviour was as in many other hard problems, but for induced isomorphisms we
+uncovered several interesting phenomena: there are hard instances far from a phase transition,
+constrainedness predicts this, and existing general techniques for designing heuristics do not work
+in certain portions of the parameter space.
 
-?? Consider dynamic heuristics.
-
-?? Other random models and non-random.
-
-?? Does not give hard instances for graph isomorphism.
+In future work, we intend to repeat the experiments using other random models, including bounded and
+regular degree. We will also look at dynamic heuristics, and switching pattern and target heuristics
+independently. Finally, we intend to investigate whether variance can be calculated efficiently
+enough to give better predictions for very sparse or dense pattern graphs.
 

### 2016-02-02 paper/paper.tex
Final typos

@@
 
-\author{First Author \thanks{This work was supported by a small pile of money and a large
-    electricity bill.} \and Second Author
+\author{First Author \thanks{This work was supported by a small pile of money from an anonymous
+        funding agency, and a large electricity bill.} \and Second Author
     \and Third Author \\
@@
 the original pattern and target as the input to the Glasgow algorithm, or to take the complements.
-(The only steps performed by the Glagsow algorithm which differs under taking the complements are
+(The only steps performed by the Glagsow algorithm which differ under taking the complements are
 the degree-based heuristics.  LAD and VF2 are not symmetric in this way: LAD performs a filtering
@@
 are usually easy, which perhaps explains the continuing popularity of VF2's connectivity-based
-heuristics \citep{Carletti:2015}. In this situation, these experiments justify re-using the
+heuristics \citep{Carletti:2015}. In this situation, these experiments justify reusing the
 non-induced heuristics on induced problems.
@@
 
-The pseudo-boolean (PB) encoding is as follows. For each pattern vertex $v$ and each target
-vertex $w$, we have a binary variable which takes the value 1 if and only if $v$ is mapped to
-$w$.  Constraints are added to ensure that each pattern vertex maps to exactly one target vertex,
-that each target vertex is mapped to by at most one pattern vertex, that adjacent vertices are
-mapped to adjacent vertices, and that non-adjacent vertices are mapped to non-adjacent vertices. We
-used the Clasp solver~\citep{gekakaosscsc11a} version 3.1.3 to solve the pseudo-boolean instances.
-The instances that are hard for the Glasgow solver remain hard for the PB solver, including
-instances inside the central region, and the easy satisfiable instances remain easy.
-
-We also implemented (but do not plot) an equivalent SAT encoding with direct encoding of the
-cardinality constraints.  Using the Glucose solver, we saw very similar performance to the PB
-encoding. We also implemented a similar integer program encoding; the Gurobi solver was only able to
-solve some of the trivial satisfiable instances, and was almost never able to prove unsatisfiability
-within the time limit.
+The pseudo-boolean (PB) encoding is as follows. For each pattern vertex $v$ and each target vertex
+$w$, we have a binary variable which takes the value 1 if and only if $v$ is mapped to $w$.
+Constraints are added to ensure that each pattern vertex maps to exactly one target vertex, that
+each target vertex is mapped to by at most one pattern vertex, that adjacent vertices are mapped to
+adjacent vertices, and that non-adjacent vertices are mapped to non-adjacent vertices. We used the
+Clasp solver~\citep{gekakaosscsc11a} version 3.1.3 to solve the pseudo-boolean instances.  The
+instances that are hard for the Glasgow solver remain hard for the PB solver, including instances
+inside the central region, and the easy satisfiable instances remain easy. (Similar results were
+seen with the Glucose SAT solver using a direct encoding of the cardinality constraints. We also
+implemented an integer program encoding; the Gurobi solver was only able to solve some of the
+trivial satisfiable instances, and was almost never able to prove unsatisfiability within the time
+limit.)
 
@@
 These results suggest that the central region may be genuinely hard, despite not being near a phase
-transition. The clique encoding in particular rules out the hypothesis that subgraph isomorphism
+transition. The clique results in particular rule out the hypothesis that subgraph isomorphism
 solvers only find this region hard due to not reasoning simultaneously about adjacency and

### 2016-04-12 paper/paper.tex
Deanonymise

@@
 
-% \author{Ciaran McCreesh\thanks{This work was supported by the Engineering and Physical Sciences
-%     Research Council [grant number EP/K503058/1]} \and Patrick Prosser \and James Trimble \\
-% University of Glasgow, Glasgow, Scotland \\
-% c.mccreesh.1@research.gla.ac.uk}
-
-\author{First Author \thanks{This work was supported by a small pile of money from an anonymous
-        funding agency, and a large electricity bill.} \and Second Author
-    \and Third Author \\
-Some Instituion, Somewhere, Some Country \\
-anonymous@example.com}
+\author{Ciaran McCreesh\thanks{This work was supported by the Engineering and Physical Sciences
+    Research Council [grant number EP/K503058/1]} \and Patrick Prosser \and James Trimble \\
+University of Glasgow, Glasgow, Scotland \\
+c.mccreesh.1@research.gla.ac.uk}
 
@@
 
-% \section*{Acknowledgements}
-% 
-% Thanks to Kitty Meeks. Also, I suppose Craig could be given an inkling of undeserved gratitude if
-% we have space left.
+\section*{Acknowledgements}
+
+The authors wish to thank Kitty Meeks and Craig Reilly for their comments.
 

### 2016-04-12 paper/paper.tex
We're buying an extra page

@@
 
-%%% The lack of unsatisfiable instances cannot be addressed by using a pattern graph from one of the
-%%% random suites with the ``wrong'' target graph: this tends to give either a trivially unsatisfiable
-%%% instance, or a satisfiable instance. (In particular, it is \emph{not} the case that a relatively
-%%% small random graph is unlikely to appear in a larger random graph.)
+The lack of unsatisfiable instances cannot be addressed by using a pattern graph from one of the
+random suites with the ``wrong'' target graph: this tends to give either a trivially unsatisfiable
+instance, or a satisfiable instance. (In particular, it is \emph{not} the case that a relatively
+small random graph is unlikely to appear in a larger random graph.)
 
@@
 satisfiable instances tend to be easier, but show greater variation than unsatisfiable instances, and
-there are exceptionally hard satisfiable instances \citep{Smith:1997}.
+there are exceptionally hard satisfiable instances \citep{Smith:1997}. (The Glasgow solver supports
+parallel search with a work-stealing strategy explicitly designed to eliminate these. We have not
+enabled this option to avoid dealing with the complexity of search tree measurements under
+parallelism.)
 

### 2016-04-12 paper/paper.tex
Typo

@@
     \input{gen-graph-phase-transition.tex}
-    \caption{With a fixed a pattern graph order of 20, a target graph order of 150, a target edge
+    \caption{With a fixed pattern graph order of 20, a target graph order of 150, a target edge
         probability of 0.40, and varying pattern edge probability, we observe a phase transition and

### 2016-04-12 paper/paper.tex
A bit more on variance

@@
 that a high expected number of solutions implies that any particular instance is likely to be
-satisfiable---a similar behaviour is seen with random constraint satisfaction problems
-\citep{Smith:1996}.
-
-% \citep{Smith:1994}
+satisfiable. (Consider, for example, a sparse graph which has several isolated vertices. If one
+solution exists, other symmetric solutions can be obtained by permuting the isolated vertices.
+Thus although the expected number of solutions may be one, there cannot be exactly one solution.) A
+similar behaviour is seen with random constraint satisfaction problems
+\citep{Smith:1994,Smith:1996}.
 

### 2016-04-12 paper/paper.tex
Scale-free
Finer data

@@
 row of \cref{figure:non-induced} we show the satisfiability phase transition for the non-induced
-variant, for patterns of order 10, 20 and 30, targets of order 150, and varying pattern (x-axis)
-and target (y-axis) edge probabilities. Each axis runs over 51 edge probabilities, from 0 to 1 in
-steps of 0.02; for each of these 2601 points, we generate ten random instances. The colour denotes
-the proportion of these instances which were found to be satisfiable.  Inside the orange region, at
-the bottom right of each plot, every instance is unsatisfiable---here we are trying to find a dense
-pattern in a sparse target. In the purple region, at the top left, every instance is
-satisfiable---we are looking for a sparse pattern in a dense target (which is easy, since we only
-have to preserve adjacency, not non-adjacency). The white band between the regions shows the
-location of the phase transition: here, roughly half the instances are satisfiable. (We discuss the
-black line below.)
+variant, for patterns of order 10, 20 and 30, targets of order 150, and varying pattern (x-axis) and
+target (y-axis) edge probabilities. Each axis runs over 101 edge probabilities, from 0 to 1 in steps
+of 0.01, except for the VF2 row which uses steps of 0.02. For each of these points, we generate ten
+random instances. The colour denotes the proportion of these instances which were found to be
+satisfiable.  Inside the orange region, at the bottom right of each plot, every instance is
+unsatisfiable---here we are trying to find a dense pattern in a sparse target. In the purple region,
+at the top left, every instance is satisfiable---we are looking for a sparse pattern in a dense
+target (which is easy, since we only have to preserve adjacency, not non-adjacency). The white band
+between the regions shows the location of the phase transition: here, roughly half the instances are
+satisfiable. (We discuss the black line below.)
 
@@
 
-In future work, we intend to repeat the experiments using other random models, including bounded and
-regular degree. We will also look at dynamic heuristics, and switching pattern and target heuristics
-independently. Finally, we intend to investigate whether variance can be calculated efficiently
-enough to give better predictions for very sparse or dense pattern graphs.
+In future work, we intend to repeat the experiments using other random models, including bounded,
+regular degree, and scale-free. We will also look at dynamic heuristics, and switching pattern and
+target heuristics independently. Finally, we intend to investigate whether variance can be
+calculated efficiently enough to give better predictions for very sparse or dense pattern graphs.
 

### 2016-04-12 prediction/induced-kappa.rb

@@ -6,10 +6,10 @@ require 'bigdecimal'
 ps = ARGV[0].to_i
 ts = ARGV[1].to_i
 
-0.upto(50) do | y |
-    0.upto(50) do | x |
-        pd = BigDecimal.new(x) / 50;
-        td = BigDecimal.new(y) / 50;
+0.upto(100) do | y |
+    0.upto(100) do | x |
+        pd = BigDecimal.new(x) / 100;
+        td = BigDecimal.new(y) / 100;
 
         cliqueedges = BigDecimal.new(ps * (ps - 1) / 2);
         nedges = pd * cliqueedges;

### 2016-04-12 prediction/induced-line-actual.rb

@@ -85,8 +85,8 @@ end
             tdresult[pass] = (upper + lower) / 1000.0
         end
 
-        pdround[pass] = (pdresult[pass] * 50.0).to_i
-        tdround[pass] = (tdresult[pass] * 50.0).to_i
+        pdround[pass] = (pdresult[pass] * 100.0).to_i
+        tdround[pass] = (tdresult[pass] * 100.0).to_i
 
         while tdround[pass] < 49 and tdround[pass] > 1 and measured[pdround[pass]][tdround[pass]].to_f < 0.5 do
             tdround[pass] += (pass == 0 ? 1 : -1)

### 2016-04-14 paper/paper.tex
Post-review tinkering

@@
 \usepackage{cleveref}
-\usepackage{natbib}
 \usepackage{dblfloatfix}
@@
 
+% Without these, figure 4 causes horrible page layout problems
 \renewcommand{\topfraction}{0.9}
@@
 
+\newcommand{\citet}[1]{\citeauthor{#1} \shortcite{#1}}
+\newcommand{\citep}[1]{\cite{#1}}
+
 % cref style
@@
 \newcommand{\shortoverline}[1]{\mkern 1.5mu\overline{\mkern-1.5mu#1\mkern-1.5mu}\mkern 1.5mu}
+
+\makeatletter\@ifpackageloaded{showframe}{
+    % show lines down the middle columns too
+    % http://tex.stackexchange.com/questions/16199/test-if-a-package-or-package-option-is-loaded
+\newlength\Fcolumnseprule\setlength\Fcolumnseprule{0.4pt}\def\@outputdblcol{%
+  \if@firstcolumn\global \@firstcolumnfalse \global \setbox\@leftcolumn \box\@outputbox
+  \else
+    \global \@firstcolumntrue
+    \setbox\@outputbox \vbox{\hb@xt@\textwidth{\hb@xt@\columnwidth{\box\@leftcolumn \hss}%
+    \vrule \@width\Fcolumnseprule\hfil{\normalcolor\vrule \@width\columnseprule}
+    \hfil\vrule \@width\Fcolumnseprule\hb@xt@\columnwidth {\box\@outputbox \hss}}}\@combinedblfloats\@outputpage
+    \begingroup\@dblfloatplacement\@startdblcolumn\@whilesw\if@fcolmade \fi{\@outputpage\@startdblcolumn}\endgroup
+  \fi
+}}{}
+\makeatother
 
@@
     \centering
-    \begin{tikzpicture}[scale=0.45,every node/.style={font=\footnotesize}]
+    \begin{tikzpicture}[scale=0.5,every node/.style={font=\footnotesize}]
         \begin{scope}
-            \node [draw, circle, fill=uofgcobalt, inner sep=1pt] (Na) at (1,  0) {1};
-            \node [draw, circle, fill=uofgpumpkin, inner sep=1pt] (Nb) at (1, -1.5) {2};
-            \node [draw, circle, fill=uofglawn, inner sep=1pt] (Nc) at (0, -3) {3};
-            \node [draw, circle, fill=uofgthistle, inner sep=1pt] (Nd) at (2, -3) {4};
+            \node [draw, circle, fill=uofgcobalt, inner sep=1.5pt] (Na) at (1,  0) {1};
+            \node [draw, circle, fill=uofgpumpkin, inner sep=1.5pt] (Nb) at (1, -1.5) {2};
+            \node [draw, circle, fill=uofglawn, inner sep=1.5pt] (Nc) at (0, -3) {3};
+            \node [draw, circle, fill=uofgthistle, inner sep=1.5pt] (Nd) at (2, -3) {4};
 
@@
 
-            \node [draw, circle, fill=uofgcobalt, inner sep=1pt] (N1) at (5.5,  0) {1};
-            \node [draw, circle, fill=white, inner sep=1pt] (N2) at (7.5,  0) {2};
-            \node [draw, circle, fill=uofgpumpkin, inner sep=1pt] (N3) at (5.5, -1.5) {3};
-            \node [draw, circle, fill=white, inner sep=1pt] (N4) at (7.5, -1.5) {4};
-            \node [draw, circle, fill=uofglawn, inner sep=1pt] (N5) at (5.5, -3) {5};
-            \node [draw, circle, fill=uofgthistle, inner sep=1pt] (N6) at (7.5, -3) {6};
+            \node [draw, circle, fill=uofgcobalt, inner sep=1.5pt] (N1) at (5.5,  0) {1};
+            \node [draw, circle, fill=white, inner sep=1.5pt] (N2) at (7.5,  0) {2};
+            \node [draw, circle, fill=uofgpumpkin, inner sep=1.5pt] (N3) at (5.5, -1.5) {3};
+            \node [draw, circle, fill=white, inner sep=1.5pt] (N4) at (7.5, -1.5) {4};
+            \node [draw, circle, fill=uofglawn, inner sep=1.5pt] (N5) at (5.5, -3) {5};
+            \node [draw, circle, fill=uofgthistle, inner sep=1.5pt] (N6) at (7.5, -3) {6};
 
@@
 
-            \node [draw, circle, fill=white, inner sep=1pt] (M1) at (-5.5,  0) {1};
-            \node [draw, circle, fill=uofgcobalt, inner sep=1pt] (M2) at (-3.5,  0) {2};
-            \node [draw, circle, fill=uofgthistle, inner sep=1pt] (M3) at (-5.5, -1.5) {3};
-            \node [draw, circle, fill=white, inner sep=1pt] (M4) at (-3.5, -1.5) {4};
-            \node [draw, circle, fill=uofglawn, inner sep=1pt] (M5) at (-5.5, -3) {5};
-            \node [draw, circle, fill=uofgpumpkin, inner sep=1pt] (M6) at (-3.5, -3) {6};
+            \node [draw, circle, fill=white, inner sep=1.5pt] (M1) at (-5.5,  0) {1};
+            \node [draw, circle, fill=uofgcobalt, inner sep=1.5pt] (M2) at (-3.5,  0) {2};
+            \node [draw, circle, fill=uofgthistle, inner sep=1.5pt] (M3) at (-5.5, -1.5) {3};
+            \node [draw, circle, fill=white, inner sep=1.5pt] (M4) at (-3.5, -1.5) {4};
+            \node [draw, circle, fill=uofglawn, inner sep=1.5pt] (M5) at (-5.5, -3) {5};
+            \node [draw, circle, fill=uofgpumpkin, inner sep=1.5pt] (M6) at (-3.5, -3) {6};
 
@@
         \end{scope}
-    \end{tikzpicture}\\[0.25cm]
+    \end{tikzpicture}\\[0.5cm]
     \begin{tikzpicture}[scale=0.45,every node/.style={font=\footnotesize}]
@@
 region---these instances provide evidence in favour of constrainedness, rather than proximity to a
-phase transition, being an accurate predictor of difficulty.
+phase transition, being an accurate predictor of difficulty, and show that constrainedness is not
+simply a refinement of a phase transition prediction.
 
@@
 \citep{McCreesh:2015}, LAD \citep{Solnon:2010}, and VF2 \citep{Cordella:2004}; each was compiled
-using GCC 4.9. Each uses backtracking search to build up an assignment of target vertices (values)
-to pattern vertices (variables), with different kinds of inference and ordering heuristics.
+using GCC 4.9.
+
+The Glasgow and LAD solvers use backtracking search to build up an assignment of target vertices
+(values) to pattern vertices (variables), but differ in terms of inference and ordering heuristics.
+The approach used by VF2 is similar, although the domains of variables are not stored (in the style
+of conventional backtracking, rather than forward-checking), and so domain wipeouts are not detected
+until an assignment is made.
 
@@
 \begin{figure}[tb]
-    \setlength{\abovecaptionskip}{6pt}
     \input{gen-graph-phase-transition.tex}
@@
 \begin{figure}[tb]
-    \setlength{\abovecaptionskip}{0pt}
     \input{gen-graph-non-induced.tex}
@@
 parameters. Since we are trying to find an \emph{injective} mapping from a pattern $P = G(p, d_p)$
-to a target $T = G(t, d_t)$, there are $t^{\underline{p}} = t \cdot (t - 1) \cdot \ldots \cdot (t -
-p + 1)$ possible assignments of target vertices to pattern vertices.  We expect the pattern to have
+to a target $T = G(t, d_t)$, there are \[ t^{\underline{p}} = t \cdot (t - 1) \cdot \ldots \cdot (t -
+p + 1) \] possible assignments of target vertices to pattern vertices.  We expect the pattern to have
 $d_p \cdot \binom{p}{2}$ edges, so we obtain the probability of all of these edges being mapped to
@@
 formula predicts a very sharp phase transition from $\langle Sol \rangle \ll 1$ to $\langle Sol
-\rangle \gg 1$. We plot where this occurs using black lines in the first row of
-\cref{figure:non-induced}.
+\rangle \gg 1$, which may easily be located numerically. We plot where this occurs using black lines
+in the first row of \cref{figure:non-induced}.
 
@@
 domains are of equal size, LAD simply breaks ties lexicographically, whereas Glasgow will pick a
-variable corresponding to a pattern vertex of highest degree. This strategy was apparently
-determined empirically, but could have been derived from the $\langle Sol \rangle$ formula: picking
-a pattern vertex of high degree will make the remaining pattern subgraph sparser, which will
-decrease the exponent in the second half of the formula, maximising the overall value. LAD does not
-apply a value ordering heuristic, but Glasgow does: it prefers target vertices of lowest degree.
-Again, this appears to have been determined empirically, but it has the effect of increasing
-$\langle Sol \rangle$ by increasing the remaining target density. The VF2 heuristics, in contrast,
-are based around preserving connectivity, which gives very little discrimination except on the
-sparsest of inputs.
+variable corresponding to a pattern vertex of highest degree. This strategy was determined
+empirically, but could have been derived from the $\langle Sol \rangle$ formula: picking a pattern
+vertex of high degree will make the remaining pattern subgraph sparser, which will decrease the
+exponent in the second half of the formula, maximising the overall value. LAD does not apply a value
+ordering heuristic, but Glasgow does: it prefers target vertices of lowest degree.  Again, this was
+determined empirically, but it has the effect of increasing $\langle Sol \rangle$ by increasing the
+remaining target density. The VF2 heuristics, in contrast, are based around preserving connectivity,
+which gives very little discrimination except on the sparsest of inputs.
 
@@
 \begin{figure*}[t]
-    \setlength{\abovecaptionskip}{0pt}
     \input{gen-graph-induced.tex}
@@
 \begin{figure*}[t]
-    \setlength{\abovecaptionskip}{0pt}
     \input{gen-graph-sat.tex}
@@
 
+The model we have proposed may be extended to graphs with labels on the vertices or edges, which
+arise naturally in many applications. Broadly speaking, randomly allocated labels make the problem
+easier by restricting the search space, although there are pathological cases where increasing the
+number of labels shifts an instance from the easy satisfiable region to being close to the phase
+transition. The model may also be extended to the (decision version of the) maximum common subgraph
+problem. The reduction to maximum clique is a competitive approach for maximum common subgraph
+problem, so it is interesting to observe that some of the easy portions of the search space become
+hard under this reduction---this suggests a possible unnecessary weakness of maximum clique
+algorithms on certain kinds of input. (Maximum clique algorithms use a greedy colouring as a bound,
+and the association graphs produced by this reduction can be very bad for greedy colourings in the
+same way that crown graphs are.)
+
+In contrast, it is worth noting that this technique does \emph{not} give a way of generating hard
+instances for graph isomorphism problems: the pattern graph must be substantially smaller than the
+target graph for independent pairs of randomly generated instances to give interesting behaviour.
+
 In future work, we intend to repeat the experiments using other random models, including bounded,
-regular degree, and scale-free. We will also look at dynamic heuristics, and switching pattern and
+regular degree, and scale-free; regular degree graphs in particular foil existing degree-based
+heuristics. Similarly, we will look at alternatives to simple randomness for labelling
+strategies---for example, in chemical datasets, label sets tend to differ based upon the degrees of
+their corresponding vertices.  We will also look at dynamic heuristics, and switching pattern and
 target heuristics independently.  Finally, we intend to investigate whether variance can be

### 2016-04-15 paper/paper.tex
James's comments

@@
 transition. The model may also be extended to the (decision version of the) maximum common subgraph
-problem. The reduction to maximum clique is a competitive approach for maximum common subgraph
+problem. The reduction to maximum clique is a competitive approach for the maximum common subgraph
 problem, so it is interesting to observe that some of the easy portions of the search space become

### 2016-04-15 paper/paper.tex
Patrick's comments

@@
 builds upon the phase transition phenomena observed in satisfiability and graph colouring problems
-first described by \citet{Cheeseman:1991} and \citet{Mitchell:1992}.  For subgraph isomorphism we
+first described by \citet{DBLP:conf/ijcai/CheesemanKT91} and \citet{Mitchell:1992}.  For subgraph isomorphism we
 identify three relevant control parameters: we can independently alter the edge probability of the

### 2016-04-15 paper/paper.tex
Expand text a bit

@@
 size equal to the order of the pattern graph exists in the association graph if and only if the
-problem is satisfiable \citep{Levi:1973}. We used this encoding with a decision version of the BBMC
-clique algorithm \citep{SanSegundo:2011}. Again, the instances in the central region remain hard,
-and additionally, some of the easy satisfiable instances become hard.
+problem is satisfiable \citep{Levi:1973}. We used this encoding with the BBMC clique algorithm
+\citep{SanSegundo:2011}, which we implemented in C++. Usually BBMC solves the optimisation version
+of the clique problem, we adapted it to solve the decision problem by initialising the incumbent to
+be one less than the decision value, and allowing it to exit as soon as a clique with size equal to
+the decision value is encountered.  Again, our results show that the instances in the central region
+remain hard, and additionally, some of the easy unsatisfiable instances become hard.
 
-These results suggest that the central region may be genuinely hard, despite not being near a phase
-transition. The clique results in particular rule out the hypothesis that subgraph isomorphism
-solvers only find this region hard due to not reasoning simultaneously about adjacency and
-non-adjacency, since the association graph encoding constraints consider compatibility rather than
-adjacency and non-adjacency.
+Together, these experiments suggest that the central region may be genuinely hard, despite not being
+near a phase transition. The clique results in particular rule out the hypothesis that subgraph
+isomorphism solvers only find this region hard due to not reasoning simultaneously about adjacency
+and non-adjacency, since the association graph encoding constraints consider compatibility rather
+than adjacency and non-adjacency.
 
-Constrainedness, denoted $\kappa$, is an alternative measure of difficulty designed to generalise
-hardness parameters across different combinatorial problems \citep{Gent:1996:Kappa}. A problem with
-$\kappa < 1$ is said to be underconstrained, and is likely to be satisfiable; a problem with $\kappa
-> 1$ is overconstrained, and is likely to be unsatisfiable. Empirically, problems with $\kappa$
-close to 1 are hard, and problems where $\kappa$ is very small or very large are usually easy. By
-handling injectivity as a restriction on the size of the state space rather than as a constraint, we
-derive \[ \kappa = 1 - \frac{\log \left( t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot {(1 -
-d_{t})}^{(1 - d_{p}) \cdot \binom{p}{2}} \right)}{\log t^{\underline{p}}}\] for induced isomorphisms, which
-we plot on the fifth row of \cref{figure:induced}. We see that constrainedness predicts that the
-central region will still be relatively difficult for larger pattern graphs: although the problem is
-overconstrained, it is less overconstrained than in the regions the Glasgow and LAD solvers found easy.
-Thus it seems that rather than just being a unification of existing generalised heuristic
+Constrainedness, denoted $\kappa$, is an alternative measure of difficulty designed to refine the
+phase transition concept, and to generalise hardness parameters across different combinatorial
+problems \citep{Gent:1996:Kappa}. A problem with $\kappa < 1$ is said to be underconstrained, and is
+likely to be satisfiable; a problem with $\kappa > 1$ is overconstrained, and is likely to be
+unsatisfiable. Empirically, problems with $\kappa$ close to 1 are hard, and problems where $\kappa$
+is very small or very large are usually easy. By handling injectivity as a restriction on the size
+of the state space rather than as a constraint, we derive \[ \kappa = 1 - \frac{\log \left(
+t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot {(1 - d_{t})}^{(1 - d_{p}) \cdot
+\binom{p}{2}} \right)}{\log t^{\underline{p}}}\] for induced isomorphisms, which we plot on the
+fifth row of \cref{figure:induced}. We see that constrainedness predicts that the central region
+will still be relatively difficult for larger pattern graphs: although the problem is
+overconstrained, it is less overconstrained than in the regions the Glasgow and LAD solvers found
+easy.  Thus it seems that rather than just being a unification of existing generalised heuristic
 techniques, constrainedness also gives a better predictor of difficulty than proximity to a phase
-transition.  Unfortunately, constrainedness does not help us with heuristics: minimising
-constrainedness gives the same predictions as maximising the expected number of solutions.
+transition---our method generates instances where constrainedness and ``close to a phase
+transition'' give very different predictions, and constrainedness gives the better predictions.
+
+Unfortunately, constrainedness does not help us with heuristics: minimising constrainedness gives
+the same predictions as maximising the expected number of solutions.
 

### 2016-04-17 paper/paper.tex
We've room for another heading
Fix weird hyphenation

@@
 
-\subsection{Why is the Central Region Hard?}
+\subsection{Is the Central Region Genuinely Hard?}
 
@@
 
+\subsection{Constrainedness}
+
 Constrainedness, denoted $\kappa$, is an alternative measure of difficulty designed to refine the
@@
 transition. The model may also be extended to the (decision version of the) maximum common subgraph
-problem. The reduction to maximum clique is a competitive approach for the maximum common subgraph
-problem, so it is interesting to observe that some of the easy portions of the search space become
-hard under this reduction---this suggests a possible unnecessary weakness of maximum clique
-algorithms on certain kinds of input. (Maximum clique algorithms use a greedy colouring as a bound,
-and the association graphs produced by this reduction can be very bad for greedy colourings in the
-same way that crown graphs are.)
+problem. The maximum clique approach is competitive for the maximum common subgraph problem, so it
+is interesting to observe that some of the easy portions of the search space become hard under this
+reduction---this suggests a possible unnecessary weakness of maximum clique algorithms on certain
+kinds of input. (Maximum clique algorithms use a greedy colouring as a bound, and the association
+graphs produced by this reduction can be very bad for greedy colourings in the same way that crown
+graphs are.)
 

### 2016-04-17 paper/paper.tex
PDF Info
Semicolon

@@
 c.mccreesh.1@research.gla.ac.uk}
+
+\pdfinfo{
+    /Title (Heuristics and Really Hard Instances for Subgraph Isomorphism Problems)
+    /Author (Ciaran McCreesh, Patrick Prosser, James Trimble)
+}
 
@@
 \citep{SanSegundo:2011}, which we implemented in C++. Usually BBMC solves the optimisation version
-of the clique problem, we adapted it to solve the decision problem by initialising the incumbent to
+of the clique problem; we adapted it to solve the decision problem by initialising the incumbent to
 be one less than the decision value, and allowing it to exit as soon as a clique with size equal to

### 2016-04-17 paper/paper.tex
Not plural

@@
 transition---our method generates instances where constrainedness and ``close to a phase
-transition'' give very different predictions, and constrainedness gives the better predictions.
+transition'' give very different predictions, and constrainedness gives the better prediction.
 

### 2016-04-17 paper/paper.tex
Preliminary

@@
 reduction---this suggests a possible unnecessary weakness of maximum clique algorithms on certain
-kinds of input. (Maximum clique algorithms use a greedy colouring as a bound, and the association
-graphs produced by this reduction can be very bad for greedy colourings in the same way that crown
-graphs are.)
+kinds of input. (Maximum clique algorithms use a greedy colouring as a bound, and a preliminary
+investigation suggests that the association graphs produced by this reduction can be very bad for
+greedy colourings in the same way that crown graphs are.)
 

### 2016-04-18 paper/paper.tex
Include Glucose plots

@@
 instances that are hard for the Glasgow solver remain hard for the PB solver, including instances
-inside the central region, and the easy satisfiable instances remain easy. (Similar results were
-seen with the Glucose SAT solver using a direct encoding of the cardinality constraints. We also
+inside the central region, and the easy satisfiable instances remain easy. Similar results were
+seen with the Glucose SAT solver \cite{glucose} using a direct encoding of the cardinality constraints. (We also
 implemented an integer program encoding; the Gurobi solver was only able to solve some of the

### 2016-04-18 paper/paper.tex
Last correction, honest

@@
         \cref{figure:non-induced}. The second row shows the number of search nodes used by the
-    Glasgow algorithm, the third row shows the number of decisions made by the Clasp pseudo-boolean
-solver, and the fourth row shows the number of search nodes used by BBMC on the clique
+    Glasgow algorithm, the third and fourth rows show the number of decisions made by the pseudo-boolean
+and SAT solvers, and the final shows the number of search nodes used on the clique
 encoding.}\label{figure:alt}

### 2016-05-11 poster/poster.tex
Start on a poster

% vim: set spell spelllang=en tw=100 et sw=4 sts=4 :

\documentclass[a0paper]{tikzposter}

\usepackage{complexity}
\usepackage{wrapfig}
\usepackage{microtype}
\usepackage{gnuplot-lua-tikz}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{sfmath}

\usepackage{lmodern}
\renewcommand*\familydefault{\sfdefault}
\usepackage[T1]{fontenc}

\title{Heuristics and Really Hard Instances for Subgraph Isomorphism Problems}
\author{Ciaran McCreesh, Patrick Prosser and James Trimble}
\institute{University of Glasgow, Glasgow, Scotland}
\titlegraphic{\includegraphics[keepaspectratio=true,scale=3.5]{UoG_keyline.pdf}}

\settitle{
    \begin{tikzpicture}
        \node (T) [inner sep=0pt] {\begin{minipage}{\linewidth}
                \color{titlefgcolor}
                {\bfseries \Huge \hspace*{10mm}Heuristics and Really Hard Instances for \\
                \hspace*{10mm}Subgraph Isomorphism Problems \par}
                \vspace*{1em}
                {\Large {\bfseries \hspace{10mm}\@author}, \@institute}
        \end{minipage}};

        \node at (T.east) [anchor=center, inner sep=0pt, xshift=-12cm] {\@titlegraphic};
    \end{tikzpicture}
}

% University of Glasgow standard colours
\definecolor{uofguniversityblue}{rgb}{0, 0.219608, 0.396078}

\definecolor{uofgheather}{rgb}{0.356863, 0.32549, 0.490196}
\definecolor{uofgaquamarine}{rgb}{0.603922, 0.72549, 0.678431}
\definecolor{uofgslate}{rgb}{0.309804, 0.34902, 0.380392}
\definecolor{uofgrose}{rgb}{0.823529, 0.470588, 0.709804}
\definecolor{uofgmocha}{rgb}{0.709804, 0.564706, 0.47451}

\definecolor{uofglawn}{rgb}{0.517647, 0.741176, 0}
\definecolor{uofgcobalt}{rgb}{0, 0.615686, 0.92549}
\definecolor{uofgturquoise}{rgb}{0, 0.709804, 0.819608}
\definecolor{uofgsunshine}{rgb}{1.0, 0.862745, 0.211765}
\definecolor{uofgpumpkin}{rgb}{1.0, 0.72549, 0.282353}
\definecolor{uofgthistle}{rgb}{0.584314, 0.070588, 0.447059}
\definecolor{uofgpillarbox}{rgb}{0.701961, 0.047059, 0}
\definecolor{uofglavendar}{rgb}{0.356863, 0.301961, 0.580392}

\definecolor{uofgsandstone}{rgb}{0.321569, 0.278431, 0.231373}
\definecolor{uofgforest}{rgb}{0, 0.317647, 0.2}
\definecolor{uofgburgundy}{rgb}{0.490196, 0.133333, 0.223529}
\definecolor{uofgrust}{rgb}{0.603922, 0.227451, 0.023529}

\definecolorstyle{UofG}{
}{
    % Background Colors
    \colorlet{backgroundcolor}{uofgsandstone!80!white}
    \colorlet{framecolor}{black}
    % Title Colors
    \colorlet{titlefgcolor}{white}
    \colorlet{titlebgcolor}{uofguniversityblue}
    % Block Colors
    \colorlet{blocktitlebgcolor}{white}
    \colorlet{blocktitlefgcolor}{uofguniversityblue}
    \colorlet{blockbodybgcolor}{white}
    \colorlet{blockbodyfgcolor}{black}
    % Innerblock Colors
    \colorlet{innerblocktitlebgcolor}{uofguniversityblue}
    \colorlet{innerblocktitlefgcolor}{black}
    \colorlet{innerblockbodybgcolor}{uofgsandstone}
    \colorlet{innerblockbodyfgcolor}{black}
    % Note colors
    \colorlet{notefgcolor}{black}
    \colorlet{notebgcolor}{uofgrust}
    \colorlet{noteframecolor}{red}
}

\usetheme{Autumn}
\usecolorstyle{UofG}

\tikzposterlatexaffectionproofoff

\useblockstyle[bodyverticalshift=-1cm, roundedcorners=1]{Default}

\renewcommand{\Huge}{\fontsize{77.2}{96}\selectfont}

% Styles for drawings

\tikzset{edge/.style={line width=3pt, color=uofgsandstone}}
\tikzset{ledge/.style={line width=3pt, color=uofgsandstone!40!white}}
\tikzset{hedge/.style={line width=3pt, color=uofgsandstone, dashed}}

\setlength\intextsep{0pt}

\begin{document}
\maketitle

{
    \colorlet{blockbodybgcolor}{uofgcobalt}
    \colorlet{blocktitlebgcolor}{uofgcobalt}
    \block[bodyverticalshift=0cm, bodyinnersep=3mm]{}{
        \centering\begin{minipage}{0.94\textwidth}
            We show how to generate \textbf{really hard} random instances for \textbf{subgraph
            isomorphism} problems. For the non-induced variant, we predict and observe a phase
            transition between satisfiable and unsatisfiable instances, with a corresponding
            complexity peak seen in \textbf{three different solvers}. For the induced variant, much
            richer behaviour is observed, and \textbf{constrainedness} gives a better measure of
            difficulty than does proximity to a phase transition. We also discuss \textbf{variable
            and value ordering} heuristics, and their relationship to the \textbf{expected number of
            solutions}.
        \end{minipage}
    }
}

\begin{columns}
\column{0.5}

\block{Subgraph Isomorphism (Finding Patterns in Graphs)}{
\begin{wrapfigure}[11]{r}{0.46\linewidth}
    \begin{center}\begin{tikzpicture}[scale=1.1]%{{{
        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (Na) at (1,  0)
        {\vphantom{0}};
        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (Nb) at (1, -2)
        {\vphantom{0}};
        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (Nc) at (0, -4)
        {\vphantom{0}};
        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (Nd) at (2, -4)
        {\vphantom{0}};

        \draw [edge] (Na) -- (Nb);
        \draw [edge] (Nb) -- (Nc);
        \draw [edge] (Nc) -- (Nd);
        \draw [edge] (Nb) -- (Nd);

        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (N1) at (7.5,  0) {\vphantom{0}};
        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N2) at (9.5,  0) {\vphantom{0}};
        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (N3) at (7.5, -2) {\vphantom{0}};
        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N4) at (9.5, -2) {\vphantom{0}};
        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (N5) at (7.5, -4) {\vphantom{0}};
        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (N6) at (9.5, -4) {\vphantom{0}};
        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N7) at (5.5,  0) {\vphantom{0}};
        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N8) at (5.5, -2) {\vphantom{0}};
        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N9) at (5.5, -4) {\vphantom{0}};

        \draw [ledge] (N1) -- (N2);
        \draw [edge] (N1) -- (N3);
        \draw [ledge] (N1) -- (N4);
        \draw [ledge] (N2) -- (N4);
        \draw [edge] (N3) -- (N5);
        \draw [edge] (N3) -- (N6);
        \draw [ledge] (N4) -- (N6);
        \draw [edge] (N5) -- (N6);
        \draw [ledge] (N2) to [in=45, out=315] (N6);
        \draw [ledge] (N1) -- (N7);
        \draw [ledge] (N5) -- (N9);
        \draw [ledge] (N7) -- (N8);
        \draw [ledge] (N8) -- (N9);
    \end{tikzpicture}\end{center}

    \vspace*{1.5cm}

    \begin{center}\begin{tikzpicture}[scale=1.1]%{{{
        \node[draw, circle, fill=uofgcobalt, inner sep=5pt, font=\bfseries] (Na) at (1,  0)
        {\vphantom{0}};
        \node[draw, circle, fill=uofgcobalt, inner sep=5pt, font=\bfseries] (Nb) at (1, -2)
        {\vphantom{0}};
        \node[draw, circle, fill=uofgcobalt, inner sep=5pt, font=\bfseries] (Nc) at (0, -4)
        {\vphantom{0}};
        \node[draw, circle, fill=uofgcobalt, inner sep=5pt, font=\bfseries] (Nd) at (2, -4)
        {\vphantom{0}};

        \draw [edge] (Na) -- (Nb);
        \draw [edge] (Nb) -- (Nc);
        \draw [edge] (Nc) -- (Nd);
        \draw [edge] (Nb) -- (Nd);

        \node[draw, circle, fill=uofgcobalt, inner sep=5pt, font=\bfseries] (N1) at (7.5,  0) {\vphantom{0}};
        \node[draw, circle, fill=uofgcobalt, inner sep=5pt, font=\bfseries] (N2) at (9.5,  0) {\vphantom{0}};
        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N3) at (7.5, -2) {\vphantom{0}};
        \node[draw, circle, fill=uofgcobalt, inner sep=5pt, font=\bfseries] (N4) at (9.5, -2) {\vphantom{0}};
        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N5) at (7.5, -4) {\vphantom{0}};
        \node[draw, circle, fill=uofgcobalt, inner sep=5pt, font=\bfseries] (N6) at (9.5, -4) {\vphantom{0}};
        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N7) at (5.5,  0) {\vphantom{0}};
        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N8) at (5.5, -2) {\vphantom{0}};
        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N9) at (5.5, -4) {\vphantom{0}};

        \draw [edge] (N1) -- (N2);
        \draw [ledge] (N1) -- (N3);
        \draw [edge] (N1) -- (N4);
        \draw [edge] (N2) -- (N4);
        \draw [ledge] (N3) -- (N5);
        \draw [ledge] (N3) -- (N6);
        \draw [edge] (N4) -- (N6);
        \draw [ledge] (N5) -- (N6);
        \draw [hedge] (N2) to [in=45, out=315] (N6);
        \draw [ledge] (N1) -- (N7);
        \draw [ledge] (N5) -- (N9);
        \draw [ledge] (N7) -- (N8);
        \draw [ledge] (N8) -- (N9);
    \end{tikzpicture}\end{center}
\end{wrapfigure}

The \textbf{non-induced subgraph isomorphism problem} is to find an injective mapping from a given
pattern graph to a given target graph which preserves adjacency---in essence, we are ``finding a
copy of'' the pattern inside the target.

\medskip

The \textbf{induced} variant of the problem additionally requires that the mapping preserve
non-adjacency, so there are no ``extra edges'' in the copy of the pattern that we find. The top
example is induced, whereas the bottom example is not, due to the dashed edge.

\medskip

Despite these problems being \NP-complete, modern practical subgraph isomorphism algorithms can
handle problem instances with many hundreds of vertices in the pattern graph, and up to \textbf{ten thousand
vertices} in the target graph, leading to successful application in areas such as \textbf{computer
vision}, \textbf{biochemistry}, and \textbf{pattern recognition}. However, these algorithms cannot
handle arbitrary instances of this size---here we generate instances with thirty and one hundred and
fifty vertices respectively which cannot be solved within one hour.
}

\block{The Expected Number of Solutions, and Heuristics}{
    If we randomly generate a pattern graph with $p$ vertices and density $d_p$, and a target graph
    with $t$ vertices and density $d_t$, the expected number of solutions to the non-induced
    isomorphism problem is \[
        \mathsf{\langle Sol \rangle = t \cdot (t - 1) \cdot \ldots \cdot (t - p + 1)\cdot {d_t}^{d_p \cdot
        \binom{p}{2}} } \textnormal{.} \]
}

\column{0.5}

\block{Non-Induced Phase Transitions}{
    If we generate random patterns and targets independently, fixing a pattern order of 20 vertices,
    a target order of 150 vertices, a target density of 0.4, and varying the pattern density, the
    familiar phase transition and (solver-independent) complexity peak occurs.

    \vspace*{1em}

    \begin{center}
        \small
        \input{gen-graph-phase-transition}
    \end{center}

    We can also look at what happens if we vary both the pattern and the target density, for
    different orders of pattern. The top row shows satisfiability---the black lines show where the
    expected number of solutions is one. The bottom row shows the search cost for one of the three
    algorithms we consider in the paper. As expected, the ``really hard'' instances are near the
    phase transition.

    \vspace*{1em}

    \begin{center}
        \small
        \input{gen-graph-non-induced}
    \end{center}
}

\end{columns}

\block{Induced Phase Transitions}{
\begin{wrapfigure}[21]{r}{0.62\linewidth}\begin{center}
    \small\input{gen-graph-induced}
\end{center}\end{wrapfigure}

    For induced isomorphisms, the expected number of solutions is \[
        \mathsf{\langle Sol \rangle = t \cdot (t - 1) \cdot \ldots \cdot (t - p + 1)\cdot
                    {d_t}^{d_p \cdot \binom{p}{2}} \cdot {(1 - d_{t})}^{(1 - d_{p}) \cdot \binom{p}{2}} } \textnormal{.} \]

    \vspace*{52ex}
}

\begin{columns}
\column{0.5}

\block{See the Paper For\ldots}{
    \begin{itemize}
        \item ~ Results from two other subgraph isomorphism algorithms.
        \item ~ Behaviour under reduction to clique, pseudo-boolean, and boolean satisfiability.
        \item ~ Should we alter heuristics based upon the expected number of solutions?
    \end{itemize}
}

\column{0.5}

\block{Future Work}{
    \begin{itemize}
        \item ~ Variance for more accurate predictions (if you can help, please get in touch).
        \item ~ Other random models (degree-bounded, scale-free, \ldots).
        \item ~ Labelled graphs.
    \end{itemize}
}

\end{columns}

{
    \colorlet{blockbodybgcolor}{uofgsandstone!80!white}
    \colorlet{blocktitlebgcolor}{uofgsandstone!80!white}
    \block[bodyverticalshift=-0.5cm]{}{
        This work was supported by the Engineering and Physical Sciences Research Council [grant
        number EP/K503058/1]. \hfill \texttt{c.mccreesh.1@research.gla.ac.uk}
    }
}

\end{document}

### 2016-05-11 poster/poster.tex
More postering

@@
 
-\medskip
+\bigskip
 
@@
 \block{The Expected Number of Solutions, and Heuristics}{
+
 If we randomly generate a pattern graph with $p$ vertices and density $d_p$, and a target graph
-    with $t$ vertices and density $d_t$, the expected number of solutions to the non-induced
+with $t$ vertices and density $d_t$, the \textbf{expected number of solutions} to the non-induced
 isomorphism problem is \[
@@
     \binom{p}{2}} } \textnormal{.} \]
+
+\bigskip
+
+By considering when $\mathsf{\langle Sol \rangle = 1}$, we can \textbf{estimate the location} of the
+satisfiable / unsatisfiable phase transition. (This is not entirely accurate due to variance,
+particularly when the pattern is very sparse or very dense).
+
+\bigskip
+
+We can also use this formula to recover \textbf{variable} and \textbf{value ordering heuristics}: we
+should make branching decisions which maximise the expected number of solutions during search. This
+tells us to prefer small domains and pattern vertices of high degree, but target vertices of low
+degree, which matches the empirically-determined heuristics used in state of the art solvers.
+
 }
@@
     We can also look at what happens if we vary both the pattern and the target density, for
-    different orders of pattern. The top row shows satisfiability---the black lines show where the
-    expected number of solutions is one. The bottom row shows the search cost for one of the three
+    different orders of pattern. The top row shows satisfiability---the black lines show where
+    $\mathsf{\langle Sol \rangle = 1}$. The bottom row shows the search cost for one of the three
     algorithms we consider in the paper. As expected, the ``really hard'' instances are near the
@@
 
-    \vspace*{1em}
+    \vspace*{0.9em}
 
@@
 \block{Induced Phase Transitions}{
-\begin{wrapfigure}[21]{r}{0.62\linewidth}\begin{center}
+\begin{wrapfigure}[24]{r}{0.64\linewidth}\begin{center}
     \small\input{gen-graph-induced}
@@
 
-    For induced isomorphisms, the expected number of solutions is \[
+    For induced isomorphisms, the behaviour is much richer. We can derive an expected number of
+    solutions of \[
         \mathsf{\langle Sol \rangle = t \cdot (t - 1) \cdot \ldots \cdot (t - p + 1)\cdot
@@
 
-    \vspace*{52ex}
+    \bigskip
+
+    Again this predicts a sharp phase transition. When the pattern is small, we see hard instances
+    near the phase transition, and easy instances elsewhere. However, when the pattern is larger,
+    the central unsatisfiable region \textbf{remains very hard}, even though it is no longer near a
+    phase transition.
+
+    \bigskip
+
+    This is not just a weakness of current subgraph isomorphism algorithms: the region is also hard
+    for pseudo-boolean and boolean satisfiability solvers, and under reduction to the clique
+    problem.
+
+    \bigskip
+
+    The central hard region \emph{is} predicted correctly by \textbf{constrainedness}, as we
+    show in the third row of plots. Although far from a phase transition, this region is only
+    slightly overconstrained.
+
+    \bigskip
+
+    Looking to maximise $\mathsf{\langle Sol \rangle}$ derives existing variable and value ordering
+    heuristics, in the case that both the pattern and target graphs are sparse. The formula suggests
+    that algorithms should swap heuristics when the pattern and target are dense---empirically, this
+    does indeed give an improvement. However, maximising the expected number of solutions is
+    \textbf{not enough to select the best heuristic} in two eighths of the search space (and
+    constrainedness does not help).
 }

### 2016-06-29 talk/talk.tex
Start IJCAI talk

--- 2015-10-20 data-presentation-talk/talk.tex
@@
 \usepackage{centernot}
-\usepackage{mathtools}
-\usepackage{listings}
 
@@
 \tikzset{bedge/.style={ultra thick}}
-\tikzset{edged/.style={color=screengrey, dashed}}
-\tikzset{edgel3/.style={color=uofgrose, ultra thick}}
 
@@
 \setbeamercolor*{structure}{fg=uofguniversityblue}
-\setbeamercolor*{palette primary}{use=structure,fg=black,bg=}
+\setbeamercolor*{palette primary}{use=structure,fg=black,bg=white}
 \setbeamercolor*{palette secondary}{use=structure,fg=white,bg=uofgcobalt}
@@
                 {\usebeamerfont{title}\usebeamercolor[white]{}\inserttitle}\\[0.1cm]
-                {\usebeamerfont{author}\usebeamercolor[white]{}\scriptsize\insertauthor}
+                {\usebeamerfont{author}\usebeamercolor[white]{}\insertauthor}
             \end{minipage}
@@
 }
-\makeatother
 
+\makeatletter
 \newenvironment{nearlyplainframe}[2][]{
-    \makeatletter
     \def\beamer@entrycode{\vspace*{-\headheight}}
-
     \setbeamertemplate{headline}
@@
     }
-    \makeatother
 
@@
 }
+\makeatother
 
@@
 
-\title[When is Finding Little Graphs Inside Big Graphs Hard?]{When is Finding Little Graphs \\ Inside Big Graphs Hard?}
-\author[Ciaran McCreesh, Patrick Prosser and James Trimble]{\textbf{Ciaran McCreesh}, Patrick Prosser and James Trimble}
+\title[Heuristics and Really Hard Instances for Subgraph Isomorphism Problems]{Heuristics and Really
+Hard Instances \\ for Subgraph Isomorphism Problems}
+\author[Ciaran McCreesh, Patrick Prosser and James Trimble]{\textbf{Ciaran McCreesh}, Patrick
+Prosser and \\ James Trimble}
 
@@
 \begin{nearlyplainframe}[t]{A Phase Transition}
-    \only<1>{
-        \input{gen-graph-phase-1}
-    }
 
-    \only<2>{
-        \input{gen-graph-phase-2}
-    }
-
-    \only<3>{
-        \input{gen-graph-phase-3}
-    }
+    \input{gen-graph-phase}
 
@@
 
-\begin{frame}{2D Phase Transitions}
+\begin{nearlyplainframe}[t]{In Two Dimensions?}
 
-    \only<1>{
-        \input{gen-graph-2d-1}
-    }
+\end{nearlyplainframe}
 
-    \only<2>{
-        \input{gen-graph-2d-2}
-    }
+\begin{frame}{See The Paper For\ldots}
 
-    \only<3>{
-        \input{gen-graph-2d-3}
-    }
+    \begin{itemize}
+        \item Is this behaviour solver-independent?
+        \item Estimating the phase transition location
+        \item Using this to rediscover variable and value ordering heuristics
+    \end{itemize}
 
@@
 
-\begin{nearlyplainframe}[t]{3D Phase Transitions?}
-
-    \only<1>{
-        \input{gen-graph-3d-1}
-    }
-
-    \only<2>{
-        \input{gen-graph-3d-2}
-    }
+\begin{nearlyplainframe}[t]{What About Induced?}
 
@@
 
-\begin{frame}{What About Induced?}
-
-    \only<1>{
-        \input{gen-graph-induced-2d-1}
-    }
-
-    \only<2>{
-        \input{gen-graph-induced-2d-2}
-    }
-
-    \only<3>{
-        \input{gen-graph-induced-2d-3}
-    }
-
-    \only<4>{
-        \input{gen-graph-induced-2d-4}
-    }
-
-    \only<5>{
-        \input{gen-graph-induced-2d-5}
-    }
-\end{frame}
-
-\begin{nearlyplainframe}[t]{Induced in 3D?}
-
-    \only<1>{
-        \input{gen-graph-induced-3d-1}
-    }
-
-    \only<2>{
-        \input{gen-graph-induced-3d-2}
-    }
+\begin{nearlyplainframe}[t]{What Changes Between 10 and 20?}
 
@@
 
-\begin{frame}{Other Solvers?}
+\begin{nearlyplainframe}[t]{Is The Central Region Really Hard?}
 
-    \only<1>{
-        \input{gen-graph-alt-1}
-    }
+\end{nearlyplainframe}
 
-    \only<2>{
-        \input{gen-graph-alt-2}
-    }
+\begin{nearlyplainframe}[t]{Constrainedness}
 
-    \only<3>{
-        \input{gen-graph-alt-3}
-    }
+\end{nearlyplainframe}
+
+\begin{frame}{See The Paper For\ldots}
+
+    \begin{itemize}
+        \item Is this behaviour solver-independent?
+            \begin{itemize}
+                \item Also SAT, PB, MIP encodings, reduction to clique.
+            \end{itemize}
+        \item Estimating the phase transition location
+        \item Using this to invent new variable and value ordering heuristics
+            \begin{itemize}
+                \item But something unexpected happens this time!
+            \end{itemize}
+    \end{itemize}
 
@@
 
-\begin{frame}{Reductions}
+\begin{frame}{Future Work}
 
-    \only<1>{
-        \input{gen-graph-red-1}
-    }
-
-    \only<2>{
-        \input{gen-graph-red-2}
-    }
-
-    \only<3>{
-        \input{gen-graph-red-3}
-    }
-
-    \only<4>{
-        \input{gen-graph-red-4}
-    }
-
-    \only<5>{
-        \input{gen-graph-red-5}
-    }
-
-\end{frame}
-
-\begin{frame}{Why the Weird Shape?}
-
-    \only<1>{
-        \input{gen-graph-shape-1}
-    }
-
-    \only<2>{
-        \input{gen-graph-shape-2}
-    }
-
-    \only<3>{
-        \input{gen-graph-shape-3}
-    }
-
-    \only<4>{
-        \input{gen-graph-shape-4}
-    }
-
-    \only<5>{
-        \input{gen-graph-shape-5}
-    }
-
-    \only<6>{
-        \input{gen-graph-shape-6}
-    }
+    \begin{itemize}
+        \item Other randomness models (bounded degree, regular, scale-free).
+        \item Better estimates of the phase transition location for very sparse or very dense
+            patterns.
+            \begin{itemize}
+                \item This needs a horrible variance calculation. Please get in touch if you like
+                    doing this sort of thing.
+            \end{itemize}
+        \item Dynamic heuristics?
+    \end{itemize}
 
@@
 \begin{frame}[plain,noframenumbering]
+    \begin{center}
+        \vspace*{4em}
+        \url{http://www.dcs.gla.ac.uk/~ciaran} \\
+        \vspace{1em}
+        \href{mailto:c.mccreesh.1@research.gla.ac.uk}{\nolinkurl{c.mccreesh.1@research.gla.ac.uk}} \\
+    \end{center}
+
     \begin{tikzpicture}[remember picture, overlay]
@@
             \begin{tikzpicture}[remember picture, overlay]
-                \fill [fill=uofguniversityblue, anchor=north west] (0, 0) rectangle (\paperwidth, -1.7cm);
+                \fill [fill=uofguniversityblue, anchor=north west] (0, 0) rectangle (\paperwidth, -2.6cm);
             \end{tikzpicture}
@@
 
-        \node (logo) [anchor=north east, shift={(-0.3cm,-0.2cm)}] at (current page.north east) {
-            \includegraphics*[keepaspectratio=true,scale=0.55]{UoG_keyline.pdf}
+        \node (logo) [anchor=north east, shift={(-0.6cm,-0.6cm)}] at (current page.north east) {
+            \includegraphics*[keepaspectratio=true,scale=0.7]{UoG_keyline.pdf}
         };

### 2016-07-04 talk/talk.tex
Animate
More talking
More talk

@@
 \newenvironment{nearlyplainframe}[2][]{
-    \def\beamer@entrycode{\vspace*{-\headheight}}
+    \def\beamer@entrycode{\vspace*{-\headheight}\vspace*{3pt}}
     \setbeamertemplate{headline}
@@
 
+\begin{frame}{Benchmarking}
+
+    \begin{itemize}
+        \item Based upon chemical and computer vision datasets, we can handle patterns with 1,000
+            vertices and targets with 10,000 vertices.
+        \item Do these results reflect the worst case, or are they too optimistic?
+        \item Can we create ``hard'' benchmark instances?
+    \end{itemize}
+
+\end{frame}
+
+\begin{frame}{Randomly Selected Subgraphs}
+
+    \begin{itemize}
+        \item Start with a random target graph.
+        \item Pick vertices at random to make a pattern.
+        \item Shuffle the numbering.
+        \item <2-> These instances will always be satisfiable!
+    \end{itemize}
+
+\end{frame}
+
+\begin{frame}{Independently Random Subgraphs}
+
+    \begin{itemize}
+        \item Make a random target graph.
+        \item Independently, make a random pattern graph.
+        \item <2-> Will these instances ever be satisfiable?
+    \end{itemize}
+
+\end{frame}
+
+\begin{nearlyplainframe}[t]{A Phase Transition}
+
+    \only<1> {
+        \input{gen-graph-phase}
+    }
+    \only<2> {
+        \input{gen-graph-phase-2}
+    }
+
+\end{nearlyplainframe}
+
+\begin{frame}{This Looks Familiar\ldots}
+    \centering
+    \includegraphics*[keepaspectratio=true,scale=0.25]{sat.jpg}
+
+    \par\flushleft
+    Understanding the Empirical Hardness of NP-Complete Problems. Kevin Leyton-Brown, Holger H.
+    Hoos, Frank Hutter, Lin Xu.  Communications of the ACM, Vol. 57 No. 5, Pages 98-107
+\end{frame}
+
+\begin{nearlyplainframe}[t]{In Two Dimensions?}
+    \only<1> {
+        \input{gen-graph-non-induced-2d}
+    }
+\end{nearlyplainframe}
+
+\begin{frame}{See The Paper For\ldots}
+
+    \begin{itemize}
+        \item Is this behaviour solver-independent?
+        \item Estimating the phase transition location.
+        \item Using this to rediscover variable and value ordering heuristics.
+    \end{itemize}
+
+\end{frame}
+
 \begin{frame}{Induced Subgraph Isomorphism}
@@
 
-\begin{frame}{Benchmarking}
-
-    \begin{itemize}
-        \item Based upon chemical and computer vision datasets, we can handle patterns with 1,000
-            vertices and targets with 10,000 vertices.
-        \item Do these results reflect the worst case, or are they too optimistic?
-        \item Can we create ``hard'' benchmark instances?
-    \end{itemize}
-
-\end{frame}
-
-\begin{frame}{Randomly Selected Subgraphs}
-
-    \begin{itemize}
-        \item Start with a random target graph.
-        \item Pick vertices at random to make a pattern.
-        \item Shuffle the numbering.
-        \item <2-> These instances will always be satisfiable!
-    \end{itemize}
-
-\end{frame}
-
-\begin{frame}{Independently Random Subgraphs}
-
-    \begin{itemize}
-        \item Make a random target graph.
-        \item Independently, make a random pattern graph.
-        \item <2-> Will these instances ever be satisfiable?
-    \end{itemize}
-
-\end{frame}
-
-\begin{nearlyplainframe}[t]{A Phase Transition}
-
-    \input{gen-graph-phase}
-
+\begin{nearlyplainframe}[t]{Induced in 2D}
+    \only<1> {
+        \input{gen-graph-induced-2d}
+    }
 \end{nearlyplainframe}
 
-\begin{frame}{This Looks Familiar\ldots}
-    \centering
-    \includegraphics*[keepaspectratio=true,scale=0.25]{sat.jpg}
+\begin{nearlyplainframe}[t]{What Changes Between 10 and 20?}
+    \only<1> {
+        \input{gen-graph-induced-2d-2}
+    }
+\end{nearlyplainframe}
 
-    \par\flushleft
-    Understanding the Empirical Hardness of NP-Complete Problems. Kevin Leyton-Brown, Holger H.
-    Hoos, Frank Hutter, Lin Xu.  Communications of the ACM, Vol. 57 No. 5, Pages 98-107
-\end{frame}
+\begin{nearlyplainframe}[t]{Is The Central Region Really Hard?}
+    \only<1> {
+        \input{gen-graph-induced-2d-middle}
+    }
+\end{nearlyplainframe}
 
-\begin{nearlyplainframe}[t]{In Two Dimensions?}
+\begin{nearlyplainframe}[t]{What About Encodings or Reductions?}
+    \only<1> {
+        \input{gen-graph-induced-2d-alt}
+    }
+\end{nearlyplainframe}
 
+\begin{nearlyplainframe}[t]{Constrainedness}
+    \only<1> {
+        \input{gen-graph-induced-2d-kappa}
+    }
 \end{nearlyplainframe}
@@
     \begin{itemize}
-        \item Is this behaviour solver-independent?
-        \item Estimating the phase transition location
-        \item Using this to rediscover variable and value ordering heuristics
-    \end{itemize}
-
-\end{frame}
-
-\begin{nearlyplainframe}[t]{What About Induced?}
-
-\end{nearlyplainframe}
-
-\begin{nearlyplainframe}[t]{What Changes Between 10 and 20?}
-
-\end{nearlyplainframe}
-
-\begin{nearlyplainframe}[t]{Is The Central Region Really Hard?}
-
-\end{nearlyplainframe}
-
-\begin{nearlyplainframe}[t]{Constrainedness}
-
-\end{nearlyplainframe}
-
-\begin{frame}{See The Paper For\ldots}
-
-    \begin{itemize}
-        \item Is this behaviour solver-independent?
-            \begin{itemize}
-                \item Also SAT, PB, MIP encodings, reduction to clique.
-            \end{itemize}
-        \item Estimating the phase transition location
-        \item Using this to invent new variable and value ordering heuristics
+        \item More on solver-independence and reductions.
+        \item Estimating the phase transition location.
+        \item Using this to invent new variable and value ordering heuristics.
             \begin{itemize}
